### 2011-11-13 tex/censored_catalog.tex
censored data: rearrange files
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@19905 03a3cea6-2c03-0410-ac9b-9271b1c66c29

% Copyright 2011 David W. Hogg (NYU).
% All rights reserved.

\documentclass[12pt]{article}

\newcommand{\dd}{\mathrm{d}}

\begin{document}\sloppy\sloppypar\raggedbottom

\textsl{This document is a draft dated 2011-11-09.  It incorporates
  direct contributions from Joseph~W.~Richards (Berkeley),
  David~W.~Hogg (NYU), James~P.~Long (Berkeley), and
  Daniel~Foreman-Mackey (NYU) and indirect contributions from many
  others.}

\vspace{1ex}

A photometric survey (think of it as operating in a single photometric
bandpass for now) scans over the celestial position of a particular
variable star at a large number of times $t_i$.  At each of these
times, the imaging data are analyzed by (not awesome) software, which
treats each scan as a \emph{completely independent} survey.  That is,
when analyzing the data from time $t_i$, none of the information from
any of the other times is used in any way.

At each time $t_i$, the star is either detected ($q_i=1$) by the
software or not ($q_i=0$).  When it is detected, the software returns
a (possibly bad) flux value $f_i$ and (likely bad) uncertainty
variance $s_i^2$.  When the source is \emph{not} detected ($q_i=0$),
nothing is reported---in this case, because the software is not
awesome, it doesn't see any reason to report anything at all, since on
the non-detect passes, it has no inkling that this patch of the sky is
interesting in any way.  Welcome to the desert of the real.

Beyond this, there are two additional issues.  The first is that each
night of imaging is different in an unreported and unknown way.  That
is, there is different transparency, sky brightness, and point-spread
function.  So the detection limit, or completeness level, or censoring
of the catalog is different every night.  The second issue is that
either because we are getting the data from a non-generous source, or
because the conditions change rapidly and unpredictably, we can't
analyze all the sources in a finite patch simultaneously.  For each
variable star of interest, we \emph{only} get the data on that star
itself.

To recap, all we get at each epoch $t_i$ is a bit $q_i$ and, when
$q_i=1$, two data $f_i$ and $s_i^2$.  We don't get any information
about the upper limits or detection thresholds at the times $t_i$ when
$q_i=0$, and we don't know anything about the data quality (catalog
veracity) at any epoch.  For notational convenience we assemble all
the data into an enormous set $D$ given by
\begin{eqnarray}\displaystyle
D &\equiv& \{D_i\}
\\
D_i &\equiv& (q_i, f_i, s_i^2)
\quad ,
\end{eqnarray}
where the $f_i$ and $s_i^2$ data will be randomly generated or
meaningless at times at which the $q_i=0$.

We don't have access to all the information we need to use responsibly
the detections and non-detections in model fitting and inference.  We
have to instantiate latent variables to stand in for the missing
information.  For example, since we don't know the censoring threshold
or completeness limit for any observation $i$, we have to make it a
model parameter $b_i$.  Because we don't believe the reported
uncertainty variances $s_i^2$, we have to add parameters that model
the departure of the true uncertainty variance $\sigma_i^2$ from the
reported variance, and so on.

For the limited purposes here, the variable star will be periodic with
angular frequency $\omega$.  The expected brightness $\mu_i$ of the
star at time $t_i$ will be given by a linear combination of periodic
functions something like
\begin{eqnarray}\displaystyle
\mu_i &=& \sum_k A_k\,x_k(t_i|\omega)
\quad ,
\end{eqnarray}
where the $A_k$ are coefficients and the $x_k$ are functions.  This
can be generalized easily.  In addition to this expected brightness,
we assume that there is some either model uncertainty or stochastic
variation, leading to a \emph{model variance} $s_\mu^2$ at each point
(assumed constant but easily generalized).

If a source is detected at epoch $t_i$, we assume that its observed
flux $f_i$ is related to the expected flux $\mu_i$ by a Gaussian with
variance that has measurement and model contributions
\begin{eqnarray}\displaystyle
p(f_i|q_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
  1 & \mbox{for $q_i=0$} \\
  N(f_i|\mu_i,s_\mu^2+\sigma_i^2) & \mbox{for $q_i=1$}
\end{array}\right.\label{eq:poff}
\\
\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, \cdots)
\\
I &\equiv& (\{t_i\}, \mbox{assumptions})
\quad ,
\end{eqnarray}
where we have made the prior information about observation times $t_i$
and other assumptions explicit in the prior information blob $I$, and
we have a parameter blob $\theta$, which we will specify completely
below.  Note that the flux datum is ignored when the source is
undetected.

Similarly for the reported uncertainty variance $s_i^2$.  We don't
assume that it is correct in any case, but when the source is
detected, we assume that it is connected to the true uncertainty
$\sigma_i$ by a probabilistic likelihood, also Gaussian
\begin{eqnarray}\displaystyle
p(s_i|q_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
  1 & \mbox{for $q_i=0$} \\
  N(s_i|\sigma_i,v_\sigma) & \mbox{for $q_i=1$}
\end{array}\right.
\\
\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, v_\sigma, \cdots)
\quad ,
\end{eqnarray}
where we have added a model parameter $v_\sigma$, which represents the
variance in the distribution of reported uncertainties given the true
uncertainty; implicitly $v_\sigma$ is a component of $\theta$.  Again,
we are ignoring the variance datum when the source is undetected.

Under the assumption that the inclusion (or not) of the source in the
catalog is based on (or very strongly related to) the measured flux
(rather than the true flux), the natural likelihood for the bit $q_i$,
on which the above likelihoods are conditioned is
\begin{eqnarray}\displaystyle
P(q_i|b_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
  \displaystyle\int_0^{b_i} N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=0$} \\
  \displaystyle\int_{b_i}^\infty N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=1$}
\end{array}\right.\label{eq:pofq}
\quad,
\end{eqnarray}
where the integrals end at an observed-flux limit $b_i$, which (as
mentioned above) is also a model parameter.  The motivation for this
expression is that it imagines all possible observations $f$ of the
star at time $t_i$ given the model and uncertainties, and finds the
fraction that would make the (imagined) detection cut.

We are going to product together these likelihoods (for $q_i$, $f_i$,
and $s_i^2$); this looks like it is using the flux twice, once in the
PDF for $f_i$ in expression (\ref{eq:poff}) and once in the
probability for $q_i$ in expression (\ref{eq:pofq}).  It is not,
however, because in the probability for $q_i$ the flux $f$ that
appears is an integration variable, not an observation or datum.  We
treat the detections or non-detections as being ``prior'' to the
measurement.  This is the weakest plank in our otherwise impenetrable
fortress of solitude.

We are going to treat the true detection limit $b_i$ for each
observation, and its true uncertainty $\sigma_i$---two parameters
\emph{per data point}---as nuisance parameters.  We want to
marginalize them out.  This requires priors on these.  Again we choose
Gaussians for simplicity, but let the means and variances of these
Gaussians be new model parameters---hyperparameters if you wish.
\begin{eqnarray}\displaystyle
p(D_i|\theta,I) &=&
  \int P(D_i|b_i,\sigma_i,\theta,I)\,p(b_i|\theta)\,p(\sigma_i|\theta)\,\dd b_i\,\dd\sigma_i
\\
p(D_i|b_i,\sigma_i,\theta,I) &=&
  P(q_i|b_i,\sigma_i,\theta,I)\,p(f_i|q_i,\sigma_i,\theta,I)\,p(s_i|q_i,\sigma_i,\theta,I)
\\
p(b_i|\theta) &=& N(b_i|B,V_B)
\\
p(\sigma_i|\theta) &=& N(\sigma_i|S,V_S)
\\
\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, v_\sigma, B, V_B, S, V_S)
\quad,
\end{eqnarray}
where we have (at last) explicitly assembled all the variable-star
parameters and hyperparameters into the big parameter vector $\theta$.
The huge sets of nuisance parameters $\{b_i\}$ and
$\{\sigma_i\}$ don't remain, because by integration we
removed them.

Finally and it goes without saying that if we treat the data points as
independent (as we are free to do, given the non-awesomeness of the
software data source), we have that
\begin{eqnarray}\displaystyle
p(D|\theta,I) &=& \prod_i p(D_i|\theta,I)
\quad.
\end{eqnarray}
This is the likelihood for variable-star parameters given the entire
data set (all the measurements and non-detections of this star from
all the epochs, as delivered by the untrustworthy robots).  It
``correctly'' or at least ``justifiably'' uses all of the information
available, without making strong assumptions about the survey or its
veracity.

\end{document}

### 2011-11-14 tex/ms.tex
Added manuscript TeX, figure, ASAS data files, and simulated data files
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@19907 03a3cea6-2c03-0410-ac9b-9271b1c66c29

--- 2011-11-13 tex/censored_catalog.tex
@@
-% Copyright 2011 David W. Hogg (NYU).
-% All rights reserved.
+% Define the document class. aastex for manuscript preparation, emulateapj for making it look polished.
+% \documentclass[12pt,preprint2]{aastex}
+%\documentclass[12pt,a4paper, iop,onecolumn,numberedappendix]{emulateapj}
+\documentclass[12pt,preprint]{aastex}
 
-\documentclass[12pt]{article}
+% Import the natbib package for bibliography inclusion. To get it to work right, first typeset the bibliography (BibTex), Command-Shift-B.
+% Then, typeset the document with LaTeX, Command-Shift-L or just Command-T.
+% To get proper reference citations using aastex (preprint manuscripts), first typeset as emulateapj with BibTex to create the BibTex 
+% files and then typeset again as aastex to get the manuscript with references.
+\usepackage{natbib}
+\usepackage{epsfig,graphicx,color}
+\usepackage{longtable}
+
+\bibliographystyle{apj}
+% A few colors to replace the defaults for certain link types
+\definecolor{orange}{cmyk}{0,0.4,0.8,0.2}
+\definecolor{darkorange}{rgb}{.71,0.21,0.01}
+\definecolor{darkgreen}{rgb}{.12,.54,.11}
+
+%-----------------------------------------------------------------------------
+% The hyperref package gives us a pdf with properly built
+% internal navigation ('pdf bookmarks' for the table of contents,
+% internal cross-reference links, web links for URLs, etc.)
+\usepackage{hyperref}
+\hypersetup{pdftex,  % needed for pdflatex
+  breaklinks=true,  % so long urls are correctly broken across lines
+  colorlinks=true,
+  urlcolor=blue,
+  linkcolor=darkorange,
+  citecolor=darkgreen,
+  }
+
+%Define this shorthand for scientific notation.
+\newcommand{\expnt}[2]{\ensuremath{#1 \times 10^{#2}}}   % scientific notation
+\newcommand{\x}{{\bf x}}
+\newcommand{\X}{{\bf X}}
+\newcommand{\y}{{\bf y}}
+\newcommand{\z}{{\bf z}}
+\newcommand{\new}{red}
+\newcommand{\ptr}{p_{\rm tr}}
+\newcommand{\pte}{p_{\rm te}}
+\newcommand{\Ex}{\mathbf{E}}
+\newcommand{\Prob}{\mathbf{P}}
+\newcommand{\PrfL}{\widehat{P}_{\rm{RF,}\mathcal{L}}(y|\x)}
+\newcommand{\PrfLx}{\widehat{P}_{\rm{RF,}\mathcal{L}\cup\x'}(y|\x)}
+\newcommand{\TreeL}{\theta_{b, \mathcal{L}}(y|\x)}
+\newcommand{\TreeLx}{\theta_{b, \mathcal{L}\cup\x'}(y|\x)}
+ \def\hipp {{\it Hipparcos~}}
 
@@
 
-\begin{document}\sloppy\sloppypar\raggedbottom
+% Commands for annotating the docs with fixme and inter-author notes.  See
+% below for how to disable these.
+%
+% Define a \fixme command to mark visually things needing fixing in the draft,
+% as well as similar commands for each author to leave initialed special
+% comments in the document.
+% For final printing or to simply disable these bright warnings, symlink
+% (there's a target 'dist_on' in the makefile that does this) the file
+% macros_state.tex to macros_off.tex
 
-\textsl{This document is a draft dated 2011-11-09.  It incorporates
-  direct contributions from Joseph~W.~Richards (Berkeley),
-  David~W.~Hogg (NYU), James~P.~Long (Berkeley), and
-  Daniel~Foreman-Mackey (NYU) and indirect contributions from many
-  others.}
+\newcommand{\fixme}[1] { \textcolor{red} {
+{\fbox{ {\bf FIX}
+\ensuremath{\blacktriangleright \blacktriangleright \blacktriangleright}}
+{\bf #1}
+\fbox{\ensuremath{\blacktriangleleft \blacktriangleleft \blacktriangleleft}
+} } } }
 
-\vspace{1ex}
+% And similarly, one (less jarring, with fewer symbols and no boldface) command
+% for each one of us to leave comments in the main text.
+\newcommand{\james}[1] { \textcolor{blue} {
+\ensuremath{\blacklozenge} {\bf james:}  {#1}
+\ensuremath{\blacklozenge} } }
+
+\newcommand{\hogg}[1] { \textcolor{darkorange} {
+\ensuremath{\blacksquare} {\bf hogg:}  {#1}
+\ensuremath{\blacksquare} } }
+
+\newcommand{\dan}[1] { \textcolor{darkgreen} {
+\ensuremath{\bigstar} {\bf dan:}  {#1}
+\ensuremath{\bigstar} } }
+
+\newcommand{\joey}[1] { \textcolor{red} {
+\ensuremath{\clubsuit} {\bf joey:}  {#1}
+\ensuremath{\clubsuit} } }%\documentclass[manuscript]{aastex}
+
+\begin{document}
+%%%%%%%% TITLE, AUTHORS, PUBLICATION STATUS, and ABSTRACT %%%%%%%%%
+
+\shorttitle{Fitting the undetected}
+\shortauthors{}
+\title{Fitting the undetected:  Light-curve feature estimation in the presence of carelessly reported non-detections}
+\author{
+Some wild \& crazy guys
+}
+%
+%\altaffiltext{1}{Astronomy Department, University of California, Berkeley, CA, 94720-7450, USA}
+%\altaffiltext{2}{Statistics Department, University of California, Berkeley, CA, 94720-7450, USA}
+%\altaffiltext{*}{E-mail: {\tt jwrichar@stat.berkeley.edu}}
+
+% Note the status (in preparation, MNRAS submitted date, ApJ accepted date, in press etc).
+\slugcomment{in preparation}
+
+\begin{abstract}
+In multi-epoch imaging surveys, faint variable sources will not be
+detectable at all epochs.  In principle, the best approach is to
+perform photometric measurements anyway, so that there is a datum
+required at at every epoch.  In practice, in most surveys we are
+provided at some epochs only with upper limits; in some not even upper
+limits but just the information that the star was not observed.  Here
+we demonstrate with real data on periodic variables from the All-Sky
+Automated Survey that these non-detections are nonetheless useful in
+fitting and parameter estimation; their inclusion improves both
+precision and accuracy.  One novel aspect of this work is that we do
+not require the data source to include accurate---or even
+any---information about the uncertainty variances on detections or the
+values of the upper limits; we find that we can model these as latent
+variables, even when they are different at every epoch.
+\end{abstract}
+
+%\keywords{stars: variables: general -- methods: data analysis -- methods: statistical -- techniques: photometric}
+
+\section{Introduction}
+\label{sec:intro}
+
+\joey{What people actually do, what problems that could cause, and what people should do}
 
@@
 itself.
+
+\joey{RRL end at 120 kpc in Sesar et al. paper}
+
+\joey{discuss in context of: parameter estimation (amplitude / period), discovery (we're gonna find a lot more variable sources if we take the non-detections into account), and classification; lossiness}
+
+\joey{Numbers of LCs affected: Stripe 82, ASAS}
+
+\section{Method}
+\label{sec:method}
 
@@
 P(q_i|b_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  \displaystyle\int_0^{b_i} N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=0$} \\
+  \displaystyle\int_{-\infty}^{b_i} N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=0$} \\
   \displaystyle\int_{b_i}^\infty N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=1$}
@@
 
+\subsection{Implementation}
+\label{ss:implemenation}
+
+\joey{Describe Python implementation here}
+
+\section{Experiments}
+\label{sec:experiments}
+
+\subsection{Simulating Faint Mira Variables}
+\label{ss:mirasim}
+
+In this first experiment, we begin with a well-observed Mira variable from the ASAS Catalog of Variable Stars (ACVS, \citealt{acvs}) , ASAS 235627-4947.2.
+This star has a pulsation period of 266.6286 days and a Lomb-Scargle amplitude of 2.38 mag (V band).
+Note there are no non-detections in the ASAS light curve.
+
+To simulate faint Mira stars from ASAS 235627-4947.2, we use the following procedure:
+\begin{enumerate}
+\item Convert the observed magnitudes, $m_i$, and errors, $s_{m,i}$ to fluxes, $f_i$ and flux errors, $s_{f,i}$ (we assume a V-band zero-point of $3.67\times10^{-9}$ erg/s/cm$^2$/\AA)
+\item For a given flux dimming parameter, $d$, sample the new fluxes, $\tilde{f}_i$, from a Gaussian distribution centered around $f_i / d$ with standard deviation sampled from the empirical ASAS distribution of $s_{f,i} | f_i$.
+\joey{I found that relationship to be linear with a slope $\approx 1$, which is probably BS since the ASAS mag errors are all crap.
+So I added a reference value to each flux error to ensure a mag limit $\approx$ 14.5 mag.}
+\item Fluxes that are not at least 5$\sigma$ above zero are denoted as non-detections and their flux estimates (and errors) are censored.
+\end{enumerate}
+Folded light curves of ASAS 235627-4947.2, dimmed by four different values of $d$, are plotted in Figure \ref{fig:mirasim}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{../plots/mira_simulated.pdf}
+\end{center}
+\caption{ Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.
+Blue tick marks along the bottom axis denote phases where there were non-detections.
+As the original flux measurements are dimmed by higher factors, the troughs of the sinusoidal light curve are censored, resulting in an incomplete view of the data.
+\label{fig:mirasim}}
+\end{figure*}
+
+\subsection{Results of Fitting Miras}
+
+\section{Results: ASAS Light Curves}
+\label{sec:results}
+
+We use the methodology of \cite{2011arXiv1106.2832R} to select the top ASAS Mira  and RR Lyrae, Fundamental Mode candidates.
+Using a posterior probability threshold of 0.8 gives us 1720 Mira  and 1029 RR Lyrae candidates.
+
+ \subsection{Mira Variables}
+
+ Show P - A relationship before and after using the method
+
+ \subsection{RR Lyrae Variables}
+
+  Show P - A relationship before and after using the method
+
+There is an RRL P - A relationship (strong linear anti-correlation)
+
+\section{Discussion}
+\label{sec:discussion}
+
+What people can do, starting from the data-taking procedure.
+
+limitations
+
+\acknowledgements
+
+\bibliography{non_detect}
+
 \end{document}

### 2012-02-21 tex/note_jwr_20120221.tex
added note from 02/21/2012
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@20280 03a3cea6-2c03-0410-ac9b-9271b1c66c29

% Define the document class. aastex for manuscript preparation, emulateapj for making it look polished.
% \documentclass[12pt,preprint2]{aastex}
%\documentclass[12pt,a4paper, iop,onecolumn,numberedappendix]{emulateapj}
\documentclass[12pt,preprint]{aastex}

% Import the natbib package for bibliography inclusion. To get it to work right, first typeset the bibliography (BibTex), Command-Shift-B.
% Then, typeset the document with LaTeX, Command-Shift-L or just Command-T.
% To get proper reference citations using aastex (preprint manuscripts), first typeset as emulateapj with BibTex to create the BibTex 
% files and then typeset again as aastex to get the manuscript with references.
\usepackage{natbib}
\usepackage{epsfig,graphicx,color}
\usepackage{longtable}

\bibliographystyle{apj}
% A few colors to replace the defaults for certain link types
\definecolor{orange}{cmyk}{0,0.4,0.8,0.2}
\definecolor{darkorange}{rgb}{.71,0.21,0.01}
\definecolor{darkgreen}{rgb}{.12,.54,.11}

%-----------------------------------------------------------------------------
% The hyperref package gives us a pdf with properly built
% internal navigation ('pdf bookmarks' for the table of contents,
% internal cross-reference links, web links for URLs, etc.)
\usepackage{hyperref}
\hypersetup{pdftex,  % needed for pdflatex
  breaklinks=true,  % so long urls are correctly broken across lines
  colorlinks=true,
  urlcolor=blue,
  linkcolor=darkorange,
  citecolor=darkgreen,
  }

%Define this shorthand for scientific notation.
\newcommand{\expnt}[2]{\ensuremath{#1 \times 10^{#2}}}   % scientific notation
\newcommand{\x}{{\bf x}}
\newcommand{\X}{{\bf X}}
\newcommand{\y}{{\bf y}}
\newcommand{\z}{{\bf z}}
\newcommand{\new}{red}
\newcommand{\ptr}{p_{\rm tr}}
\newcommand{\pte}{p_{\rm te}}
\newcommand{\Ex}{\mathbf{E}}
\newcommand{\Prob}{\mathbf{P}}
\newcommand{\PrfL}{\widehat{P}_{\rm{RF,}\mathcal{L}}(y|\x)}
\newcommand{\PrfLx}{\widehat{P}_{\rm{RF,}\mathcal{L}\cup\x'}(y|\x)}
\newcommand{\TreeL}{\theta_{b, \mathcal{L}}(y|\x)}
\newcommand{\TreeLx}{\theta_{b, \mathcal{L}\cup\x'}(y|\x)}
 \def\hipp {{\it Hipparcos~}}

\newcommand{\dd}{\mathrm{d}}

\newcommand{\fobs}{f_i^*}
\newcommand{\sobs}{s_i^*}

% Commands for annotating the docs with fixme and inter-author notes.  See
% below for how to disable these.
%
% Define a \fixme command to mark visually things needing fixing in the draft,
% as well as similar commands for each author to leave initialed special
% comments in the document.
% For final printing or to simply disable these bright warnings, symlink
% (there's a target 'dist_on' in the makefile that does this) the file
% macros_state.tex to macros_off.tex

\newcommand{\fixme}[1] { \textcolor{red} {
{\fbox{ {\bf FIX}
\ensuremath{\blacktriangleright \blacktriangleright \blacktriangleright}}
{\bf #1}
\fbox{\ensuremath{\blacktriangleleft \blacktriangleleft \blacktriangleleft}
} } } }

% And similarly, one (less jarring, with fewer symbols and no boldface) command
% for each one of us to leave comments in the main text.
\newcommand{\james}[1] { \textcolor{blue} {
\ensuremath{\blacklozenge} {\bf james:}  {#1}
\ensuremath{\blacklozenge} } }

\newcommand{\hogg}[1] { \textcolor{darkorange} {
\ensuremath{\blacksquare} {\bf hogg:}  {#1}
\ensuremath{\blacksquare} } }

\newcommand{\dan}[1] { \textcolor{darkgreen} {
\ensuremath{\bigstar} {\bf dan:}  {#1}
\ensuremath{\bigstar} } }

\newcommand{\joey}[1] { \textcolor{red} {
\ensuremath{\clubsuit} {\bf joey:}  {#1}
\ensuremath{\clubsuit} } }%\documentclass[manuscript]{aastex}

\begin{document}
%%%%%%%% TITLE, AUTHORS, PUBLICATION STATUS, and ABSTRACT %%%%%%%%%

\shorttitle{Note: Feb 8, 2012}
\shortauthors{}
\title{Note on Non-Detections Model}
\author{
Joseph W. Richards
}

% Note the status (in preparation, MNRAS submitted date, ApJ accepted date, in press etc).
\slugcomment{February 21, 2012}

\section{Formulating the Model}
\label{sec:model}

This note, based on the original note from JWR on Feb. 8, 2012, includes revisions from David Hogg, James Long, Nat Butler, and Josh Bloom.

As a reminder, we are modeling light curve data in the presence of non-detections, which are epochs of observation in which no detection of the source of interest was made.

\subsection{Preliminaries}
\label{ss:prelim}

For each astronomical object, a photometric survey measures a multi-epoch light curve over $N$ (typically unevenly spaced) epochs.  At each epoch, $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object in the presence of the noise from the sky and instrument.  In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, which is the brightest the object could have been given that it was not detected at a significant level by the software.

There are many reasons that a source might not show up in a catalog.  These include:
\begin{itemize}
\item low S/N of the source, due either to a higher noise level or a fainter signal,
\item the source falling outside of the detection window (e.g., near chip edge),
\item occulting of the source by an artifact of the detector (e.g., hot pixels, masked out), or
\item the source was out-shown by an intervening object (e.g., asteroid, comet, airplane, etc.).
\end{itemize}
Here, we present a statistical model that can be used to detect variable sources and model their variability using multi-epoch light curves containing epochs of non-detection.  Previous efforts to detect and model variability using multi-epoch photometry have ignored non-detections (REFs) or used them in an ad-hoc manner lacking statistical rigor (REFs).

In this paper, we will assume that $\fobs$ and $\sobs$ take on real-valued numbers in epochs for which the source was detected and receive the reference value {\tt NA} in epochs where no detection was made. For notational convenience we assemble all the light curve data into a data set $D$ given by
\begin{eqnarray}\displaystyle
D &\equiv& \{D_i\}
\\
D_i &\equiv& (\fobs, \sobs)
\quad ,
\end{eqnarray}
where $(\fobs, \sobs)$ are the light curve measurements  at $t_i$.  The goal, for each astronomical object, is to maximize the likelihood of the light curve data, $D$, given a set of model parameters that describe the variability of the object and the characteristics of the observations.

\subsection{The Light Curve Model}

To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with oscillation frequency $\omega$,
\begin{eqnarray}\displaystyle
\mu_i &=& \sum_{k=1}^K A_k\, \sin (2 \pi t_i \omega  k + \phi_k)
\quad ,
\end{eqnarray}
where $A_k$ is the amplitude of the $k^{\rm th}$ harmonic of frequency $\omega$ and $\phi_k$ is the relative phase offset of harmonic $k$.  In this analysis, we will fix the number of harmonics to $K=3$, though in principle we can use the data to choose the appropriate $K$ to capture the complexity of the light curve.  In addition to this expected brightness, we assume that there is model uncertainty or stochastic variation, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized).

Additionally, we need to instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, for epoch $i$.  The $b_i$ parameter is essential because it constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.  By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.   At each epoch, we connect the observed flux, $\fobs$ to the observable flux of the object, $f_i$, via the detection limit, $b_i$ by
\begin{eqnarray}\displaystyle
\fobs &=& \left\{\begin{array}{ll}
  f_i & \mbox{if $f_i \ge b_i$} \\
  \texttt{NA} & \mbox{if $f_i < b_i$} \\
\end{array} \right.
\end{eqnarray}
so that the observed flux is \texttt{NA} only when the true observable flux, $f_i$, is below the detection limit, $b_i$.  And similarly,
\begin{eqnarray}\displaystyle
\sobs &=& \left\{\begin{array}{ll}
  s_i & \mbox{if $f_i \ge b_i$} \\
  \texttt{NA} & \mbox{if $f_i < b_i$} \\
\end{array} \right.
\end{eqnarray}
so that the true observable standard error of the flux measurement is \texttt{NA} only when the true observable flux is below the detection limit.

Also, because we do not necessarily believe the reported measurement uncertainties, $\sobs$, we choose to introduce a parameter, $\sigma_i$, to represent the true uncertainty variance for the brightness measurement at epoch $i$.

Hence, our initial model consists of the parameter vectors $\{b_1,...,b_N\}$ and $\{\sigma_1,...,\sigma_N\}$, and the model parameters
\begin{eqnarray}\displaystyle
\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2) \quad ,
\end{eqnarray}
along with prior information about observation times, $\{t_1,...,t_N\}$ and other prior assumptions, which we make explicit by creating the prior information set
\begin{eqnarray}\displaystyle
I &\equiv& (\{t_i\}, \mbox{assumptions}) \quad .
\end{eqnarray}
Our goal is to write down the form of the likelihood of the data, $D$, given $\theta$ and $I$.  Then, for each light curve we can find the vector $\theta$ that maximizes the data likelihood.

\subsection{Statistical Model for Light Curves with Non-Detections}

We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma_i$) and model ($s_{\mu}$) contributions.  In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = f_i \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($f_i < b_i$).  To derive the likelihood of the observed $\fobs$, given $\sigma_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
\begin{eqnarray}\displaystyle
p(\fobs |\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
  N(f_i | \mu_i,  \sigma_i^2 + s_{\mu}^2)\,  \int_0^{f_i} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
  \int_{0}^{\infty} \int_{0}^{b_i} N(f_i | \mu_i, \sigma_i^2 + s_{\mu}^2)\, p(b_i | \theta)\, \dd f_i\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik}
\\
p(b_i|\theta) &=& N(b_i|B,V_B)
\label{eq:bprior}
\\
\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2, B, V_B) \quad ,
\end{eqnarray}
where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.  In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $f_i$, enforcing that the detection limit be fainter than the observed brightness.  Likewise, in the epochs for which no detection was made ($\fobs = \texttt{NA}$), we integrate the joint $(f_i, b_i)$ likelihood over all possible values of $b_i$ and over the unknown $f_i$ from $0$ to $b_i$, ensuring that the brightness (if it were able to be observed) be fainter than the detection limit.

In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made on that epoch.  Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.  If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle is estimable from the raw telescope images), we can replace Equation \ref{eq:bprior} by these epoch-wise distributions.  In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
\begin{eqnarray}\displaystyle
p(\fobs |\sigma_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
  N(f_i | \mu_i,  \sigma_i^2 + s_{\mu}^2)\,  I(f_i \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
 \int_{0}^{b_i} N(f_i | \mu_i, \sigma_i^2 + s_{\mu}^2)\, \dd f_i & \mbox{if $\fobs = \texttt{NA}$} \\
\end{array}\right.\label{eq:mlik_s}
\end{eqnarray}
where the boolean indictor function, $I(f_i \ge b_i)$, is 1 if $f_i \ge b_i$ and 0 otherwise.

Next, we model the reported standard error, $\sobs$, on the uncertainty of the magnitude measurement.  Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true uncertainty standard error, $\sigma_i$ through a truncated Gaussian likelihood, where the truncation enforces a non-negative variance for each observation.  Our likelihood of observed $\sobs$, given $\sigma_i$, $\theta$ and $I$, is
\begin{eqnarray}\displaystyle
p(\sobs | \sigma_i, \theta,I) &=& \left\{\begin{array}{ll}
  N(s_i|\sigma_i,v_\sigma) & \mbox{if $\sobs \ne \texttt{NA}$} \\
  \int_{0}^{\infty}  N(s_i | \sigma_i,v_\sigma)\,  \dd s_i & \mbox{if $\sobs = \texttt{NA}$}
\end{array}\right.\label{eq:slik}
\\
\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
\end{eqnarray}
where we have added a model parameter $v_\sigma$, which represents the
variance in the distribution of reported uncertainties given the true
uncertainty.  The only purpose of
including the likelihood in Equation \ref{eq:slik} to the model, in practice, to keep the $\sigma_i$ from
drifting very far away from the $s_i$, as set by the hyperparameter
$v_\sigma$.

Putting it all together, we can write down the likelihood for the data, $D_i$, on a single epoch, as
\begin{eqnarray}\displaystyle
p(D_i|\theta,I) &=& \int_0^{\infty} p(\fobs |\sigma_i,\theta,I)\, p(\sobs |\sigma_i,\theta,I)\, p(\sigma_i | \theta)\,\dd \sigma_i
\\
p(\sigma_i|\theta) &\propto& \sigma_i^{-1}
\\
\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
\end{eqnarray}
where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), and integrated out the nuisance parameter, $\sigma_i$.  We have assumed a Jeffrey's prior on $\sigma_i$, which is non-informative and invariant to reparametrization of the variance.  We could alternatively use an inverse-gamma prior, which takes two hyperparameters (and is conjugate in the case of a normal likelihood with unknown variance).  We have also assumed that $\fobs$ and $\sobs$ are conditionally independent given $\sigma_i$, $\theta$, and $I$.  This is a reasonable assumption in the case that a detection is made at epoch $t_i$.

Finally if we assume that the data collected at each epoch are independent given the model parameters, we have that
\begin{eqnarray}\displaystyle
p(D|\theta,I) &=& \prod_i p(D_i|\theta,I)
\quad.
\end{eqnarray}
This is the likelihood for the entire
data set (all the measurements and non-detections of this star from
all the epochs, as delivered by the survey) given the $2K + 5$ parameter vector $\theta$.  This model
``correctly'' or at least ``justifiably'' uses all of the information
available, without making strong assumptions about the survey or its
veracity.

\subsection{Detecting Variability}

We will describe here how we will do the likelihood ratio test of the model of constant flux to that of periodically varying flux.

\bibliography{non_detect}

\end{document}

### 2012-02-29 tex/ms.tex
integrated note into tex
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@20302 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
+\newcommand{\fobs}{f_i^*}
+\newcommand{\sobs}{s_i^*}
+
 % Commands for annotating the docs with fixme and inter-author notes.  See
@@
 \author{
-Some wild \& crazy guys
+TBD
 }
@@
 
-\joey{What people actually do, what problems that could cause, and what people should do}
+For each astronomical object, a photometric survey (think of it as operating in a single photometric bandpass for now) repeatedly scans over the celestial position of a source, recording a multi-epoch light curve over $N$ (typically unevenly spaced) epochs.
+At each epoch $i$, with associated time $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the root variance of the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object.
+In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, $b_i$, which is the brightest the object could have been given that it was not detected at a significant level by the software.
 
-A photometric survey (think of it as operating in a single photometric
-bandpass for now) scans over the celestial position of a particular
-variable star at a large number of times $t_i$.  At each of these
-times, the imaging data are analyzed by (not awesome) software, which
-treats each scan as a \emph{completely independent} survey.  That is,
-when analyzing the data from time $t_i$, none of the information from
-any of the other times is used in any way.
+There are many reasons that a source might not show up in a catalog.  These include:
+\begin{itemize}
+\item low S/N of the source, due either to a higher noise level (e.g., clouds) or a fainter signal (e.g. intrinsic dimming of the source),
+\item the source falling outside of the detection window (e.g., near chip edge),
+\item occulting of the source by an artifact of the detector (e.g., hot pixels, masked out), or
+\item the source was out-shone by an intervening object (e.g., asteroid, comet, variable star, airplane, etc.).
+\end{itemize}
+Here, we present a statistical model that can be used to detect variable sources and model their variability using multi-epoch light curves containing epochs of non-detection.
+Previous efforts to detect and model variability using multi-epoch photometry have typically ignored non-detections (REFs) or used them in an ad-hoc manner lacking statistical rigor (REFs).
+An exception is \citet{2009AJ....137.4400L}, who measure proper motions of sources in SDSS falling below the detection limit.
 
-At each time $t_i$, the star is either detected ($q_i=1$) by the
-software or not ($q_i=0$).  When it is detected, the software returns
-a (possibly bad) flux value $f_i$ and (likely bad) uncertainty
-variance $s_i^2$.  When the source is \emph{not} detected ($q_i=0$),
-nothing is reported---in this case, because the software is not
-awesome, it doesn't see any reason to report anything at all, since on
-the non-detect passes, it has no inkling that this patch of the sky is
-interesting in any way.  Welcome to the desert of the real.
+%A photometric survey (think of it as operating in a single photometric
+%bandpass for now) scans over the celestial position of a particular
+%variable star at a large number of times $t_i$.  At each of these
+%times, the imaging data are analyzed by (not awesome) software, which
+%treats each scan as a \emph{completely independent} survey.  That is,
+%when analyzing the data from time $t_i$, none of the information from
+%any of the other times is used in any way.
+%
+%At each time $t_i$, the star is either detected ($q_i=1$) by the
+%software or not ($q_i=0$).  When it is detected, the software returns
+%a (possibly bad) flux value $f_i$ and (likely bad) uncertainty
+%variance $s_i^2$.  When the source is \emph{not} detected ($q_i=0$),
+%nothing is reported---in this case, because the software is not
+%awesome, it doesn't see any reason to report anything at all, since on
+%the non-detect passes, it has no inkling that this patch of the sky is
+%interesting in any way.  Welcome to the desert of the real.
+%
+%Beyond this, there are two additional issues.  The first is that each
+%night of imaging is different in an unreported and unknown way.  That
+%is, there is different transparency, sky brightness, and point-spread
+%function.  So the detection limit, or completeness level, or censoring
+%of the catalog is different every night.  The second issue is that
+%either because we are getting the data from a non-generous source, or
+%because the conditions change rapidly and unpredictably, we can't
+%analyze all the sources in a finite patch simultaneously.  For each
+%variable star of interest, we \emph{only} get the data on that star
+%itself.
 
-Beyond this, there are two additional issues.  The first is that each
-night of imaging is different in an unreported and unknown way.  That
-is, there is different transparency, sky brightness, and point-spread
-function.  So the detection limit, or completeness level, or censoring
-of the catalog is different every night.  The second issue is that
-either because we are getting the data from a non-generous source, or
-because the conditions change rapidly and unpredictably, we can't
-analyze all the sources in a finite patch simultaneously.  For each
-variable star of interest, we \emph{only} get the data on that star
-itself.
+\joey{discuss in context of: parameter estimation (amplitude / period), discovery (we're gonna find a lot more variable sources if we take the non-detections into account), and classification; lossiness}
 
 \joey{RRL end at 120 kpc in Sesar et al. paper}
-
-\joey{discuss in context of: parameter estimation (amplitude / period), discovery (we're gonna find a lot more variable sources if we take the non-detections into account), and classification; lossiness}
 
@@
 
-To recap, all we get at each epoch $t_i$ is a bit $q_i$ and, when
-$q_i=1$, two data $f_i$ and $s_i^2$.  We don't get any information
-about the upper limits or detection thresholds at the times $t_i$ when
-$q_i=0$, and we don't know anything about the data quality (catalog
-veracity) at any epoch.  For notational convenience we assemble all
-the data into an enormous set $D$ given by
+\subsection{Preliminaries}
+\label{ss:prelim}
+
+For each epoch, $t_i$, the photometric light curve of an astronomical source contains an estimate the mean brightness, $\fobs$, and root-variance of the error in that estimate, $\sobs$.
+These quantities are estimated from pixelated images taken of the source.
+In this paper, we will assume that $\fobs$ and $\sobs$ take on real-valued numbers in epochs for which the source was detected and receive the reference value {\tt NA} in epochs where no detection was made.
+For notational convenience we assemble all the light curve data for one source into a data set $D$ given by
 \begin{eqnarray}\displaystyle
@@
 \\
-D_i &\equiv& (q_i, f_i, s_i^2)
+D_i &\equiv& (\fobs, \sobs)
 \quad ,
 \end{eqnarray}
-where the $f_i$ and $s_i^2$ data will be randomly generated or
-meaningless at times at which the $q_i=0$.
+where $(\fobs, \sobs)$ are the light curve measurements  at $t_i$.
+The goal, for each astronomical object, is to construct the likelihood, $f^n$, for the data $D$, given a set of model parameters that describe the variability of the object and the characteristics of the observations, and either to maximize the likelihood with respect to the model parameters or to use it in further inference.
 
-We don't have access to all the information we need to use responsibly
-the detections and non-detections in model fitting and inference.  We
-have to instantiate latent variables to stand in for the missing
-information.  For example, since we don't know the censoring threshold
-or completeness limit for any observation $i$, we have to make it a
-model parameter $b_i$.  Because we don't believe the reported
-uncertainty variances $s_i^2$, we have to add parameters that model
-the departure of the true uncertainty variance $\sigma_i^2$ from the
-reported variance, and so on.
+\subsection{The Light Curve Model}
 
-For the limited purposes here, the variable star will be periodic with
-angular frequency $\omega$.  The expected brightness $\mu_i$ of the
-star at time $t_i$ will be given by a linear combination of periodic
-functions something like
+To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with angular oscillation frequency $\omega$,
 \begin{eqnarray}\displaystyle
-\mu_i &=& \sum_k A_k\,x_k(t_i|\omega)
-\quad ,
+\mu_i &=& \sum_{k=1}^K A_k\, \sin (t_i \, \omega  k) + B_k\, \cos (t_i \, \omega  k)
+\quad ,\label{eq:fourier}
 \end{eqnarray}
-where the $A_k$ are coefficients and the $x_k$ are functions.  This
-can be generalized easily.  In addition to this expected brightness,
-we assume that there is some either model uncertainty or stochastic
-variation, leading to a \emph{model variance} $s_\mu^2$ at each point
-(assumed constant but easily generalized).
+where $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(B_k,A_k)$ is the relative phase offset of harmonic $k$.
+In this analysis, we will fix the number of harmonics to $K=3$, though in principle we can use the data to choose the appropriate $K$ to capture the complexity of the light curve.
+In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized).
 
-If a source is detected at epoch $t_i$, we assume that its observed
-flux $f_i$ is related to the expected flux $\mu_i$ by a Gaussian with
-variance that has measurement and model contributions
+Additionally, we need to instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.
+The $b_i$ parameter is essential because it constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.
+By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.
+At each epoch, we connect the observed flux, $\fobs$, to the latent variable, $f_i$, signifying the true observable flux of the object, via the detection limit, $b_i$ by
 \begin{eqnarray}\displaystyle
-p(f_i|q_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  1 & \mbox{for $q_i=0$} \\
-  N(f_i|\mu_i,s_\mu^2+\sigma_i^2) & \mbox{for $q_i=1$}
-\end{array}\right.\label{eq:poff}
+\fobs &=& \left\{\begin{array}{ll}
+  f_i & \mbox{if $f_i \ge b_i$} \\
+  \texttt{NA} & \mbox{if $f_i < b_i$} \\
+\end{array} \right.
+\end{eqnarray}
+so that the observed flux is \texttt{NA} only when the true observable flux, $f_i$, is below the detection limit, $b_i$.
+And similarly, we introduce a latent variable, $s_i$, for the true observable standard error,
+\begin{eqnarray}\displaystyle
+\sobs &=& \left\{\begin{array}{ll}
+  s_i & \mbox{if $f_i \ge b_i$} \\
+  \texttt{NA} & \mbox{if $f_i < b_i$} \\
+\end{array} \right.
+\end{eqnarray}
+so that the true observable standard error of the flux measurement is \texttt{NA} only when the true observable flux is below the detection limit.
+
+Also, because we do not necessarily believe the reported measurement uncertainties, $\sobs$, we choose to introduce a parameter, $\sigma_i$, to represent the true uncertainty variance for the brightness measurement at epoch $i$.
+
+Hence, our initial model consists of the parameter vectors $\{f_1,...,f_N\}$, $\{s_1,...,s_N\}$, $\{b_1,...,b_N\}$ and $\{\sigma_1,...,\sigma_N\}$, and the model parameters
+\begin{eqnarray}\displaystyle
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2) \quad ,
+\end{eqnarray}
+along with prior information about observation times, $\{t_1,...,t_N\}$ and other prior assumptions, which we make explicit by creating the prior information set
+\begin{eqnarray}\displaystyle
+I &\equiv& (\{t_i\}, \mbox{assumptions}) \quad .
+\end{eqnarray}
+Our goal is to write down the form of the likelihood of the data, $D$, given $\theta$ and $I$.  Then, for each light curve we can find the vector $\theta$ that maximizes the data likelihood.
+
+\subsection{Statistical Model for Light Curves with Non-Detections}
+
+We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma_i$) and model ($s_{\mu}$) contributions.
+In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = f_i \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($f_i < b_i$).
+To derive the likelihood of the observed $\fobs$, given $\sigma_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
+\begin{eqnarray}\displaystyle
+p(\fobs |\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
+  N(f_i | \mu_i,  \sigma_i^2 + s_{\mu}^2)\,  \int_0^{f_i} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
+  \int_{0}^{\infty} \int_{0}^{b_i} N(f_i | \mu_i, \sigma_i^2 + s_{\mu}^2)\, p(b_i | \theta)\, \dd f_i\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
+\end{array}\right.\label{eq:mlik}
 \\
-\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, \cdots)
+p(b_i|\theta) &=& N(b_i|B,V_B)
+\label{eq:bprior}
 \\
-I &\equiv& (\{t_i\}, \mbox{assumptions})
-\quad ,
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2, B, V_B) \quad ,
 \end{eqnarray}
-where we have made the prior information about observation times $t_i$
-and other assumptions explicit in the prior information blob $I$, and
-we have a parameter blob $\theta$, which we will specify completely
-below.  Note that the flux datum is ignored when the source is
-undetected.
+where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.
+In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $f_i$, enforcing that the detection limit be fainter than the observed brightness.
+Likewise, in the epochs for which no detection was made ($\fobs = \texttt{NA}$), we integrate the joint $(f_i, b_i)$ likelihood over all possible values of $b_i$ and over the unknown $f_i$ from $0$ to $b_i$, ensuring that the brightness (if it were able to be observed) be fainter than the detection limit.
 
-Similarly for the reported uncertainty variance $s_i^2$.  We don't
-assume that it is correct in any case, but when the source is
-detected, we assume that it is connected to the true uncertainty
-$\sigma_i$ by a probabilistic likelihood, also Gaussian
+In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made on that epoch.
+Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.
+If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle can be inferred from the raw telescope images), we can replace Equation \ref{eq:bprior} with a different distribution per epoch.
+In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
 \begin{eqnarray}\displaystyle
-p(s_i|q_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  1 & \mbox{for $q_i=0$} \\
-  N(s_i|\sigma_i,v_\sigma) & \mbox{for $q_i=1$}
-\end{array}\right.
+p(\fobs |\sigma_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
+  N(f_i | \mu_i,  \sigma_i^2 + s_{\mu}^2)\,  I(f_i \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
+ \int_{0}^{b_i} N(f_i | \mu_i, \sigma_i^2 + s_{\mu}^2)\, \dd f_i & \mbox{if $\fobs = \texttt{NA}$} \\
+\end{array}\right.\label{eq:mlik_s}
+\end{eqnarray}
+where the boolean indictor function, $I(f_i \ge b_i)$, is 1 if $f_i \ge b_i$ and 0 otherwise.
+
+Next, we model the reported standard error, $\sobs$, on the uncertainty of the magnitude measurement.
+Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true uncertainty standard error, $\sigma_i$ through a truncated Gaussian likelihood, where the truncation enforces a non-negative variance for each observation.
+Our likelihood of observed $\sobs$, given $\sigma_i$, $\theta$ and $I$, is
+\begin{eqnarray}\displaystyle
+p(\sobs | \sigma_i, \theta,I) &=& \left\{\begin{array}{ll}
+  N(s_i|\sigma_i,v_\sigma) & \mbox{if $\sobs \ne \texttt{NA}$} \\
+  \int_{0}^{\infty}  N(s_i | \sigma_i,v_\sigma)\,  \dd s_i & \mbox{if $\sobs = \texttt{NA}$}
+\end{array}\right.\label{eq:slik}
 \\
-\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, v_\sigma, \cdots)
-\quad ,
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
 \end{eqnarray}
+\joey{Need a distribution that is truncated at 0. Something better? }
 where we have added a model parameter $v_\sigma$, which represents the
 variance in the distribution of reported uncertainties given the true
-uncertainty; implicitly $v_\sigma$ is a component of $\theta$.  Again,
-we are ignoring the variance datum when the source is undetected.
+uncertainty.  The only purpose of 
+ including the likelihood in Equation \ref{eq:slik} to the model, in practice, to keep the $\sigma_i$ from
+drifting very far away from the $s_i$, as set by the hyperparameter
+$v_\sigma$.
 
-Under the assumption that the inclusion (or not) of the source in the
-catalog is based on (or very strongly related to) the measured flux
-(rather than the true flux), the natural likelihood for the bit $q_i$,
-on which the above likelihoods are conditioned is
+Putting it all together, we can write down the likelihood of the data, $D_i$, for the parameters $\theta$, on a single epoch, as
 \begin{eqnarray}\displaystyle
-P(q_i|b_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  \displaystyle\int_{-\infty}^{b_i} N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=0$} \\
-  \displaystyle\int_{b_i}^\infty N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=1$}
-\end{array}\right.\label{eq:pofq}
-\quad,
+p(D_i|\theta,I) &=& \int_0^{\infty} p(\fobs |\sigma_i,\theta,I)\, p(\sobs |\sigma_i,\theta,I)\, p(\sigma_i | \theta)\,\dd \sigma_i
+\\
+p(\sigma_i|\theta) &\propto& \sigma_i^{-1}
+\\
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
 \end{eqnarray}
-where the integrals end at an observed-flux limit $b_i$, which (as
-mentioned above) is also a model parameter.  The motivation for this
-expression is that it imagines all possible observations $f$ of the
-star at time $t_i$ given the model and uncertainties, and finds the
-fraction that would make the (imagined) detection cut.
+where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), and integrated out the nuisance parameter, $\sigma_i$.
+We have assumed a Jeffrey's prior on $\sigma_i$, which is non-informative and invariant to reparametrization of the variance.
+We could alternatively use an inverse-gamma prior, which takes two hyperparameters (and is conjugate in the case of a normal likelihood with unknown variance).
+We have also assumed that $\fobs$ and $\sobs$ are conditionally independent given $\sigma_i$, $\theta$, and $I$.
+This is a reasonable assumption in the case that a detection is made at epoch $t_i$.
 
-We are going to product together these likelihoods (for $q_i$, $f_i$,
-and $s_i^2$); this looks like it is using the flux twice, once in the
-PDF for $f_i$ in expression (\ref{eq:poff}) and once in the
-probability for $q_i$ in expression (\ref{eq:pofq}).  It is not,
-however, because in the probability for $q_i$ the flux $f$ that
-appears is an integration variable, not an observation or datum.  We
-treat the detections or non-detections as being ``prior'' to the
-measurement.  This is the weakest plank in our otherwise impenetrable
-fortress of solitude.
-
-We are going to treat the true detection limit $b_i$ for each
-observation, and its true uncertainty $\sigma_i$---two parameters
-\emph{per data point}---as nuisance parameters.  We want to
-marginalize them out.  This requires priors on these.  Again we choose
-Gaussians for simplicity, but let the means and variances of these
-Gaussians be new model parameters---hyperparameters if you wish.
-\begin{eqnarray}\displaystyle
-p(D_i|\theta,I) &=&
-  \int P(D_i|b_i,\sigma_i,\theta,I)\,p(b_i|\theta)\,p(\sigma_i|\theta)\,\dd b_i\,\dd\sigma_i
-\\
-p(D_i|b_i,\sigma_i,\theta,I) &=&
-  P(q_i|b_i,\sigma_i,\theta,I)\,p(f_i|q_i,\sigma_i,\theta,I)\,p(s_i|q_i,\sigma_i,\theta,I)
-\\
-p(b_i|\theta) &=& N(b_i|B,V_B)
-\\
-p(\sigma_i|\theta) &=& N(\sigma_i|S,V_S)
-\\
-\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, v_\sigma, B, V_B, S, V_S)
-\quad,
-\end{eqnarray}
-where we have (at last) explicitly assembled all the variable-star
-parameters and hyperparameters into the big parameter vector $\theta$.
-The huge sets of nuisance parameters $\{b_i\}$ and
-$\{\sigma_i\}$ don't remain, because by integration we
-removed them.
-
-Finally and it goes without saying that if we treat the data points as
-independent (as we are free to do, given the non-awesomeness of the
-software data source), we have that
+Finally if we assume that the data collected at each epoch are independent given the model parameters, we have that
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-This is the likelihood for variable-star parameters given the entire
+This is the likelihood for the entire
 data set (all the measurements and non-detections of this star from
-all the epochs, as delivered by the untrustworthy robots).  It
+all the epochs, as delivered by the survey) given the $2K + 5$ parameter vector $\theta$.  This model
 ``correctly'' or at least ``justifiably'' uses all of the information
@@
 veracity.
+
+\subsection{Detecting Variability}
+
+We will describe here how we will do the likelihood ratio test of the model of constant flux to that of periodically varying flux.
+
+\subsection{Mixture Model for $\fobs$ or $\sobs$}
+
+This may be too much for this paper, but will be useful in the presence of crappy data.
 

### 2012-03-22 tex/note_jpl_20120321/note_jpl_20120321.tex
version controlling tests of nat's software
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@20472 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\documentclass[10pt]{article}
\usepackage{verbatim, amsmath,amssymb,amsthm,graphicx}
\usepackage[margin=.5in,nohead,nofoot]{geometry}
\usepackage{sectsty}
\usepackage{float}
\sectionfont{\normalsize}
\subsectionfont{\small}

\title{}
\date{}
\author{}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\text{ }}
\newcommand{\argmax}[1]{\underset{#1}{\operatorname{argmax}}}
\newcommand{\minimax}[2]{\argmin{#1}\underset{#2}{\operatorname{max}}}
\newcommand{\bb}{\textbf{b}}

\newcommand{\Var}{\text{Var }}
\newcommand{\Cov}{\text{Cov }}

\newenvironment{my_enumerate}{
  \begin{enumerate}
    \setlength{\itemsep}{1pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}}{\end{enumerate}
}

% Alter some LaTeX defaults for better treatment of figures:
% See p.105 of [yas] elisp error!TeX Unbound'' for suggested values.
% See pp. 199-200 of Lamport's [yas] elisp error!LaTeX'' book for details.
%   General parameters, for ALL pages:
\renewcommand{\topfraction}{0.9}% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}% max fraction of floats at bottom
%   Parameters for TEXT pages (not float pages):
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}     % 2 may work better
\setcounter{dbltopnumber}{2}    % for 2-column pages
\renewcommand{\dbltopfraction}{0.9}% fit big float above 2-col. text
\renewcommand{\textfraction}{0.07}% allow minimal text w. figs
%   Parameters for FLOAT pages (not text pages):
\renewcommand{\floatpagefraction}{0.7}% require fuller float pages
% N.B.: floatpagefraction MUST be less than topfraction !!
\renewcommand{\dblfloatpagefraction}{0.7}% require fuller float pages

% remember to use [htp] or [htpb] for placement

\begin{document}
\section*{Initial Test of Nat's Code on Miras and RR Lyrae - JPL - March 21}

I ran Nat's censored code on all our Miras and RR Lyrae both with censored observations and without censored observations.

\subsection{Miras}
\begin{itemize}
\item of the 1720 Miras, 247 had different period estimates (by more than 1\%) after including the censored observations
\item from visual inspectation almost all these 247 had the correct period after using the censored data
\item for the 247, using censored data nearly always doubled the period estimate (see plots below)
\end{itemize}

\noindent For RR Lyrae:
\begin{itemize}
\item i like RR lyraes
\end{itemize}

- density of miras periods w / wo censored obs, density of rr lyrae periods w / w/o censored 
- trouble getting unnormalized amplitude out of code
- may be issue with selection of miras, other interesting patterns may arise if we run code on non-miras

\begin{figure}[h]
  \begin{center}
    \begin{includegraphics}[height=4in,width=4in]{50.pdf}
      \caption{Mira where without using censored data the period is half of the estimate using censored. \textbf{period w/o censored = half period with censored} is the most common scenario among the Miras.}
    \end{includegraphics}
  \end{center}
\end{figure}

\begin{figure}[h]
  \begin{center}
    \begin{includegraphics}[height=4in,width=4in]{52.pdf}
      \caption{Another Mira.}
    \end{includegraphics}
  \end{center}
\end{figure}

\begin{figure}[h]
  \begin{center}
    \begin{includegraphics}[height=4in,width=4in]{96.pdf}
      \caption{And a third Mira.}
    \end{includegraphics}
  \end{center}
\end{figure}

\end{document}

### 2012-03-22 tex/note_jpl_20120321/note_jpl_20120321.tex
finished version of this note
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@20473 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-I ran Nat's censored code on all our Miras and RR Lyrae both with censored observations and without censored observations.
+I ran Nat's censored code on all our Miras and 100 of our RR Lyrae both with censored observations and without censored observations. For Miras:
 
-\subsection{Miras}
 \begin{itemize}
@@
 \begin{itemize}
-\item i like RR lyraes
+\item tested 100 of the RR Lyrae, period was identical for both models for all 100 curves
+\item visual inspection of folded curves suggests we get period right very high fraction of the time
+\item perhaps these RR Lyrae are too easy
 \end{itemize}
 
-- density of miras periods w / wo censored obs, density of rr lyrae periods w / w/o censored 
-- trouble getting unnormalized amplitude out of code
-- may be issue with selection of miras, other interesting patterns may arise if we run code on non-miras
+\noindent I wasn't able to back amplitudes out of Nat's code so I can't make a period -- amplitude plot just yet.
+For Miras where the naive model gets half the correct period, the amplidutes are going to way undershoot the true value.
+When the period is right but a lot of censoring, the amplidues will be biased down a bit.
 
@@
     \begin{includegraphics}[height=4in,width=4in]{50.pdf}
-      \caption{Mira where without using censored data the period is half of the estimate using censored.
-\textbf{period w/o censored = half period with censored} is the most common scenario among the Miras.}
+      \caption{Mira where without using censored data the period is half of the estimate using censored. This is the issue with the majority of the 247 where naive and censored models disagree.}
     \end{includegraphics}
@@
 
+\begin{figure}[h]
+  \begin{center}
+    \begin{includegraphics}[height=4in,width=4in]{mira_period_density.pdf}
+      \caption{Densities of Mira periods using censored model and ignoring censoring. A few very low period estimates are clipped in the graph.\label{mira_period_plot}}
+    \end{includegraphics}
+  \end{center}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{includegraphics}[height=4in,width=4in]{4.pdf}
+      \caption{RR Lyrae. Period is correct for both models.}
+    \end{includegraphics}
+  \end{center}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{includegraphics}[height=4in,width=4in]{5.pdf}
+      \caption{Another RR Lyrae with period correct in both models. Notice that non-detections are more frequent at higher magnitudes as we expect.}
+    \end{includegraphics}
+  \end{center}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{includegraphics}[height=4in,width=4in]{7.pdf}
+      \caption{Another RR Lyrae with period correct in both models.
+Interesting censoring pattern.
+Perhaps the censoring threshold was at a low magnitude during the first two observing cycles and then increased.}
+    \end{includegraphics}
+  \end{center}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{center}
+    \begin{includegraphics}[height=4in,width=4in]{rrl_period_density.pdf}
+      \caption{RRL period densities are the same for the two models.\label{rrl_period_plot}}
+    \end{includegraphics}
+  \end{center}
+\end{figure}
+
 \end{document}

### 2012-04-09 py/censoring.py
last commits before moving to git
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/censored@20563 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -2,10 +2,16 @@ import numpy as np
 from scipy.optimize import fmin
 from scipy import special
 
-
-## this is weighted least squares, weights on points are given uncertainties
-## we ignore censored observations here
 def naive_opt(myData):
+    """ 
+    This model ignores uncertainty, just weighted least squares.
+
+    Inputs:
+        myData: data object
+
+    Ouput:
+        fitted model parameters
+    """
     ## compute weight matrix, remember errors are in SD, not var
     Wsqrt = np.diag(myData.known_errors[:,0])
     ## multiply predictors and response by weight matrix
@@ -17,13 +23,19 @@ def naive_opt(myData):
     return np.dot(XtXinv,XtY)
 
 
-## a simple version of our model where:
-## 1) sigmas = s's (i.e. reported standard deviations are correct)
-## 2) sigma_mu = 0 (no model uncertainty)
-## 3) known censoring threshold
-## compute_censored=False ignores censoring, use this to test
-## that optimization methods are working
 def simple_model_likelihood(params,myData,b,compute_censored):
+    """
+    Version of our model that treats reported uncertainties as true, sets model 
+    uncertainty to 0, and assumes we know the censoring threshold.
+    
+    Inputs:
+        params: model parameters to compute likelihood for
+        myData: data object
+        b: censoring threshold
+        compute_censored: should we compute the liklihood for censored observations
+    Output:
+        likelihood of model for params
+    """
     params = np.array(params).reshape((len(params),1))
     ## compute (-2.0 * log likelihood) for uncensored observations
     normalized_differences = (myData.known_response - np.dot(myData.known_predictors,params)) / myData.known_errors
@@ -44,6 +56,7 @@ class SimParameters:
         self.sigmas = np.ones(n).reshape((n,1))
         self.n = n
 
+## for running tests, evaluating code
 class SimData:
     def __init__(self,Params):
         p = Params.beta.shape[0] - 1
@@ -78,7 +91,8 @@ class Data:
         print "censored_predictors:"
         print self.censored_predictors
 
-    
+##def lc_to_data_object(time, signal, error, time_missing):
+
 
 if __name__ == "__main__":
     ## generate parameters and data for the model

### 2012-04-09 py/visualize.py

@@ -30,3 +30,41 @@ def plot_curve(tfe,true,naive,hrl,model):
         ptimes = np.linspace(tfe[:,0].min(),tfe[:,0].max(),1000)
         plt.plot(ptimes,model.hrl.mean(ptimes),'g-')
     plt.show()
+
+
+
+## shows folded and/or unfolded curves
+def plot_two_curves(tfe,censored_model,naive_model,
+                    plot_name="plot.pdf",in_title=""):
+    q1 = tfe[:,1] < 29
+    q0 = tfe[:,1] > 29
+    plt.figure(1)
+
+    plt.subplots_adjust(hspace=.7)
+    plt.subplot(311)
+    plt.title("Unfolded " + in_title)
+    plt.plot(tfe[q1,0],-tfe[q1,1],'r.')
+    plt.plot(tfe[q0,0],-np.repeat(tfe[q1,1].max(),q0.sum()),'b.')
+    plt.xlabel("Time")
+    plt.ylabel("Flux")
+
+    plt.subplot(312)
+    plt.title("Not using Censored. Period: " + repr(naive_model))
+    plt.plot((tfe[q1,0] % naive_model) / naive_model,
+             -tfe[q1,1],'r.')
+    plt.plot((tfe[q0,0] % naive_model) / naive_model,
+             -np.repeat(tfe[q1,1].max(),q0.sum()),'b.')
+    plt.xlabel("Phase")
+    plt.ylabel("Flux")
+
+    plt.subplot(313)
+    plt.title(" Using Censored. Period: " + repr(censored_model))
+    plt.plot((tfe[q1,0] % censored_model) / censored_model,
+             -tfe[q1,1],'r.')
+    plt.plot((tfe[q0,0] % censored_model) / censored_model,
+             -np.repeat(tfe[q1,1].max(),q0.sum()),'b.')
+    values = np.linspace(0,1,100)
+    plt.xlabel("Phase")
+    plt.ylabel("Flux")
+    plt.savefig(plot_name, format='pdf')
+    ##plt.show()

### 2012-04-09 tex/ms.tex
Added Nat's code (py/lomb_scargle_censor.py), my note from 20120221, and the current ms tex/pdf

--- 2011-11-14 tex/ms.tex
@@
 
-\shorttitle{Fitting the undetected}
+\shorttitle{Fitting the Undetected}
 \shortauthors{}
-\title{Fitting the undetected:  Light-curve feature estimation in the presence of carelessly reported non-detections}
+\title{Fitting the Undetected:  Light-curve Feature Estimation in the Presence of Non-detections}
 \author{
@@
 perform photometric measurements anyway, so that there is a datum
-required at at every epoch.  In practice, in most surveys we are
-provided at some epochs only with upper limits; in some not even upper
+recorded at at every epoch.  In practice, most surveys only provide
+detection upper limits at some epochs; in some not even upper
 limits but just the information that the star was not observed.  Here
 we demonstrate with real data on periodic variables from the All-Sky
-Automated Survey that these non-detections are nonetheless useful in
+Automated Survey (ASAS) that these non-detections are nonetheless useful in
 fitting and parameter estimation; their inclusion improves both
@@
 values of the upper limits; we find that we can model these as latent
-variables, even when they are different at every epoch.
+variables, even when they are different at every epoch.  Using realistic
+simulations of Mira variable stars, we obtain amplitude estimates that
+are {\bf XXX\%} more accurate by incorporating the non-detections into
+the analysis.  For  Mira and RR Lyrae variables observed by ASAS, our
+method obtains a tighter period - amplitude relationship than the standard
+technique.
 \end{abstract}
@@
 
+Surveys working in real-time tend to analyze each image with an eye to populating databases quickly.
+Unless a particular region of the sky gets special treatment (such as it is a known place of interest), discovery of variability happens by comparison of flux values at the database level.
+Analysis on such databases of transient/variables thus deals with censored data where there is often less information available that there should be.
+Even then, we might be able to infer what the thresholds were based on S/N measurements of objects neighboring the source of interest.
+
 \joey{What people actually do, what problems that could cause, and what people should do}
@@
 bandpass for now) scans over the celestial position of a particular
-variable star at a large number of times $t_i$.  At each of these
-times, the imaging data are analyzed by (not awesome) software, which
+variable star  a large number of times $t_i$.  At each of these
+times, the imaging data are analyzed by software, which
 treats each scan as a \emph{completely independent} survey.  That is,
@@
 \joey{discuss in context of: parameter estimation (amplitude / period), discovery (we're gonna find a lot more variable sources if we take the non-detections into account), and classification; lossiness}
+
+\joey{let's make a point to stress variable discovery}
 

### 2012-04-09 tex/note_jwr_20120221.tex
Added Nat's code (py/lomb_scargle_censor.py), my note from 20120221, and the current ms tex/pdf

@@
 For each astronomical object, a photometric survey measures a multi-epoch light curve over $N$ (typically unevenly spaced) epochs.
-At each epoch, $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object in the presence of the noise from the sky and instrument.
-In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, which is the brightest the object could have been given that it was not detected at a significant level by the software.
+At each epoch $i$, with associated time $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the root variance of the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object.
+In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, $b_i$, which is the brightest the object could have been given that it was not detected at a significant level by the software.
 
@@
 \item occulting of the source by an artifact of the detector (e.g., hot pixels, masked out), or
-\item the source was out-shown by an intervening object (e.g., asteroid, comet, airplane, etc.).
+\item the source was out-shone by an intervening object (e.g., asteroid, comet, variable star, airplane, etc.).
 \end{itemize}
 Here, we present a statistical model that can be used to detect variable sources and model their variability using multi-epoch light curves containing epochs of non-detection.
-Previous efforts to detect and model variability using multi-epoch photometry have ignored non-detections (REFs) or used them in an ad-hoc manner lacking statistical rigor (REFs).
+Previous efforts to detect and model variability using multi-epoch photometry have typically ignored non-detections (REFs) or used them in an ad-hoc manner lacking statistical rigor (REFs).
+An exception is \citet{2009AJ....137.4400L}, who measure proper motions of sources in SDSS falling below the detection limit.
 
 In this paper, we will assume that $\fobs$ and $\sobs$ take on real-valued numbers in epochs for which the source was detected and receive the reference value {\tt NA} in epochs where no detection was made.
-For notational convenience we assemble all the light curve data into a data set $D$ given by
+For notational convenience we assemble all the light curve data for one source into a data set $D$ given by
 \begin{eqnarray}\displaystyle
@@
 where $(\fobs, \sobs)$ are the light curve measurements  at $t_i$.
-The goal, for each astronomical object, is to maximize the likelihood of the light curve data, $D$, given a set of model parameters that describe the variability of the object and the characteristics of the observations.
+The goal, for each astronomical object, is to construct the likelihood, $f^n$, for the data $D$, given a set of model parameters that describe the variability of the object and the characteristics of the observations, and either to maximize the likelihood with respect to the model parameters or to use it in further inference.
 
@@
 
-To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with oscillation frequency $\omega$,
+To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with angular oscillation frequency $\omega$,
 \begin{eqnarray}\displaystyle
-\mu_i &=& \sum_{k=1}^K A_k\, \sin (2 \pi t_i \omega  k + \phi_k)
-\quad ,
+\mu_i &=& \sum_{k=1}^K A_k\, \sin (t_i \, \omega  k) + B_k\, \cos (t_i \, \omega  k)
+\quad ,\label{eq:fourier}
 \end{eqnarray}
-where $A_k$ is the amplitude of the $k^{\rm th}$ harmonic of frequency $\omega$ and $\phi_k$ is the relative phase offset of harmonic $k$.
+where $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(B_k,A_k)$ is the relative phase offset of harmonic $k$.
 In this analysis, we will fix the number of harmonics to $K=3$, though in principle we can use the data to choose the appropriate $K$ to capture the complexity of the light curve.
-In addition to this expected brightness, we assume that there is model uncertainty or stochastic variation, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized).
+In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized).
 
-Additionally, we need to instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, for epoch $i$.
+Additionally, we need to instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.
 The $b_i$ parameter is essential because it constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.
 By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.
-At each epoch, we connect the observed flux, $\fobs$ to the observable flux of the object, $f_i$, via the detection limit, $b_i$ by
+At each epoch, we connect the observed flux, $\fobs$, to the latent variable, $f_i$, signifying the true observable flux of the object, via the detection limit, $b_i$ by
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-so that the observed flux is \texttt{NA} only when the true observable flux, $f_i$, is below the detection limit, $b_i$.  And similarly,
+so that the observed flux is \texttt{NA} only when the true observable flux, $f_i$, is below the detection limit, $b_i$.
+And similarly, we introduce a latent variable, $s_i$, for the true observable standard error,
 \begin{eqnarray}\displaystyle
@@
 
-Hence, our initial model consists of the parameter vectors $\{b_1,...,b_N\}$ and $\{\sigma_1,...,\sigma_N\}$, and the model parameters
+Hence, our initial model consists of the parameter vectors $\{f_1,...,f_N\}$, $\{s_1,...,s_N\}$, $\{b_1,...,b_N\}$ and $\{\sigma_1,...,\sigma_N\}$, and the model parameters
 \begin{eqnarray}\displaystyle
-\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2) \quad ,
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2) \quad ,
 \end{eqnarray}
@@
 \\
-\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2, B, V_B) \quad ,
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2, B, V_B) \quad ,
 \end{eqnarray}
@@
 Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.
-If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle is estimable from the raw telescope images), we can replace Equation \ref{eq:bprior} by these epoch-wise distributions.
+If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle can be inferred from the raw telescope images), we can replace Equation \ref{eq:bprior} with a different distribution per epoch.
 In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
@@
 \\
-\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
 \end{eqnarray}
+\joey{Need a distribution that is truncated at 0. Something better? }
 where we have added a model parameter $v_\sigma$, which represents the
@@
 
-Putting it all together, we can write down the likelihood for the data, $D_i$, on a single epoch, as
+Putting it all together, we can write down the likelihood of the data, $D_i$, for the parameters $\theta$, on a single epoch, as
 \begin{eqnarray}\displaystyle
@@
 \\
-\theta &\equiv& (\omega, \{A_k\}, \{\phi_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
+\theta &\equiv& (\omega, \{A_k\}, \{B_k\}, s_\mu^2, B, V_B ,v_\sigma) \quad ,
 \end{eqnarray}
@@
 
+\subsection{Mixture Model for $\fobs$ or $\sobs$}
+
+This may be too much for this paper, but will be useful in the presence of crappy data.
+
 \bibliography{non_detect}

### 2012-04-09 py/lomb_scargle_censor.py

@@ -1,0 +1,176 @@
+from numpy import empty,pi,sqrt,sin,cos,var,dot,array,int,clip
+from scipy import weave
+from scipy.optimize import fmin
+
+def lomb(time, signal, error, time_missing, f1, df, numf, err_thresh=0.3, err_zpt=0.01):
+    """
+    C version of lomb_scargle
+
+    Inputs:
+        time: time vector
+        signal: data vector
+        error: data uncertainty vector
+        time_missing: time vector for missing observations
+        df: frequency step
+        numf: number of frequencies to consider
+        err_thresh: uncertainty of data not observed
+        err_zpt: uncertainty in the zero-point, corresponding to uncertaintly in the limit prior
+
+    Output:
+        psd: power spectrum on frequency grid: f1,f1+df,...,f1+numf*df
+    """
+    numt = len(time)
+    numt_missing = len(time_missing)
+
+    wth = (1./error).astype('float64')
+    s0 = dot(wth,wth)
+    wth /= sqrt(s0)
+
+    cn = (signal*wth).astype('float64')
+    cn -= dot(cn,wth)*wth
+
+    tt = 2*pi*time.astype('float64')
+    sinx,cosx = sin(tt*f1)*wth,cos(tt*f1)*wth
+    wpi = sin(df*tt); wpr=sin(0.5*df*tt); wpr = -2.*wpr*wpr
+
+    tt = 2*pi*time_missing.astype('float64')
+    sinx1,cosx1 = sin(tt*f1),cos(tt*f1)
+    wpi1 = sin(df*tt); wpr1=sin(0.5*df*tt); wpr1 = -2.*wpr1*wpr1
+
+    b0 = (cn/wth).max()
+    thresh = array([b0,err_zpt,sqrt(err_thresh**2+err_zpt**2)],dtype='float64')
+
+    # lookup tables for error function and it's derivatives
+    #from scipy import derivative
[…]

### 2012-04-10 tex/ms.tex
revised manuscript, implementing changes discussed on 04/09 with hogg and jpl

@@
 \newcommand{\dd}{\mathrm{d}}
+
+\newcommand{\fobs}{f_i}
+\newcommand{\sobs}{s^2_i}
+
+\newcommand{\ftrue}{f_i^*}
 
@@
 
-\section{Method}
-\label{sec:method}
+\section{Formulating the Model}
+\label{sec:model}
 
-To recap, all we get at each epoch $t_i$ is a bit $q_i$ and, when
-$q_i=1$, two data $f_i$ and $s_i^2$.  We don't get any information
-about the upper limits or detection thresholds at the times $t_i$ when
-$q_i=0$, and we don't know anything about the data quality (catalog
-veracity) at any epoch.  For notational convenience we assemble all
-the data into an enormous set $D$ given by
+%This note, based on the original note from JWR on Feb. 8, 2012, includes revisions from David Hogg, James Long, Nat Butler, and Josh Bloom.
+
+As a reminder, we are modeling light curve data in the presence of non-detections, which are epochs of observation in which no detection of the source of interest was made.
+
+\subsection{Preliminaries}
+\label{ss:prelim}
+
+For each astronomical object, a photometric survey measures a multi-epoch light curve over $N$ (typically unevenly spaced) epochs.
+At each epoch $i$, with associated time $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the  variance of the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object.
+In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, $b_i$, which is the brightest the object could have been given that it was not detected at a significant level by the software.
+
+There are many reasons that a source might not show up in a catalog.  These include:
+\begin{itemize}
+\item low S/N of the source, due either to a higher noise level or a fainter signal,
+\item the source falling outside of the detection window (e.g., near chip edge),
+\item occulting of the source by an artifact of the detector (e.g., hot pixels, masked out), or
+\item the source was out-shone by an intervening object (e.g., asteroid, comet, variable star, airplane, etc.).
+\end{itemize}
+Here, we present a statistical model that can be used to detect variable sources and model their variability using multi-epoch light curves containing epochs of non-detection.
+Previous efforts to detect and model variability using multi-epoch photometry have typically ignored non-detections (REFs) or used them in an ad-hoc manner lacking statistical rigor (REFs).
+An exception is \citet{2009AJ....137.4400L}, who measure proper motions of sources in SDSS falling below the detection limit.
+
+In this paper, we will assume that $\fobs$ and $\sobs$ take on real-valued numbers in epochs for which the source was detected and receive the reference value {\tt NA} in epochs where no detection was made.
+For notational convenience we assemble all the light curve data for one source into a data set $D$ given by
 \begin{eqnarray}\displaystyle
-D &\equiv& \{D_i\}
+D &\equiv& \{D_i\}_{i=1}^N
 \\
-D_i &\equiv& (q_i, f_i, s_i^2)
+D_i &\equiv& (\fobs, \sobs)
 \quad ,
 \end{eqnarray}
-where the $f_i$ and $s_i^2$ data will be randomly generated or
-meaningless at times at which the $q_i=0$.
+where $(\fobs, \sobs)$ are the light curve measurements  at $t_i$.
+The goal, for each astronomical object, is to construct the likelihood for the data $D$, given a set of model parameters that describe the variability of the object and the characteristics of the observations, and then to maximize the likelihood with respect to the model parameters.% or to use it in further inference.
 
-We don't have access to all the information we need to use responsibly
-the detections and non-detections in model fitting and inference.  We
-have to instantiate latent variables to stand in for the missing
-information.  For example, since we don't know the censoring threshold
-or completeness limit for any observation $i$, we have to make it a
-model parameter $b_i$.  Because we don't believe the reported
-uncertainty variances $s_i^2$, we have to add parameters that model
-the departure of the true uncertainty variance $\sigma_i^2$ from the
-reported variance, and so on.
+\subsection{The Light Curve Model}
 
-For the limited purposes here, the variable star will be periodic with
-angular frequency $\omega$.  The expected brightness $\mu_i$ of the
-star at time $t_i$ will be given by a linear combination of periodic
-functions something like
+To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with angular oscillation frequency $\omega$,
 \begin{eqnarray}\displaystyle
-\mu_i &=& \sum_k A_k\,x_k(t_i|\omega)
-\quad ,
+\mu_i &=& A_0 + \sum_{k=1}^K A_k\, \sin (t_i \, \omega  k) + B_k\, \cos (t_i \, \omega  k)
+\quad ,\label{eq:fourier}
 \end{eqnarray}
-where the $A_k$ are coefficients and the $x_k$ are functions.  This
-can be generalized easily.  In addition to this expected brightness,
-we assume that there is some either model uncertainty or stochastic
-variation, leading to a \emph{model variance} $s_\mu^2$ at each point
-(assumed constant but easily generalized).
+where $\mu_i$ is the flux at time $t_i$, $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(B_k,A_k)$ is the relative phase offset of harmonic $k$.
+The number of harmonics, $K$, can either be fixed or treated as a free parameter over which to optimize the likelihood.
+In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized).
 
-If a source is detected at epoch $t_i$, we assume that its observed
-flux $f_i$ is related to the expected flux $\mu_i$ by a Gaussian with
-variance that has measurement and model contributions
+Additionally, we need to instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.
+The $b_i$ parameter is essential because it constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.
+By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.
+At each epoch, we connect the observed flux, $\fobs$, to the latent variable, $\ftrue$, signifying the true observable flux of the object, via the detection limit, $b_i$ by
 \begin{eqnarray}\displaystyle
-p(f_i|q_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  1 & \mbox{for $q_i=0$} \\
-  N(f_i|\mu_i,s_\mu^2+\sigma_i^2) & \mbox{for $q_i=1$}
-\end{array}\right.\label{eq:poff}
-\\
-\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, \cdots)
-\\
-I &\equiv& (\{t_i\}, \mbox{assumptions})
-\quad ,
+\fobs &=& \left\{\begin{array}{ll}
+  \ftrue & \mbox{if $\ftrue \ge b_i$} \\
+  \texttt{NA} & \mbox{if $\ftrue < b_i$} \\
+\end{array} \right.
 \end{eqnarray}
-where we have made the prior information about observation times $t_i$
-and other assumptions explicit in the prior information blob $I$, and
-we have a parameter blob $\theta$, which we will specify completely
-below.  Note that the flux datum is ignored when the source is
-undetected.
+so that the observed flux is \texttt{NA} only when the true observable flux, $\ftrue$, is below the detection limit, $b_i$.
 
-Similarly for the reported uncertainty variance $s_i^2$.  We don't
-assume that it is correct in any case, but when the source is
-detected, we assume that it is connected to the true uncertainty
-$\sigma_i$ by a probabilistic likelihood, also Gaussian
+Also, because we do not necessarily believe the reported measurement uncertainties, $\sobs$, we choose to introduce a parameter, $\sigma^2_i$, to represent the true uncertainty variance for the brightness measurement at epoch $i$.
+
+Hence, our initial model consists of the parameter vectors $\{f^*_1,...,f^*_N\}$, $\{b_1,...,b_N\}$ and $\{\sigma^2_1,...,\sigma^2_N\}$, and the model parameters
 \begin{eqnarray}\displaystyle
-p(s_i|q_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  1 & \mbox{for $q_i=0$} \\
-  N(s_i|\sigma_i,v_\sigma) & \mbox{for $q_i=1$}
-\end{array}\right.
-\\
-\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, v_\sigma, \cdots)
-\quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s_\mu^2, \cdots) \quad ,
 \end{eqnarray}
-where we have added a model parameter $v_\sigma$, which represents the
-variance in the distribution of reported uncertainties given the true
-uncertainty; implicitly $v_\sigma$ is a component of $\theta$.  Again,
-we are ignoring the variance datum when the source is undetected.
+along with prior information about observation times, $\{t_1,...,t_N\}$ and other prior assumptions \joey{we should be more explicit here}, which we make explicit by creating the prior information set
+\begin{eqnarray}\displaystyle
+I &\equiv& (\{t_i\}_{i=1}^N, \mbox{assumptions}) \quad .
+\end{eqnarray}
+Our goal is to write down the form of the likelihood of the data, $D$, given $\theta$ and $I$.
+Then, for each light curve we can search for the vector $\theta$ that maximizes the data likelihood, and use those maximum likelihood estimates for downstream astrophysical inference.
 
-Under the assumption that the inclusion (or not) of the source in the
-catalog is based on (or very strongly related to) the measured flux
-(rather than the true flux), the natural likelihood for the bit $q_i$,
-on which the above likelihoods are conditioned is
+\subsection{Statistical Model for Light Curves with Non-Detections}
+
+We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma^2_i$) and model ($s_{\mu}$) contributions.
+In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = \ftrue \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($\ftrue < b_i$).
+To derive the likelihood of the observed $\fobs$, given $\sigma^2_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
 \begin{eqnarray}\displaystyle
-P(q_i|b_i,\sigma_i,\theta,I) &=& \left\{\begin{array}{ll}
-  \displaystyle\int_{-\infty}^{b_i} N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=0$} \\
-  \displaystyle\int_{b_i}^\infty N(f|\mu_i,s_\mu^2+\sigma_i^2)\,\dd f & \mbox{for $q_i=1$}
-\end{array}\right.\label{eq:pofq}
-\quad,
-\end{eqnarray}
-where the integrals end at an observed-flux limit $b_i$, which (as
-mentioned above) is also a model parameter.  The motivation for this
-expression is that it imagines all possible observations $f$ of the
-star at time $t_i$ given the model and uncertainties, and finds the
-fraction that would make the (imagined) detection cut.
-
-We are going to product together these likelihoods (for $q_i$, $f_i$,
-and $s_i^2$); this looks like it is using the flux twice, once in the
-PDF for $f_i$ in expression (\ref{eq:poff}) and once in the
-probability for $q_i$ in expression (\ref{eq:pofq}).  It is not,
-however, because in the probability for $q_i$ the flux $f$ that
-appears is an integration variable, not an observation or datum.  We
-treat the detections or non-detections as being ``prior'' to the
-measurement.  This is the weakest plank in our otherwise impenetrable
-fortress of solitude.
-
-We are going to treat the true detection limit $b_i$ for each
-observation, and its true uncertainty $\sigma_i$---two parameters
-\emph{per data point}---as nuisance parameters.  We want to
-marginalize them out.  This requires priors on these.  Again we choose
-Gaussians for simplicity, but let the means and variances of these
-Gaussians be new model parameters---hyperparameters if you wish.
-\begin{eqnarray}\displaystyle
-p(D_i|\theta,I) &=&
-  \int P(D_i|b_i,\sigma_i,\theta,I)\,p(b_i|\theta)\,p(\sigma_i|\theta)\,\dd b_i\,\dd\sigma_i
-\\
-p(D_i|b_i,\sigma_i,\theta,I) &=&
-  P(q_i|b_i,\sigma_i,\theta,I)\,p(f_i|q_i,\sigma_i,\theta,I)\,p(s_i|q_i,\sigma_i,\theta,I)
+p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
+  N(\ftrue | \mu_i,  \sigma^2_i + s_{\mu}^2)\,  \int_0^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
+  \int_{0}^{\infty} \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s_{\mu}^2)\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
+\end{array}\right.\label{eq:mlik}
 \\
 p(b_i|\theta) &=& N(b_i|B,V_B)
+\label{eq:bprior}
 \\
-p(\sigma_i|\theta) &=& N(\sigma_i|S,V_S)
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s_\mu^2, B, V_B, \cdots) \quad ,
+\end{eqnarray}
+where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.
+In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $\ftrue$, enforcing that the detection limit be fainter than the observed brightness.
+Likewise, in the epochs for which no detection was made ($\fobs = \texttt{NA}$), we integrate the joint $(\ftrue, b_i)$ likelihood over all possible values of $b_i$ and over the unknown $\ftrue$ from $0$ to $b_i$, ensuring that the brightness (if it were able to be observed) be fainter than the detection limit.
+
+In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made on that epoch.
+Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.
+If instead, we are given an estimate of $b_i$ plus its error distribution for each epoch (which, in principle can be inferred from the raw telescope images), we can replace Equation \ref{eq:bprior} with a different distribution per epoch.
+In the case that the $b_i$ are assumed to be completely known (without error), the data likelihood of $\fobs$ becomes
+\begin{eqnarray}\displaystyle
+p(\fobs |\sigma^2_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
+  N(\ftrue | \mu_i,  \sigma^2_i + s_{\mu}^2)\,  I(\ftrue \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
+ \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma_i^2 + s_{\mu}^2)\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
+\end{array}\right.\label{eq:mlik_s}
+\end{eqnarray}
+where the boolean indictor function, $I(\ftrue \ge b_i)$, is 1 if $\ftrue \ge b_i$ and 0 otherwise.
+
+Next, we model the reported variance, $\sobs$, on the uncertainty of the magnitude measurement.
+Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true uncertainty variance, $\sigma^2_i$ through a Gamma likelihood, which is defined to take on non-negative values.
+Our likelihood of observed $\sobs$, given $\sigma^2_i$, $\theta$ and $I$, is
+\begin{eqnarray}\displaystyle
+p(\sobs | \fobs, \sigma^2_i, \theta,I) &=& \left\{\begin{array}{ll}
+  \Gamma (\sobs | \sigma^2_i, V_{\sigma} ) & \mbox{if $\fobs \ne \texttt{NA}$} \\
+ 1& \mbox{if $\fobs = \texttt{NA}$}
+\end{array}\right.\label{eq:slik}
+\end{eqnarray}
+where $\Gamma(x | m, V)$ is the standard Gamma distribution with mean $m$ and variance $V$, 
+and we have added a model parameter $V_\sigma$ to represent the
+variance in the distribution of reported uncertainties given the true
+uncertainty.  (Typically, the Gamma distribution is parameterized by parameters $(\alpha, \beta)$, where the mean is $\alpha\beta$ and the variance is $\alpha\beta^2$.)  The  purpose of 
+ including the likelihood in Equation \ref{eq:slik} to the model, in practice, to keep the $\sigma^2_i$ from
+drifting very far away from the $s^2_i$, as set by the hyperparameter
+$V_\sigma$.
+
+Putting it all together, we can write down the likelihood of the data, $D_i$, for the parameters $\theta$, on a single epoch, as
+\begin{eqnarray}\displaystyle
+p(D_i|\theta,I) &=& \int_0^{\infty} p(\fobs |\sigma^2_i,\theta,I)\, p(\sobs | \fobs, \sigma^2_i,\theta,I)\, p(\sigma^2_i | \theta, I)\,\dd \sigma^2_i
 \\
-\theta &\equiv& (\omega, \{A_k\}, s_\mu^2, v_\sigma, B, V_B, S, V_S)
-\quad,
+p(\sigma^2_i|\theta, I) &\propto& \sigma_i^{-1}
+\\
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s_\mu^2, B, V_B, V_{\sigma}) \quad ,
 \end{eqnarray}
-where we have (at last) explicitly assembled all the variable-star
-parameters and hyperparameters into the big parameter vector $\theta$.
-The huge sets of nuisance parameters $\{b_i\}$ and
-$\{\sigma_i\}$ don't remain, because by integration we
-removed them.
+where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), and integrated out the nuisance parameter, $\sigma^2_i$.
+We have assumed a Jeffrey's prior on $\sigma^2_i$, which is non-informative and invariant to reparametrization of the variance.
+We could alternatively use an inverse-gamma prior, which takes two hyperparameters (and is conjugate in the case of a normal likelihood with unknown variance).
+% We have also assumed that $\fobs$ and $\sobs$ are conditionally independent given $\sigma^2_i$, $\theta$, and $I$.
+This is a reasonable assumption in the case that a detection is made at epoch $t_i$.
 
-Finally and it goes without saying that if we treat the data points as
-independent (as we are free to do, given the non-awesomeness of the
-software data source), we have that
+Finally if we assume that the data collected at each epoch are independent given the model parameters, we have that
 \begin{eqnarray}\displaystyle
@@
 \label{sec:experiments}
+
+\joey{Note: ASAS provides periods and amplitude for all of the objects in ACVS!  We can and should use this as a comparison set}
+
+\joey{From a brief reading of a few Mira papers, it seems that some papers throw away Miras for which no trough of the LC was observed.
+This seems silly.
+I have an idea of the P-A plots from Hipparcos.
+Should be a good comparison set.
+Also, we can consider applying this to Hipparcos, which has publicly available data.}
 

### 2012-04-10 tex/ms.tex
small typo corrected
add two hyperparameters, small Latex tweaks, drop and ignore pdf file

@@
 \begin{eqnarray}\displaystyle
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s_\mu^2, \cdots) \quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, \cdots) \quad ,
 \end{eqnarray}
@@
 
-We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma^2_i$) and model ($s_{\mu}$) contributions.
+We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma^2_i$) and model ($s^2_{\mu}$) contributions.
 In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = \ftrue \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($\ftrue < b_i$).
@@
 p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
-  N(\ftrue | \mu_i,  \sigma^2_i + s_{\mu}^2)\,  \int_0^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
-  \int_{0}^{\infty} \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s_{\mu}^2)\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
+  N(\ftrue | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_0^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
+  \int_{0}^{\infty} \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik}
@@
 \\
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s_\mu^2, B, V_B, \cdots) \quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, \cdots) \quad ,
 \end{eqnarray}
@@
 p(\fobs |\sigma^2_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
-  N(\ftrue | \mu_i,  \sigma^2_i + s_{\mu}^2)\,  I(\ftrue \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
- \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma_i^2 + s_{\mu}^2)\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
+  N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
+ \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma_i^2 + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik_s}
+\quad ,
 \end{eqnarray}
@@
 \end{array}\right.\label{eq:slik}
+\quad ,
 \end{eqnarray}
@@
 \\
-p(\sigma^2_i|\theta, I) &\propto& \sigma_i^{-1}
+p(\sigma^2_i|\theta, I) &=& \Gamma(\sigma^2_i | S, V_S)
 \\
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s_\mu^2, B, V_B, V_{\sigma}) \quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, V_{\sigma}, S, V_S) \quad ,
 \end{eqnarray}
-where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), and integrated out the nuisance parameter, $\sigma^2_i$.
-We have assumed a Jeffrey's prior on $\sigma^2_i$, which is non-informative and invariant to reparametrization of the variance.
-We could alternatively use an inverse-gamma prior, which takes two hyperparameters (and is conjugate in the case of a normal likelihood with unknown variance).
-% We have also assumed that $\fobs$ and $\sobs$ are conditionally independent given $\sigma^2_i$, $\theta$, and $I$.
-This is a reasonable assumption in the case that a detection is made at epoch $t_i$.
+where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), we have integrated out the nuisance parameter $\sigma^2_i$, again the $\Gamma$ is the Gamma function with specifid mean and variance, and we have two more hyperparameters.
 
@@
 \begin{enumerate}
-\item Convert the observed magnitudes, $m_i$, and errors, $s_{m,i}$ to fluxes, $f_i$ and flux errors, $s_{f,i}$ (we assume a V-band zero-point of $3.67\times10^{-9}$ erg/s/cm$^2$/\AA)
-\item For a given flux dimming parameter, $d$, sample the new fluxes, $\tilde{f}_i$, from a Gaussian distribution centered around $f_i / d$ with standard deviation sampled from the empirical ASAS distribution of $s_{f,i} | f_i$.
+\item Convert the observed magnitudes $m_i$ and uncertainty variances $s^2_{m,i}$ to fluxes $f_i$ and flux uncertainty variances $s^2_{f,i}$ (we assume a V-band zero-point of $3.67\times10^{-9}$ erg/s/cm$^2$/\AA)
+\item For a given flux dimming parameter, $d$, sample the new fluxes, $\tilde{f}_i$, from a Gaussian distribution centered around $f_i / d$ with standard deviation sampled from the empirical ASAS distribution of $s^2_{f,i} | f_i$.
 \joey{I found that relationship to be linear with a slope $\approx 1$, which is probably BS since the ASAS mag errors are all crap.

### 2012-04-10 tex/ms.tex
fix typos

@@
 In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = \ftrue \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($\ftrue < b_i$).
-To derive the likelihood of the observed $\fobs$, given $\sigma^2_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
+To derive the likelihood of the observed $\fobs$ given $\sigma^2_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
 \begin{eqnarray}\displaystyle
@@
   N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
- \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma_i^2 + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
+ \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik_s}
@@
 Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true uncertainty variance, $\sigma^2_i$ through a Gamma likelihood, which is defined to take on non-negative values.
-Our likelihood of observed $\sobs$, given $\sigma^2_i$, $\theta$ and $I$, is
+Our likelihood of observed $\sobs$ given $\fobs$, $\sigma^2_i$, $\theta$ and $I$, is
 \begin{eqnarray}\displaystyle
@@
  including the likelihood in Equation \ref{eq:slik} to the model, in practice, to keep the $\sigma^2_i$ from
-drifting very far away from the $s^2_i$, as set by the hyperparameter
+drifting very far away from the $\sobs$, as set by the hyperparameter
 $V_\sigma$.
@@
 \end{eqnarray}
-where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), we have integrated out the nuisance parameter $\sigma^2_i$, again the $\Gamma$ is the Gamma function with specifid mean and variance, and we have two more hyperparameters.
+where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), we have integrated out the nuisance parameter $\sigma^2_i$, again the $\Gamma$ is the Gamma function with specified mean and variance, and we have two more hyperparameters $S$ and $V_S$.
 

### 2012-04-11 py/censoring2.py
version controlling new censoring code

@@ -1,0 +1,41 @@
+###
+### implements censoring model
+### date April 10, 2012
+###
+
+import numpy as np
+from scipy.stats import norm
+from scipy.stats import gamma
+
+def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
+    """ 
+    computes log(p(D|theta,I))
+    t: times when not censored (1-d ndarray of length n)
+    f: flux when not censored (1-d ndarray of length n)
+    e: reported error when not censored (1-d ndarray of length n)
+    tc: times of censored observations (1-d ndarray of length m)
+    everything else: exactly as in paper (all floats)
+    """
+    ## convert all times to expected flux (i.e. t_i -> u_i)
+    u = A0 + A1*np.sin(t*w*2*np.pi) + B1*np.cos(t*w*2*np.pi)
+    uc = A0 + A1*np.sin(tc*w*2*np.pi) + B1*np.cos(tc*w*2*np.pi)
+
+    ## compute loglikelihood
+    ll = 0
+    for i in uc:
+        ll += likelihood_censored(i,su2,B,VB)
+    for i in u:
+        ll += likelihood_observed()
+    return ll
+
+def likelihood_censored(uc,su2,B,VB):
+    ## integrand of equation (7), integrate this across bi
+    def erf_integral(bi,sig2,uc,su2,B,VB):
+        return norm.cdf((bi - uc) / np.sqrt(sig2 + su2)) * (1 / (np.sqrt(2*np.pi*VB))) * np.exp(-(bi-B)**2/(2*VB))
+    ## integrand of equation (12), integrate this across sig2
+    def sig_integral(sig2,uc,su2,B,VB):
+        return integrate.quad(erf_integral, 0.5, 1.5, [sig2,uc,su2,B,VB]) * (1/VB)*gamma.pdf(sig2,B/VB)
+    return integrate.quad(sig_integral,.5,1.5,[uc,su2,B,VB])
+    
+def liklihood_observed(c):
+    print "hello"

### 2012-04-11 py/censoring.py
old code, don't use, replaced by censoring2.py

@@ -1,3 +1,19 @@
+#################
+################# WARNING: OLD, DON'T USE
+#################
+
+
+
+
+
+
+
+
+
+
+
+
+
 import numpy as np
 from scipy.optimize import fmin
 from scipy import special

### 2012-04-11 py/censoring2.py
new code for the censoring model

@@ -9,12 +9,12 @@ from scipy.stats import gamma
 
 def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     """ 
-    computes log(p(D|theta,I))
-    t: times when not censored (1-d ndarray of length n)
-    f: flux when not censored (1-d ndarray of length n)
-    e: reported error when not censored (1-d ndarray of length n)
-    tc: times of censored observations (1-d ndarray of length m)
-    everything else: exactly as in paper (all floats)
+    computes log p(D|theta,I)
+    t: times when not censored (1-d ndarray)
+    f: flux when not censored (1-d ndarray)
+    e: reported error when not censored (1-d ndarray)
+    tc: times of censored observations (1-d ndarray)
+    everything else: same notation as paper (all floats)
     """
     ## convert all times to expected flux (i.e. t_i -> u_i)
     u = A0 + A1*np.sin(t*w*2*np.pi) + B1*np.cos(t*w*2*np.pi)
@@ -24,8 +24,8 @@ def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     ll = 0
     for i in uc:
         ll += likelihood_censored(i,su2,B,VB)
-    for i in u:
-        ll += likelihood_observed()
+    for i in np.range(u.size):
+        ll += likelihood_observed(u[i],f[i],e[i],su2,B,VB,Vsig,S,VS)
     return ll
 
 def likelihood_censored(uc,su2,B,VB):
@@ -34,8 +34,19 @@ def likelihood_censored(uc,su2,B,VB):
         return norm.cdf((bi - uc) / np.sqrt(sig2 + su2)) * (1 / (np.sqrt(2*np.pi*VB))) * np.exp(-(bi-B)**2/(2*VB))
     ## integrand of equation (12), integrate this across sig2
     def sig_integral(sig2,uc,su2,B,VB):
-        return integrate.quad(erf_integral, 0.5, 1.5, [sig2,uc,su2,B,VB]) * (1/VB)*gamma.pdf(sig2,B/VB)
-    return integrate.quad(sig_integral,.5,1.5,[uc,su2,B,VB])
[…]

### 2012-04-11 py/censoring2.py
fixed a bug, forgot to take logs

@@ -23,9 +23,9 @@ def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     ## compute loglikelihood
     ll = 0
     for i in uc:
-        ll += likelihood_censored(i,su2,B,VB)
+        ll += np.log(likelihood_censored(i,su2,B,VB))
     for i in np.range(u.size):
-        ll += likelihood_observed(u[i],f[i],e[i],su2,B,VB,Vsig,S,VS)
+        ll += np.log(likelihood_observed(u[i],f[i],e[i],su2,B,VB,Vsig,S,VS))
     return ll
 
 def likelihood_censored(uc,su2,B,VB):

### 2012-04-11 py/censoring2.py
minor bug fixes in censoring2.py.  it now runs for me in Python 2.7.2 | EPD 7.2-2

@@ -6,6 +6,7 @@
 import numpy as np
 from scipy.stats import norm
 from scipy.stats import gamma
+from scipy import integrate
 
 def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     """ 
@@ -24,7 +25,7 @@ def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     ll = 0
     for i in uc:
         ll += np.log(likelihood_censored(i,su2,B,VB))
-    for i in np.range(u.size):
+    for i in np.arange(u.size):
         ll += np.log(likelihood_observed(u[i],f[i],e[i],su2,B,VB,Vsig,S,VS))
     return ll
 
@@ -34,17 +35,17 @@ def likelihood_censored(uc,su2,B,VB):
         return norm.cdf((bi - uc) / np.sqrt(sig2 + su2)) * (1 / (np.sqrt(2*np.pi*VB))) * np.exp(-(bi-B)**2/(2*VB))
     ## integrand of equation (12), integrate this across sig2
     def sig_integral(sig2,uc,su2,B,VB):
-        return integrate.quad(erf_integral,B-3*np.sqrt(VB),B+3*np.sqrt(VB), [sig2,uc,su2,B,VB]) * (1/VB)*gamma.pdf(sig2,B/VB)
-    return integrate.quad(sig_integral,0.0,10,[uc,su2,B,VB])
+        return integrate.quad(erf_integral,B-3*np.sqrt(VB),B+3*np.sqrt(VB), (sig2,uc,su2,B,VB))[0] * (1/VB)*gamma.pdf(sig2,B/VB)
+    return integrate.quad(sig_integral,0.0,10,(uc,su2,B,VB))[0]
     
-def liklihood_observed(u,f,e,su2,B,VB,Vsig,S,VS):
+def likelihood_observed(u,f,e,su2,B,VB,Vsig,S,VS):
     def integrand(sig2,u,f,e,su2,B,VB,Vsig,S,VS):
         p_not_cens = norm.cdf((f - B) / np.sqrt(VB))
         p_flux = norm.pdf((f - u) / np.sqrt(sig2 + su2))
         p_si = (1/Vsig)*gamma.pdf(e,sig2/Vsig)
         p_sig2 = (1/VB)*gamma.pdf(sig2,B/VB)
[…]

### 2012-04-11 py/censoring2.py
changed gamma calls

@@ -5,7 +5,7 @@
 
 import numpy as np
 from scipy.stats import norm
-from scipy.stats import gamma
+from scipy.special import gamma
 from scipy import integrate
 
 def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
@@ -35,7 +35,8 @@ def likelihood_censored(uc,su2,B,VB):
         return norm.cdf((bi - uc) / np.sqrt(sig2 + su2)) * (1 / (np.sqrt(2*np.pi*VB))) * np.exp(-(bi-B)**2/(2*VB))
     ## integrand of equation (12), integrate this across sig2
     def sig_integral(sig2,uc,su2,B,VB):
-        return integrate.quad(erf_integral,B-3*np.sqrt(VB),B+3*np.sqrt(VB), (sig2,uc,su2,B,VB))[0] * (1/VB)*gamma.pdf(sig2,B/VB)
+        p_sig2 = (1/((VB**B)*gamma(V)))*(sig2**(B-1))*np.exp(-sig2/VB)
+        return integrate.quad(erf_integral,B-3*np.sqrt(VB),B+3*np.sqrt(VB), (sig2,uc,su2,B,VB))[0] * p_sig2
     return integrate.quad(sig_integral,0.0,10,(uc,su2,B,VB))[0]
     
 def likelihood_observed(u,f,e,su2,B,VB,Vsig,S,VS):
@@ -43,7 +44,7 @@ def likelihood_observed(u,f,e,su2,B,VB,Vsig,S,VS):
         p_not_cens = norm.cdf((f - B) / np.sqrt(VB))
         p_flux = norm.pdf((f - u) / np.sqrt(sig2 + su2))
         p_si = (1/Vsig)*gamma.pdf(e,sig2/Vsig)
-        p_sig2 = (1/VB)*gamma.pdf(sig2,B/VB)
+        p_sig2 = (1/((VB**B)*gamma(V)))*(sig2**(B-1))*np.exp(-sig2/VB)
         return p_not_cens * p_flux * p_si * p_sig2
     return integrate.quad(integrand,0.0,10,(u,f,e,su2,B,VB,Vsig,S,VS))[0]
 

### 2012-04-11 py/censoring2.py
changes to censoring2.py made during coffee hour on 04112012

@@ -4,9 +4,15 @@
 ###
 
 import numpy as np
-from scipy.stats import norm
-from scipy.stats import gamma
+import scipy.stats as stats
+#from scipy.stats import norm as scipynorm
+#from scipy.stats import gamma as scipygamma
 from scipy import integrate
+
+
+
+def newgamma(mean,var):
+    return stats.gamma(mean**2/var, scale = var/mean)
 
 def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     """ 
@@ -22,32 +28,29 @@ def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
     uc = A0 + A1*np.sin(tc*w*2*np.pi) + B1*np.cos(tc*w*2*np.pi)
 
     ## compute loglikelihood
-    ll = 0
-    for i in uc:
-        ll += np.log(likelihood_censored(i,su2,B,VB))
-    for i in np.arange(u.size):
-        ll += np.log(likelihood_observed(u[i],f[i],e[i],su2,B,VB,Vsig,S,VS))
-    return ll
-
-def likelihood_censored(uc,su2,B,VB):
+    return np.sum(loglikelihood_censored(uc,su2,B,VB,S,VS)) + \
+         np.sum(loglikelihood_observed(u,f,e,su2,B,VB,Vsig,S,VS))
+    
+def loglikelihood_censored(uc,su2,B,VB,S,VS):
     ## integrand of equation (7), integrate this across bi
     def erf_integral(bi,sig2,uc,su2,B,VB):
-        return norm.cdf((bi - uc) / np.sqrt(sig2 + su2)) * (1 / (np.sqrt(2*np.pi*VB))) * np.exp(-(bi-B)**2/(2*VB))
+        return stats.norm.cdf(uc, bi, np.sqrt(sig2 + su2)) \
+               * stats.gamma.pdf(bi,B**2/VB, scale = VB/B)
     ## integrand of equation (12), integrate this across sig2
-    def sig_integral(sig2,uc,su2,B,VB):
-        return integrate.quad(erf_integral,B-3*np.sqrt(VB),B+3*np.sqrt(VB), (sig2,uc,su2,B,VB))[0] * (1/VB)*gamma.pdf(sig2,B/VB)
[…]

### 2012-04-11 py/censoring2.py
consistent norm

@@ -35,7 +35,7 @@ def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
 def loglikelihood_censored(uc,su2,B,VB,S,VS):
     ## integrand of equation (7), integrate this across bi
     def erf_integral(bi,sig2,uc,su2,B,VB):
-        return stats.norm.cdf(uc, bi, np.sqrt(sig2 + su2)) \
+        return stats.norm.cdf((uc - bi) / np.sqrt(sig2 + su2)) \
                * stats.gamma.pdf(bi,B**2/VB, scale = VB/B)
     ## integrand of equation (12), integrate this across sig2
     def sig_integrand(sig2,uc,su2,B,VB,S,VS):

### 2012-04-11 py/fake_data.py
generate data according to the model

@@ -1,0 +1,71 @@
+'''
+This file is part of the Censored Data project.
+Copyright 2012 the authors.
+
+issues:
+-------
+- not written
+- doesn't run
+- not tested
+'''
+
+if __name__ == '__main__':
+    import matplotlib
+    matplotlib.use('Agg')
+    from matplotlib import rc
+    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':12})
+    rc('text', usetex=True)
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.stats import norm as Norm
+from scipy.stats import gamma as Gamma
+
+def mu(times, omega, A0, A1, B1):
+    return A0 \
+        + A1 * np.cos(2. * np.pi * omega * times) \
+        + A1 * np.sin(2. * np.pi * omega * times)
+
+def make_fake_data(omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS, size=1):
+    times = 3. * np.random.uniform(size=size)
+    mus = mu(times, omega, A0, A1, B1)
+    sigma2s = Gamma.rvs(S**2/VS, scale=VS/S, size=times.size)
+    print sigma2s
+    assert(np.all(sigma2s >= 0.))
+    sobs2s = np.array([Gamma.rvs(m**2/Vsigma, scale=Vsigma/m, size=1)[0] for m in sigma2s])
+    print sobs2s
+    assert(np.all(sobs2s >= 0.))
+    fobss = mus + np.sqrt(sigma2s + s2mu) * Norm.rvs(size=times.size)
+    bs = Gamma.rvs(B**2/VB, scale=VB/B, size=times.size)
+    assert(np.all(bs >= 0.))
+    good = fobss > bs
+    bad = good == False
+    print np.sum(good), times[good]
+    print np.sum(bad), times[bad]
+    return times[good], fobss[good], sobs2s[good], times[bad]
+
+def plot_data(ax, goodtimes, fobss, sobs2s, badtimes):
+    ax.errorbar(goodtimes, fobss, fmt='ko', yerr=np.sqrt(sobs2s))
+    ax.plot(badtimes, np.zeros_like(badtimes), 'ro')
+    return None
+
+def plot_model(ax, omega, A0, A1, B1):
+    tlim = ax.get_xlim()
[…]

### 2012-04-11 py/fake_data.py
comments

@@ -1,11 +1,9 @@
 '''
 This file is part of the Censored Data project.
-Copyright 2012 the authors.
+Copyright 2012 David W. Hogg (NYU).
 
 issues:
 -------
-- not written
-- doesn't run
 - not tested
 '''
 
@@ -21,18 +19,23 @@ from scipy.stats import norm as Norm
 from scipy.stats import gamma as Gamma
 
 def mu(times, omega, A0, A1, B1):
+    '''
+    Make the model mean flux given lightcurve parameters.
+    '''
     return A0 \
         + A1 * np.cos(2. * np.pi * omega * times) \
         + A1 * np.sin(2. * np.pi * omega * times)
 
 def make_fake_data(omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS, size=1):
+    '''
+    Implement the full generative model from the paper.  Is this
+    correct?
+    '''
     times = 3. * np.random.uniform(size=size)
     mus = mu(times, omega, A0, A1, B1)
     sigma2s = Gamma.rvs(S**2/VS, scale=VS/S, size=times.size)
-    print sigma2s
     assert(np.all(sigma2s >= 0.))
     sobs2s = np.array([Gamma.rvs(m**2/Vsigma, scale=Vsigma/m, size=1)[0] for m in sigma2s])
-    print sobs2s
     assert(np.all(sobs2s >= 0.))
     fobss = mus + np.sqrt(sigma2s + s2mu) * Norm.rvs(size=times.size)
     bs = Gamma.rvs(B**2/VB, scale=VB/B, size=times.size)
@@ -59,10 +62,9 @@ if __name__ == '__main__':
     omega, A0, A1, B1 = 2.5, 5.0, 3.0, 0.5
     s2mu = 0.2**2
     S, VS = 0.2, 0.04**2
-    B, VB = 4.0, 0.4**2
+    B, VB = 3.5, 1.0**2
     Vsigma = 0.1**2
     gt, f, s2, bt = make_fake_data(omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS, size=64)
-    print gt.size, f.size, s2.size, bt.size
     ax = plt.subplot(111)
     plot_data(ax, gt, f, s2, bt)
     plot_model(ax, omega, A0, A1, B1)

### 2012-04-11 py/fake_data.py
small changes, bug fixes

@@ -5,6 +5,7 @@ Copyright 2012 David W. Hogg (NYU).
 issues:
 -------
 - not tested
+- not properly commented
 '''
 
 if __name__ == '__main__':
@@ -23,27 +24,21 @@ def mu(times, omega, A0, A1, B1):
     Make the model mean flux given lightcurve parameters.
     '''
     return A0 \
-        + A1 * np.cos(2. * np.pi * omega * times) \
-        + A1 * np.sin(2. * np.pi * omega * times)
+        + A1 * np.cos(omega * times) \
+        + B1 * np.sin(omega * times)
 
-def make_fake_data(omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS, size=1):
+def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     '''
     Implement the full generative model from the paper.  Is this
     correct?
     '''
-    times = 3. * np.random.uniform(size=size)
     mus = mu(times, omega, A0, A1, B1)
     sigma2s = Gamma.rvs(S**2/VS, scale=VS/S, size=times.size)
-    assert(np.all(sigma2s >= 0.))
     sobs2s = np.array([Gamma.rvs(m**2/Vsigma, scale=Vsigma/m, size=1)[0] for m in sigma2s])
-    assert(np.all(sobs2s >= 0.))
     fobss = mus + np.sqrt(sigma2s + s2mu) * Norm.rvs(size=times.size)
     bs = Gamma.rvs(B**2/VB, scale=VB/B, size=times.size)
-    assert(np.all(bs >= 0.))
     good = fobss > bs
     bad = good == False
-    print np.sum(good), times[good]
-    print np.sum(bad), times[bad]
     return times[good], fobss[good], sobs2s[good], times[bad]
 
 def plot_data(ax, goodtimes, fobss, sobs2s, badtimes):
@@ -59,12 +54,12 @@ def plot_model(ax, omega, A0, A1, B1):
     return None
 
 if __name__ == '__main__':
-    omega, A0, A1, B1 = 2.5, 5.0, 3.0, 0.5
+    omega, A0, A1, B1 = 2. * np.pi / 400., 5.0, 3.0, 0.5
     s2mu = 0.2**2
-    S, VS = 0.2, 0.04**2
-    B, VB = 3.5, 1.0**2
-    Vsigma = 0.1**2
[…]

### 2012-04-11 py/censoring2.py
fixed censoring2.py code

@@ -1,6 +1,8 @@
 ###
 ### implements censoring model
 ### date April 10, 2012
+
+### Updated by JWR on 20120411
 ###
 
 import numpy as np
@@ -11,46 +13,57 @@ from scipy.stats import norm
 from scipy import integrate
 
 
+class Censored:
 
-def newgamma(mean,var):
-    return stats.gamma(mean**2/var, scale = var/mean)
-
-def log_likelihood(t,f,e,tc,w,A0,A1,B1,su2,B,VB,Vsig,S,VS):
-    """ 
-    computes log p(D|theta,I)
-    t: times when not censored (1-d ndarray)
-    f: flux when not censored (1-d ndarray)
-    e: reported error when not censored (1-d ndarray)
-    tc: times of censored observations (1-d ndarray)
-    everything else: same notation as paper (all floats)
-    """
+    def __init__(self, t, f, e, tc, tolquad = 0.1):
+        """ 
+        t: times when not censored (1-d ndarray)
+        f: flux when not censored (1-d ndarray)
+        e: reported error when not censored (1-d ndarray)
+        tc: times of censored observations (1-d ndarray)
+        """      
+        self.tolquad = tolquad # numerical tolerance in integration
+        # insert data 
+        self.t = t
+        self.f = f
+        self.e = e
+        self.tc = tc
+        
+    def log_likelihood(self,P,A0,A1,B1,su2,B,VB,Vsig,S,VS):
+        """ 
+        computes log p(D|theta,I)
+        everything else: same notation as paper (all floats)
+        """
     ## convert all times to expected flux (i.e. t_i -> u_i)
-    u = A0 + A1*np.sin(t*w*2*np.pi) + B1*np.cos(t*w*2*np.pi)
-    uc = A0 + A1*np.sin(tc*w*2*np.pi) + B1*np.cos(tc*w*2*np.pi)
+        w =  (2 * np.pi)/ P 
+        self.u = A0 + A1*np.cos(self.t*w) + B1*np.sin(self.t*w)
+        self.uc = A0 + A1*np.cos(self.tc*w) + B1*np.sin(self.tc*w)
 
     ## compute loglikelihood
[…]

### 2012-04-11 tex/ms.tex
adding graphical model
commented graphical model b/c wouldn't compile
changed prior on b_i to normal

@@
 \usepackage{longtable}
+\usepackage{amsmath}
 
@@
 p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
-  N(\ftrue | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_0^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
-  \int_{0}^{\infty} \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
+  N(\ftrue | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_{-\infty}^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
+  \int_{-\infty}^{\infty} \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik}
@@
 In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $\ftrue$, enforcing that the detection limit be fainter than the observed brightness.
-Likewise, in the epochs for which no detection was made ($\fobs = \texttt{NA}$), we integrate the joint $(\ftrue, b_i)$ likelihood over all possible values of $b_i$ and over the unknown $\ftrue$ from $0$ to $b_i$, ensuring that the brightness (if it were able to be observed) be fainter than the detection limit.
+In the epochs for which no detection was made ($\fobs = \texttt{NA}$), we compute the probability that the unknown detection threshold $b_i$ is greater than the unknown true flux $f_i$.
+When both $b_i$ and $f_i$ are normally distributed, this can be computed using a single call to a Gaussian cumulative distribution function, specifically
+\begin{equation*}
+\mathbb{P}(b_i \geq f_i) = \int_{-\infty}^0 N(x|B-\mu_i,V_B + \sigma_i^2 + s_{\mu}^2) dx
+\end{equation*}
 
@@
   N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
- \int_{0}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
+ \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik_s}
@@
 veracity.
+
+\begin{figure}[hbtp]
+  \begin{center}
+    \includegraphics[width=\textwidth,%
+      trim=3cm 12cm 5cm 3cm, clip=true]{gm/gm.pdf}
+  \end{center}
+  \caption{The graphical representation of the probability distribution.
+    \label{fig:graph}}
+\end{figure}
 

### 2012-04-11 py/censoring2.py
removed inside integral for censored observations. turned into call to normal cdf

@@ -44,14 +44,10 @@ class Censored:
                np.sum(self.loglikelihood_observed(su2,B,VB,Vsig,S,VS))
     
     def loglikelihood_censored(self,su2,B,VB,S,VS):
-    ## integrand of equation (7), integrate this across bi
-        def erf_integral(bi,ui,sig2,su2,B,VB):
-            return stats.norm.cdf((ui - bi) / np.sqrt(sig2 + su2)) \
-                   * stats.gamma.pdf(bi,B**2/VB, scale = VB/B)
     ## integrand of equation (12), integrate this across sig2
         def sig_integrand(sig2,ui,su2,B,VB,S,VS):
-            return integrate.quad(erf_integral,0.,B+5.*np.sqrt(VB), (ui,sig2,su2,B,VB),epsabs = self.tolquad)[0] \
-                   * stats.gamma.pdf(sig2,S**2/VS, scale = VS/S)
+            return  stats.norm.cdf((B-ui) / np.sqrt(VB + sig2 + su2)) * \
+                stats.gamma.pdf(sig2,S**2/VS, scale = VS/S)
         return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0] \
                        for ui in self.uc])
 

### 2012-04-12 py/censoring2.py
rewrote parameters as par (single input to llik function)

@@ -29,11 +29,22 @@ class Censored:
         self.e = e
         self.tc = tc
         
-    def log_likelihood(self,P,A0,A1,B1,su2,B,VB,Vsig,S,VS):
+    def log_likelihood(self,par):
         """ 
         computes log p(D|theta,I)
         everything else: same notation as paper (all floats)
         """
+        P = par[0]
+        A0 = par[1]
+        A1 = par[2]
+        B1 = par[3]
+        su2 = par[4]
+        B = par[5]
+        VB = par[6]
+        Vsig = par[7]
+        S = par[8]
+        VS = par[9]
+        
     ## convert all times to expected flux (i.e. t_i -> u_i)
         w =  (2 * np.pi)/ P 
         self.u = A0 + A1*np.cos(self.t*w) + B1*np.sin(self.t*w)

### 2012-04-12 py/censoring2.py
changed scipy pdf calls to our own functions

@@ -11,6 +11,7 @@ import scipy.stats as stats
 #from scipy.stats import gamma as scipygamma
 from scipy.stats import norm
 from scipy import integrate
+from scipy.special import gamma
 
 
 class Censored:
@@ -58,19 +59,26 @@ class Censored:
     ## integrand of equation (12), integrate this across sig2
         def sig_integrand(sig2,ui,su2,B,VB,S,VS):
             return  stats.norm.cdf((B-ui) / np.sqrt(VB + sig2 + su2)) * \
-                stats.gamma.pdf(sig2,S**2/VS, scale = VS/S)
+                gamma_pdf(sig2,S,VS)
         return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0] \
                        for ui in self.uc])
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
         def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
             p_not_cens = stats.gamma.cdf(fi,B**2/VB, scale = VB/B)
-            p_flux = stats.norm.pdf((fi - ui) / np.sqrt(sig2 + su2))
-            p_si = stats.gamma.pdf(ei, sig2**2/Vsig, scale = Vsig/sig2)
-            p_sig2 = stats.gamma.pdf(sig2,S**2/VS, scale = VS/S)
+            p_flux = gaussian_pdf((fi - ui) / np.sqrt(sig2 + su2))
+            p_si = gamma_pdf(ei,sig2,Vsig)
+            p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
         return np.log([integrate.quad(integrand,0.,S+5.*np.sqrt(VS),(ui,fi,ei,su2,B,VB,Vsig,S,VS),epsabs=self.tolquad)[0]\
                        for (ui,fi,ei) in zip(self.u,self.f,self.e)])
+
+def gaussian_pdf(x):
+    return (1 / (np.sqrt(2.*np.pi))) * np.exp(-(x**2) / (2.))    
+
+def gamma_pdf(x,k,theta):
+    return (1 / ((theta**k)*gamma(k))) *  (x**(k-1.)) * np.exp(-x / theta)
+
 
 ### TODO: 1. intelligently choose limits for gamma integral
 ###       2. check that my gamma interpretation is right

### 2012-04-12 py/censoring2.py
corrected gamma call

@@ -76,7 +76,9 @@ class Censored:
 def gaussian_pdf(x):
     return (1 / (np.sqrt(2.*np.pi))) * np.exp(-(x**2) / (2.))    
 
-def gamma_pdf(x,k,theta):
+def gamma_pdf(x,mean,var):
+    theta = var / mean
+    k = mean / theta
     return (1 / ((theta**k)*gamma(k))) *  (x**(k-1.)) * np.exp(-x / theta)
 
 

### 2012-04-12 py/censoring2.py
using our own gaussian_cdf, not scipy

@@ -12,7 +12,7 @@ import scipy.stats as stats
 from scipy.stats import norm
 from scipy import integrate
 from scipy.special import gamma
-
+from scipy.special import erf
 
 class Censored:
 
@@ -58,7 +58,7 @@ class Censored:
     def loglikelihood_censored(self,su2,B,VB,S,VS):
     ## integrand of equation (12), integrate this across sig2
         def sig_integrand(sig2,ui,su2,B,VB,S,VS):
-            return  stats.norm.cdf((B-ui) / np.sqrt(VB + sig2 + su2)) * \
+            return  gaussian_cdf((B-ui) / np.sqrt(VB + sig2 + su2)) * \
                 gamma_pdf(sig2,S,VS)
         return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0] \
                        for ui in self.uc])
@@ -75,6 +75,9 @@ class Censored:
 
 def gaussian_pdf(x):
     return (1 / (np.sqrt(2.*np.pi))) * np.exp(-(x**2) / (2.))    
+
+def gaussian_cdf(x):
+    return .5*(1. + erf(x/np.sqrt(2.)))
 
 def gamma_pdf(x,mean,var):
     theta = var / mean

### 2012-04-12 py/censoring2.py
gamma_cdf is now called with our own function, not scipy

@@ -13,6 +13,7 @@ from scipy.stats import norm
 from scipy import integrate
 from scipy.special import gamma
 from scipy.special import erf
+from scipy.special import gammainc
 
 class Censored:
 
@@ -65,7 +66,7 @@ class Censored:
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
         def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
-            p_not_cens = stats.gamma.cdf(fi,B**2/VB, scale = VB/B)
+            p_not_cens = gamma_cdf(fi,B**2/VB, scale = VB/B)
             p_flux = gaussian_pdf((fi - ui) / np.sqrt(sig2 + su2))
             p_si = gamma_pdf(ei,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
@@ -83,6 +84,12 @@ def gamma_pdf(x,mean,var):
     theta = var / mean
     k = mean / theta
     return (1 / ((theta**k)*gamma(k))) *  (x**(k-1.)) * np.exp(-x / theta)
+
+def gamma_cdf(x,mean,var):
+    theta = var / mean
+    k = mean / theta
+    return  gammainc(k,x/theta) ## wikipedia and scipy define imcomplete gamma differently
+
 
 
 ### TODO: 1. intelligently choose limits for gamma integral

### 2012-04-12 py/censoring2.py
commented scipy.stats imports

@@ -6,10 +6,10 @@
 ###
 
 import numpy as np
-import scipy.stats as stats
+#import scipy.stats as stats
 #from scipy.stats import norm as scipynorm
 #from scipy.stats import gamma as scipygamma
-from scipy.stats import norm
+#from scipy.stats import norm
 from scipy import integrate
 from scipy.special import gamma
 from scipy.special import erf

### 2012-04-12 py/censoring2.py
added fmin method to censoring2 file and committed test_censoring2 file, which profiles code & runs fmin on fake data

@@ -14,6 +14,7 @@ from scipy import integrate
 from scipy.special import gamma
 from scipy.special import erf
 from scipy.special import gammainc
+import scipy.optimize as op
 
 class Censored:
 
@@ -66,13 +67,22 @@ class Censored:
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
         def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
-            p_not_cens = gamma_cdf(fi,B**2/VB, scale = VB/B)
+            p_not_cens = gamma_cdf(fi,B,VB)#fi,B**2/VB, scale = VB/B)
             p_flux = gaussian_pdf((fi - ui) / np.sqrt(sig2 + su2))
             p_si = gamma_pdf(ei,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
         return np.log([integrate.quad(integrand,0.,S+5.*np.sqrt(VS),(ui,fi,ei,su2,B,VB,Vsig,S,VS),epsabs=self.tolquad)[0]\
                        for (ui,fi,ei) in zip(self.u,self.f,self.e)])
+
+    def negll(self,par):
+        return -1*self.log_likelihood(par)
+
+    def optim_fmin(self,p0,maxiter=1000,ftol=0.0001,xtol=0.0001):
+        opt = op.fmin(self.negll, p0, maxiter=maxiter,ftol=ftol)
+        #opt = op.fmin_bfgs(self.negll, p0, gtol=ftol, maxiter=maxiter)
+        return opt
+
 
 def gaussian_pdf(x):
     return (1 / (np.sqrt(2.*np.pi))) * np.exp(-(x**2) / (2.))    
@@ -93,5 +103,5 @@ def gamma_cdf(x,mean,var):
 
 
 ### TODO: 1. intelligently choose limits for gamma integral
-###       2. check that my gamma interpretation is right
-###       3. audit all gamma and norm calls
+###       2. check that my gamma interpretation is right; DONE
+###       3. audit all gamma and norm calls; DONE, replaced by hand-built functions

### 2012-04-12 py/test_censoring2.py

@@ -1,0 +1,110 @@
+
+import censoring2 as c2
+import cProfile
+import scipy.stats as stats
+import numpy as np
+import scipy.optimize as op
+import matplotlib.pylab as plt
+
+
+def plot_llmarg(cens, pstar, ind=0, p0=0, pn=1, n=11, name='P'):
+
+    ll_p = []
+    true = pstar[ind]
+    parval = np.linspace(p0,pn,n)
+    for pp in parval:
+        ptmp = pstar + []; ptmp[ind] = pp
+        ll_p.append(cens.log_likelihood(ptmp))
+    plt.plot(parval, ll_p, '-o')
+    plt.xlabel(name)
+    plt.ylabel('log likelihood')
+    plt.text(0.5, 0.1, 'True value = ' + str(round(true,3)), transform = ax.transAxes)
+    plt.show()
+
+
+if __name__ == "__main__":
+
+    import fake_data as fd
+    
+    A0 = 5.; A1 = 1.;    B1 = 1.
+    P = 400.
+    w = (2 * np.pi) / P
+    s2mu = 0.2**2
+    S, VS = 0.7, 0.15**2
+    B, VB = 4.5, 0.4**2
+    Vsigma = 0.2**2
+
+    gt, f, s2, bt = fd.make_fake_data(3600. * np.random.uniform(size=250), \
+                                      w, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS)
+
+    cmodel = c2.Censored(gt,f,s2,bt,tolquad=1.)
+    popt = [P,A0,A1,B1,s2mu,B,VB,Vsigma,S,VS]
+
+    prof = True
+    if(prof): # profile code?
+        cProfile.run("cmodel.log_likelihood(popt)")
+
+
+    plotmarg = False
+
+    if(plotmarg):
+        # make a plot of ll vs. parameter
+        fig = plt.figure(figsize=(16, 10))
+        fig.subplots_adjust(hspace=0.7)
+        fig.subplots_adjust(wspace=0.2)
+
+        ax = fig.add_subplot(5,2,1)
+        plot_llmarg(cmodel,popt,0,p0=350, pn = 450, n=11, name='Period')
+
+        ax = fig.add_subplot(5,2,2)
+        plot_llmarg(cmodel,popt,1,p0=2, pn = 8, n=11, name='A0')
+
[…]

### 2012-04-12 py/censoring2.py
fixed cdf for loglikelihood_observed. now using gaussian cdf instead of gamma. now matches cdf for censored data (which was changed for computational reasons)

@@ -67,7 +67,7 @@ class Censored:
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
         def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
-            p_not_cens = gamma_cdf(fi,B,VB)#fi,B**2/VB, scale = VB/B)
+            p_not_cens = gaussian_cdf(fi,B,VB)#fi,B**2/VB, scale = VB/B)
             p_flux = gaussian_pdf((fi - ui) / np.sqrt(sig2 + su2))
             p_si = gamma_pdf(ei,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)

### 2012-04-12 py/fake_data.py
in fake_data Gamma ---> Norm for bs

@@ -36,7 +36,7 @@ def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     sigma2s = Gamma.rvs(S**2/VS, scale=VS/S, size=times.size)
     sobs2s = np.array([Gamma.rvs(m**2/Vsigma, scale=Vsigma/m, size=1)[0] for m in sigma2s])
     fobss = mus + np.sqrt(sigma2s + s2mu) * Norm.rvs(size=times.size)
-    bs = Gamma.rvs(B**2/VB, scale=VB/B, size=times.size)
+    bs = B + np.sqrt(VB) * Norm.rvs(size=times.size)
     good = fobss > bs
     bad = good == False
     return times[good], fobss[good], sobs2s[good], times[bad]

### 2012-04-12 py/test_censoring2.py

@@ -45,7 +45,7 @@ if __name__ == "__main__":
         cProfile.run("cmodel.log_likelihood(popt)")
 
 
-    plotmarg = False
+    plotmarg = True
 
     if(plotmarg):
         # make a plot of ll vs. parameter
@@ -69,13 +69,13 @@ if __name__ == "__main__":
         plot_llmarg(cmodel,popt,4,p0=0.01, pn = 0.07, n=11, name=r'$s_{\mu}^2$')
 
         ax = fig.add_subplot(5,2,6)
-        plot_llmarg(cmodel,popt,5,p0=3, pn = 6, n=11, name='B')
+        plot_llmarg(cmodel,popt,7,p0=0.01, pn = 0.07, n=11, name=r'$V_{\sigma}$')
 
         ax = fig.add_subplot(5,2,7)
-        plot_llmarg(cmodel,popt,6,p0=0.1, pn = 0.22, n=11, name=r'$V_B$')
+        plot_llmarg(cmodel,popt,5,p0=3, pn = 6, n=11, name='B')
 
         ax = fig.add_subplot(5,2,8)
-        plot_llmarg(cmodel,popt,7,p0=0.01, pn = 0.07, n=11, name=r'$V_{\sigma}$')
+        plot_llmarg(cmodel,popt,6,p0=0.1, pn = 0.22, n=11, name=r'$V_B$')
 
         ax = fig.add_subplot(5,2,9)
         plot_llmarg(cmodel,popt,8,p0=0.4, pn = 1.0, n=11, name='S')

### 2012-04-12 py/censoring2.py
fixed bug with call to gaussian_cdf

@@ -67,7 +67,7 @@ class Censored:
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
         def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
-            p_not_cens = gaussian_cdf(fi,B,VB)#fi,B**2/VB, scale = VB/B)
+            p_not_cens = gaussian_cdf((fi-B)/np.sqrt(VB + sig2 + su2))#fi,B**2/VB, scale = VB/B)
             p_flux = gaussian_pdf((fi - ui) / np.sqrt(sig2 + su2))
             p_si = gamma_pdf(ei,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)

### 2012-04-12 py/censoring2.py
added pdf/cdf unit tests and fixed bugs in those definitions

@@ -1,9 +1,13 @@
-###
-### implements censoring model
-### date April 10, 2012
+"""
+This file is part of the Censored Data Project
+Copyright 2012, Joseph W. Richards, David W. Hogg, James Long
 
-### Updated by JWR on 20120411
-###
+This code is synchronized with the document tex/ms.tex
+
+Issues:
+- limits on gamma integrations are made up (sensitivity not checked)
+- make it possible to insert estimated (or known) upper limits
+"""
 
 import numpy as np
 #import scipy.stats as stats
@@ -31,6 +35,8 @@ class Censored:
         self.f = f
         self.e = e
         self.tc = tc
+
+        unittests()
         
     def log_likelihood(self,par):
         """ 
@@ -60,15 +66,14 @@ class Censored:
     def loglikelihood_censored(self,su2,B,VB,S,VS):
     ## integrand of equation (12), integrate this across sig2
         def sig_integrand(sig2,ui,su2,B,VB,S,VS):
-            return  gaussian_cdf((B-ui) / np.sqrt(VB + sig2 + su2)) * \
-                gamma_pdf(sig2,S,VS)
+            return  gaussian_cdf(B, ui, VB + sig2 + su2) * gamma_pdf(sig2,S,VS)
         return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0] \
                        for ui in self.uc])
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
         def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
-            p_not_cens = gaussian_cdf(fi,B,VB)#fi,B**2/VB, scale = VB/B)
-            p_flux = gaussian_pdf((fi - ui) / np.sqrt(sig2 + su2))
+            p_not_cens = gaussian_cdf(fi, B, VB)
+            p_flux = gaussian_pdf(fi, ui, sig2 + su2)
             p_si = gamma_pdf(ei,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
@@ -83,17 +88,20 @@ class Censored:
[…]

### 2012-04-12 py/fake_data.py

@@ -1,6 +1,8 @@
 '''
 This file is part of the Censored Data project.
-Copyright 2012 David W. Hogg (NYU).
+Copyright 2012 David W. Hogg (NYU), Joseph W. Richards, and James Long (UCB)
+
+Should be synced with censoring2.py
 
 issues:
 -------

### 2012-04-12 py/test_censoring2.py

@@ -24,17 +24,19 @@ def plot_llmarg(cens, pstar, ind=0, p0=0, pn=1, n=11, name='P'):
 
 if __name__ == "__main__":
 
+    np.random.seed(47)
+
     import fake_data as fd
     
     A0 = 5.; A1 = 1.;    B1 = 1.
     P = 400.
     w = (2 * np.pi) / P
-    s2mu = 0.2**2
+    s2mu = 0.5**2
     S, VS = 0.7, 0.15**2
     B, VB = 4.5, 0.4**2
     Vsigma = 0.2**2
 
-    gt, f, s2, bt = fd.make_fake_data(3600. * np.random.uniform(size=250), \
+    gt, f, s2, bt = fd.make_fake_data(3600. * np.random.uniform(size=512), \
                                       w, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS)
 
     cmodel = c2.Censored(gt,f,s2,bt,tolquad=1.)
@@ -66,7 +68,7 @@ if __name__ == "__main__":
         plot_llmarg(cmodel,popt,3,p0=0.5, pn = 1.5, n=11, name='B1')
 
         ax = fig.add_subplot(5,2,5)
-        plot_llmarg(cmodel,popt,4,p0=0.01, pn = 0.07, n=11, name=r'$s_{\mu}^2$')
+        plot_llmarg(cmodel,popt,4,p0=0, pn = 0.5, n=11, name=r'$s_{\mu}^2$')
 
         ax = fig.add_subplot(5,2,6)
         plot_llmarg(cmodel,popt,7,p0=0.01, pn = 0.07, n=11, name=r'$V_{\sigma}$')

### 2012-04-12 py/censoring2.py
flux in micromaggies

@@ -22,7 +22,7 @@ import scipy.optimize as op
 
 class Censored:
 
-    def __init__(self, t, f, e, tc, tolquad = 0.1):
+    def __init__(self, t, m, em, tc, tolquad = 0.1):
         """ 
         t: times when not censored (1-d ndarray)
         f: flux when not censored (1-d ndarray)
@@ -32,9 +32,12 @@ class Censored:
         self.tolquad = tolquad # numerical tolerance in integration
         # insert data 
         self.t = t
-        self.f = f
-        self.e = e
+        self.m = m
+        self.em = em
         self.tc = tc
+
+        self.f = mag2flux(self.m)
+        self.ef = magerr2fluxerr(self.m, self.em)
 
         unittests()
         
@@ -78,7 +81,7 @@ class Censored:
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
         return np.log([integrate.quad(integrand,0.,S+5.*np.sqrt(VS),(ui,fi,ei,su2,B,VB,Vsig,S,VS),epsabs=self.tolquad)[0]\
-                       for (ui,fi,ei) in zip(self.u,self.f,self.e)])
+                       for (ui,fi,ei) in zip(self.u,self.f,self.ef)])
 
     def negll(self,par):
         return -1*self.log_likelihood(par)
@@ -108,6 +111,16 @@ def gamma_cdf(x,mean,var):
     k = mean / theta
     return  gammainc(k,x/theta) ## wikipedia and scipy define imcomplete gamma differently
 
+# flux to mag conversions
+
+def mag2flux(m, f0 = 1.e-6):
+    return f0 * 10**(-0.4*m)
+
+def flux2mag(f, f0 = 1.e-6):
+    return -2.5 * np.log10(f/f0)
+
+def magerr2fluxerr(m, merr, f0 = 1.e-6):
+    return 0.5 * (mag2flux(m - merr, f0) - mag2flux(m + merr, f0))
 
 def unittests():
     assert(np.abs(1. - integrate.quad(gaussian_pdf,-20.,20.,(1.322,1.532))[0]) < 1.e-7)
@@ -118,6 +131,10 @@ def unittests():
     assert(np.abs(1. - gamma_cdf(100,0.332,1.323)) < 1.e-7)
     assert(np.abs(gamma_cdf(0,0.332,1.323)) < 1.e-7)
[…]

### 2012-04-12 py/fake_data.py
improved plotting

@@ -14,7 +14,7 @@ if __name__ == '__main__':
     import matplotlib
     matplotlib.use('Agg')
     from matplotlib import rc
-    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':12})
+    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
     rc('text', usetex=True)
 import matplotlib.pyplot as plt
 import numpy as np
@@ -43,16 +43,36 @@ def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     bad = good == False
     return times[good], fobss[good], sobs2s[good], times[bad]
 
-def plot_data(ax, goodtimes, fobss, sobs2s, badtimes):
-    ax.errorbar(goodtimes, fobss, fmt='ko', yerr=np.sqrt(sobs2s))
-    ax.plot(badtimes, np.zeros_like(badtimes), 'ro')
+def hogg_errorbar(ax, x, y, yerr, color='k', alpha=0.25):
+    for xi,yi,yerri in zip(x,y,yerr):
+        ax.plot([xi, xi], [yi - yerri, yi + yerri], color+'-', alpha=alpha)
     return None
 
-def plot_model(ax, omega, A0, A1, B1):
-    tlim = ax.get_xlim()
+def plot_data(ax, goodtimes, fobss, sobs2s, badtimes, title=''):
+    alltimes = np.append(goodtimes, badtimes)
+    mediant = np.round(np.median(alltimes)).astype(int)
+    alltimes -= mediant
+    sobss = np.sqrt(sobs2s)
+    ax.axhline(0., color='k', alpha=0.25)
+    ax.plot(goodtimes - mediant, fobss, 'ko', alpha=0.5, mec='k')
+    hogg_errorbar(ax, goodtimes - mediant, fobss, sobss)
+    ax.plot(badtimes - mediant, np.zeros_like(badtimes), 'r.', alpha=0.5, mec='r')
+    foo = np.max(fobss + sobss)
+    ax.set_xlim(np.min(alltimes), np.max(alltimes))
+    ax.set_ylim(-0.1 * foo, 1.1 * foo)
+    ax.set_xlabel(r'time $t$ (MJD - %d~d)' % mediant)
+    ax.set_ylabel(r'flux $f$ ($\mu$Mgy)')
+    ax.set_title(title)
[…]

### 2012-04-12 py/censoring2.py
in Gamma pdf, ei should be ei2; fixed this

@@ -38,6 +38,7 @@ class Censored:
 
         self.f = mag2flux(self.m)
         self.ef = magerr2fluxerr(self.m, self.em)
+        self.ef2 = self.ef**2
 
         unittests()
         
@@ -74,14 +75,14 @@ class Censored:
                        for ui in self.uc])
 
     def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
-        def integrand(sig2,ui,fi,ei,su2,B,VB,Vsig,S,VS):
+        def integrand(sig2,ui,fi,ei2,su2,B,VB,Vsig,S,VS):
             p_not_cens = gaussian_cdf(fi, B, VB)
             p_flux = gaussian_pdf(fi, ui, sig2 + su2)
-            p_si = gamma_pdf(ei,sig2,Vsig)
+            p_si = gamma_pdf(ei2,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
-        return np.log([integrate.quad(integrand,0.,S+5.*np.sqrt(VS),(ui,fi,ei,su2,B,VB,Vsig,S,VS),epsabs=self.tolquad)[0]\
-                       for (ui,fi,ei) in zip(self.u,self.f,self.ef)])
+        return np.log([integrate.quad(integrand,0.,S+5.*np.sqrt(VS),(ui,fi,ei2,su2,B,VB,Vsig,S,VS),epsabs=self.tolquad)[0]\
+                       for (ui,fi,ei2) in zip(self.u,self.f,self.ef2)])
 
     def negll(self,par):
         return -1*self.log_likelihood(par)

### 2012-04-12 py/censoring2.py
separate mu function, some plotting

@@ -35,16 +35,19 @@ class Censored:
         self.m = m
         self.em = em
         self.tc = tc
-
         self.f = mag2flux(self.m)
         self.ef = magerr2fluxerr(self.m, self.em)
-
         unittests()
+        return None
+
+    def mu(self, times, omega, A0, A1, B1):
+        return A0 + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
         
     def log_likelihood(self,par):
         """ 
         computes log p(D|theta,I)
         everything else: same notation as paper (all floats)
+        hella ugly unpacking from par[]
         """
         P = par[0]
         A0 = par[1]
@@ -56,11 +59,11 @@ class Censored:
         Vsig = par[7]
         S = par[8]
         VS = par[9]
-        
+
     ## convert all times to expected flux (i.e. t_i -> u_i)
         w =  (2 * np.pi)/ P 
-        self.u = A0 + A1*np.cos(self.t*w) + B1*np.sin(self.t*w)
-        self.uc = A0 + A1*np.cos(self.tc*w) + B1*np.sin(self.tc*w)
+        self.u  = self.mu(self.t,  w, A0, A1, B1)
+        self.uc = self.mu(self.tc, w, A0, A1, B1)
 
     ## compute loglikelihood
         return np.sum(self.loglikelihood_censored(su2,B,VB,S,VS)) + \
@@ -90,6 +93,47 @@ class Censored:
         opt = op.fmin(self.negll, p0, maxiter=maxiter,ftol=ftol)
         #opt = op.fmin_bfgs(self.negll, p0, gtol=ftol, maxiter=maxiter)
         return opt
+
+    def plot(self, ax, par):
+        '''
+        input:
+        - ax: matplotlib axes object
+        - par: set of hyperparameters
+
+        usage:
+            plt.clf()
+            ax = plt.subplot(111)
+            cens.plot(ax, par)
+            plt.savefig('wtf.png')
+        '''
+        omega = 2. * np.pi / par[0]
+        A0 = par[1]
+        A1 = par[2]
+        B1 = par[3]
+        s2mu = par[4]
+        alltimes = np.append(self.t, self.tc)
+        mediant = np.round(np.median(alltimes)).astype(int)
+        ax.axhline(0., color='k', alpha=0.25)
[…]

### 2012-04-12 py/fake_data.py

@@ -43,11 +43,6 @@ def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     bad = good == False
     return times[good], fobss[good], sobs2s[good], times[bad]
 
-def hogg_errorbar(ax, x, y, yerr, color='k', alpha=0.25):
-    for xi,yi,yerri in zip(x,y,yerr):
-        ax.plot([xi, xi], [yi - yerri, yi + yerri], color+'-', alpha=alpha)
-    return None
-
 def plot_data(ax, goodtimes, fobss, sobs2s, badtimes, title=''):
     alltimes = np.append(goodtimes, badtimes)
     mediant = np.round(np.median(alltimes)).astype(int)

### 2012-04-12 py/censored_code.py
hating on deprecated code

@@ -1,233 +1,0 @@
-##################
-################## OBSOLETE CODE ---> WORKING CODE NOW IN censoring.py
-##################
-
-
-
-########## 
-########## 
-########## CENSORED REGRESSION ALGORITHM 
-########## 
-##########
-########## by James Long 
-########## date: 11/8/2011 
-########## 
-
-import cProfile
-import numpy as np
-import scipy.optimize
-import scipy.stats
-import visualize
-import pstats
-from scipy import special
-
-def gaussian_pdf_1d(x,mean,var):
-    return (1./np.sqrt(2.* np.pi * var))*np.exp(-(0.5*(x-mean)**2)/var)
-
-def gaussian_cdf_1d(x,mean,var):
-    return .5*(1. + special.erf((x-mean)/np.sqrt(2.*var)))
-
-# physical explanation, data, 
-class LightCurve:
-    def __init__(self, data, model, prior):
-        self.data = data
-        self.model = model
-        self.prior = prior
-    ## makes a plot of the light curv
-    def plot(self,true=False,naive=False,hrl=False):
-        visualize.plot_curve(self.data.get_all(),true=true,naive=naive,hrl=hrl,model=self.model)
-    def naive_optimize(self):
-        tfe = self.data.get_all()
-        tfe = tfe[tfe[:,3]==1,:]
-        def cost(p, const):
-            self.model.naive.set_opt_params(p)
-            return ((self.model.naive.mean(tfe[:,0]) - tfe[:,1]) / tfe[:,2])**2
-        p = self.model.naive.get_opt_params()
-        const = self.model.naive.get_const_params()
-        betterp = scipy.optimize.leastsq(cost, p, const)
-    def ln_likelihood(self):
-        self.bs = self.prior.bsample(100)
-        self.sigmas = self.prior.sigmasample(100)
-        return sum(map(lambda datum: self.ln_likelihood_point(datum,self.bs,self.sigmas), self.data))
-    #compute p(D_n|theta) 
-    def ln_likelihood_point(self, datum, bs, sigmas):
-        return np.log(np.mean(self.likelihood_point(datum,self.bs,self.sigmas)))
-    def likelihood_point(self, datum, bs, sigmas):
-        t, q, f, s = datum
-        mu = self.model.hrl.mean(t)
-        variances = self.model.hrl.get_modelvariance() + sigmas**2
-        if q == 1:
-            ## p(qi|b,sigma,theta,I) * p(f|q,sigma,theta,I) * p(s|q,sigma,theta,I)   where qi = 1
[…]

### 2012-04-12 py/fake_data.py
output mags not fluxes

@@ -29,6 +29,10 @@ def mu(times, omega, A0, A1, B1):
         + A1 * np.cos(omega * times) \
         + B1 * np.sin(omega * times)
 
+def flux2mag(f):
+    f0 = 1.e-6
+    return -2.5 * np.log10(fobss[good] / f0)
+
 def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     '''
     Implement the full generative model from the paper.  Is this
@@ -41,43 +45,8 @@ def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     bs = B + np.sqrt(VB) * Norm.rvs(size=times.size)
     good = fobss > bs
     bad = good == False
-    return times[good], fobss[good], sobs2s[good], times[bad]
-
-def plot_data(ax, goodtimes, fobss, sobs2s, badtimes, title=''):
-    alltimes = np.append(goodtimes, badtimes)
-    mediant = np.round(np.median(alltimes)).astype(int)
-    alltimes -= mediant
-    sobss = np.sqrt(sobs2s)
-    ax.axhline(0., color='k', alpha=0.25)
-    ax.plot(goodtimes - mediant, fobss, 'ko', alpha=0.5, mec='k')
-    hogg_errorbar(ax, goodtimes - mediant, fobss, sobss)
-    ax.plot(badtimes - mediant, np.zeros_like(badtimes), 'r.', alpha=0.5, mec='r')
-    foo = np.max(fobss + sobss)
-    ax.set_xlim(np.min(alltimes), np.max(alltimes))
-    ax.set_ylim(-0.1 * foo, 1.1 * foo)
-    ax.set_xlabel(r'time $t$ (MJD - %d~d)' % mediant)
-    ax.set_ylabel(r'flux $f$ ($\mu$Mgy)')
-    ax.set_title(title)
-    return mediant
-
-def plot_model(ax, t_offset, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
-    tlim = np.array(ax.get_xlim()) + t_offset
-    flim = ax.get_ylim()
-    tp = np.linspace(tlim[0], tlim[1], 10000)
-    fp = mu(tp, omega, A0, A1, B1)
-    ax.plot(tp - t_offset, fp + np.sqrt(s2mu), 'b-', alpha=0.25)
[…]

### 2012-04-13 py/censoring2.py
added init_par method and fmin

@@ -22,7 +22,7 @@ import scipy.optimize as op
 
 class Censored:
 
-    def __init__(self, t, m, em, tc, tolquad = 0.1):
+    def __init__(self, t, m, em, tc, name = "Awesome LC", tolquad = 0.1):
         """ 
         t: times when not censored (1-d ndarray)
         f: flux when not censored (1-d ndarray)
@@ -38,6 +38,8 @@ class Censored:
         self.f = mag2flux(self.m)
         self.ef = magerr2fluxerr(self.m, self.em)
         self.ef2 = self.ef**2
+        self.name = name
+        
         unittests()
         return None
 
@@ -48,7 +50,7 @@ class Censored:
         """ 
         computes log p(D|theta,I)
         everything else: same notation as paper (all floats)
-        hella ugly unpacking from par[]
+        hella ugly unpacking from par[]; you're so hella NorCal
         """
         P = par[0]
         A0 = par[1]
@@ -89,6 +91,30 @@ class Censored:
 
     def negll(self,par):
         return -1*self.log_likelihood(par)
+
+    def get_init_par(self,Period):
+        # params:     P,A0,A1,B1,su2,B,VB,Vsig,S,VS
+        par = np.zeros(10)
+        par[0] = Period
+        par[1:4] = self.least_sq(Period)
+        par[4] = (0.2 * par[1])**2
+        par[5] = 2.*np.abs(np.min(self.f))
+        par[6] = par[5]**2
+        par[7] = np.median(self.ef2)
+        par[8] = par[7]
+        par[9] = par[8]**2.
+        return par
+
+    def least_sq(self,Period):
+        Amat = np.zeros( (3,len(self.t)))
+        Amat[0,:] = 1.
+        Amat[1,:] = self.mu(self.t,2.*np.pi / Period, 0,1,0)
+        Amat[2,:] = self.mu(self.t,2.*np.pi / Period, 0,0,1)
+        Atb = np.dot(Amat, self.f)
+        print Atb.shape
+        AtAinv = np.matrix(np.dot(Amat,Amat.T)).I
+        return np.dot(AtAinv,Atb)
+        
 
     def optim_fmin(self,p0,maxiter=1000,ftol=0.0001,xtol=0.0001):
[…]

### 2012-04-13 py/fake_data.py

@@ -10,13 +10,6 @@ issues:
 - not properly commented
 '''
 
-if __name__ == '__main__':
-    import matplotlib
-    matplotlib.use('Agg')
-    from matplotlib import rc
-    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
-    rc('text', usetex=True)
-import matplotlib.pyplot as plt
 import numpy as np
 from scipy.stats import norm as Norm
 from scipy.stats import gamma as Gamma
@@ -30,8 +23,8 @@ def mu(times, omega, A0, A1, B1):
         + B1 * np.sin(omega * times)
 
 def flux2mag(f):
-    f0 = 1.e-6
-    return -2.5 * np.log10(fobss[good] / f0)
+    f0 = 1.e6
+    return -2.5 * np.log10(f / f0)
 
 def make_fake_data(times, omega, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS):
     '''

### 2012-04-13 py/test_censoring2.py

@@ -1,4 +1,11 @@
 
+if __name__ == '__main__':
+    import matplotlib
+    matplotlib.use('Agg')
+    from matplotlib import rc
+    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
+    rc('text', usetex=True)
+
 import censoring2 as c2
 import cProfile
 import scipy.stats as stats
@@ -28,19 +35,33 @@ if __name__ == "__main__":
 
     import fake_data as fd
     
-    A0 = 5.; A1 = 1.;    B1 = 1.
+    A0 = 100.; A1 = 50.;    B1 = 50.
     P = 400.
-    w = (2 * np.pi) / P
-    s2mu = 0.5**2
-    S, VS = 0.7, 0.15**2
-    B, VB = 4.5, 0.4**2
-    Vsigma = 0.2**2
+    w = (2. * np.pi) / P
+    s2mu = 25.**2
+    S, VS = 25., 5.**2
+    B, VB = 100., 20.**2
+    Vsigma = 10.**2
 
-    gt, f, s2, bt = fd.make_fake_data(3600. * np.random.uniform(size=512), \
+    gt, m, mer, bt = fd.make_fake_data(3600. * np.random.uniform(size=512), \
                                       w, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS)
-
-    cmodel = c2.Censored(gt,f,s2,bt,tolquad=1.)
+    
+    cmodel = c2.Censored(gt,m,mer,bt,tolquad=1.)
     popt = [P,A0,A1,B1,s2mu,B,VB,Vsigma,S,VS]
+
+    p0 = cmodel.get_init_par(400.)
+    print p0
+
+    plt.clf()
+    ax = plt.subplot(111)
+    cmodel.plot(ax,p0)
+    plt.savefig('init_params.png')
+
+    plt.clf()
+    ax = plt.subplot(111)
+    cmodel.plot(ax,popt)
+    plt.savefig('true_params.png')
+
 
     prof = True
     if(prof): # profile code?
@@ -56,43 +77,49 @@ if __name__ == "__main__":
         fig.subplots_adjust(wspace=0.2)
 
         ax = fig.add_subplot(5,2,1)
-        plot_llmarg(cmodel,popt,0,p0=350, pn = 450, n=11, name='Period')
[…]

### 2012-04-13 py/censoring2.py
folded plot added

@@ -122,7 +122,7 @@ class Censored:
         #opt = op.fmin_bfgs(self.negll, p0, gtol=ftol, maxiter=maxiter)
         return opt
 
-    def plot(self, ax, par):
+    def plot(self, ax, par, fold = False):
         '''
         input:
         - ax: matplotlib axes object
@@ -134,27 +134,43 @@ class Censored:
             cens.plot(ax, par)
             plt.savefig('wtf.png')
         '''
+        def hogg_errorbar(ax, x, y, yerr, color='k', alpha=0.25):
+            for xi,yi,yerri in zip(x,y,yerr):
+                ax.plot([xi, xi], [yi - yerri, yi + yerri], color+'-', alpha=alpha)
+            return None
+        def phase(t, P):
+            return (t/P) % 1.
+        if(fold):
+            x, xc = phase(self.t, par[0]), phase(self.tc, par[0])
+            mediant = 0
+            tlim = np.array([0., 2.])
+            tp = np.linspace(0., par[0]*2., 10000)
+            xp = np.linspace(0., 2., len(tp))
+        else:
+            x, xc = self.t, self.tc
+            alltimes = np.append(self.t, self.tc)
+            mediant = np.round(np.median(alltimes)).astype(int)
+            tlim = np.array([np.min(alltimes), np.max(alltimes)])
+            tp = np.linspace(tlim[0], tlim[1], 10000)
+            xp = tp
         omega = 2. * np.pi / par[0]
         A0 = par[1]
         A1 = par[2]
         B1 = par[3]
         s2mu = par[4]
-        alltimes = np.append(self.t, self.tc)
-        mediant = np.round(np.median(alltimes)).astype(int)
         ax.axhline(0., color='k', alpha=0.25)
-        ax.plot(self.t - mediant, self.f, 'ko', alpha=0.5, mec='k')
-        def hogg_errorbar(ax, x, y, yerr, color='k', alpha=0.25):
-            for xi,yi,yerri in zip(x,y,yerr):
-                ax.plot([xi, xi], [yi - yerri, yi + yerri], color+'-', alpha=alpha)
-            return None
-        hogg_errorbar(ax, self.t - mediant, self.f, self.ef)
[…]

### 2012-04-13 py/test_censoring2.py

@@ -59,8 +59,18 @@ if __name__ == "__main__":
 
     plt.clf()
     ax = plt.subplot(111)
+    cmodel.plot(ax,p0, fold=True)
+    plt.savefig('init_params_folded.png')
+
+    plt.clf()
+    ax = plt.subplot(111)
     cmodel.plot(ax,popt)
     plt.savefig('true_params.png')
+
+    plt.clf()
+    ax = plt.subplot(111)
+    cmodel.plot(ax,popt, fold=True)
+    plt.savefig('true_params_folded.png')
 
 
     prof = True
@@ -68,7 +78,7 @@ if __name__ == "__main__":
         cProfile.run("cmodel.log_likelihood(popt)")
 
 
-    plotmarg = True
+    plotmarg = False
 
     if(plotmarg):
         # make a plot of ll vs. parameter
@@ -108,7 +118,7 @@ if __name__ == "__main__":
         plt.savefig('likelihood_marg.png')
 
 
-    maxlik = True
+    maxlik = False
     if(maxlik):
         #p0 = popt
         #p0 = popt + stats.norm.rvs(0,.01,size=10)

### 2012-04-13 py/censoring2.py
changed init_par function because variances were to large, causing fmin to run too slowly

@@ -101,10 +101,10 @@ class Censored:
         par[1:4] = self.least_sq(Period)
         par[4] = (0.2 * par[1])**2
         par[5] = 2.*np.abs(np.min(self.f))
-        par[6] = par[5]**2
+        par[6] = par[5]
         par[7] = np.median(self.ef2)
         par[8] = par[7]
-        par[9] = par[8]**2.
+        par[9] = par[8] # must be kept small or integration will take too long
         return par
 
     def least_sq(self,Period):
@@ -117,8 +117,8 @@ class Censored:
         AtAinv = np.matrix(np.dot(Amat,Amat.T)).I
         return np.dot(AtAinv,Atb)
         
-    def optim_fmin(self,p0,maxiter=1000,ftol=0.0001,xtol=0.0001):
-        opt = op.fmin(self.negll, p0, maxiter=maxiter,ftol=ftol)
+    def optim_fmin(self,p0,maxiter=1000,ftol=0.0001,xtol=0.0001,mfev=1.e8):
+        opt = op.fmin(self.negll, p0, maxiter=maxiter,ftol=ftol,maxfun=mfev)
         #opt = op.fmin_bfgs(self.negll, p0, gtol=ftol, maxiter=maxiter)
         return opt
 

### 2012-04-13 py/test_censoring2.py

@@ -39,9 +39,9 @@ if __name__ == "__main__":
     P = 400.
     w = (2. * np.pi) / P
     s2mu = 25.**2
-    S, VS = 25., 5.**2
     B, VB = 100., 20.**2
-    Vsigma = 10.**2
+    Vsigma = 5.**2
+    S, VS = 25., 5.**2
 
     gt, m, mer, bt = fd.make_fake_data(3600. * np.random.uniform(size=512), \
                                       w, A0, A1, B1, s2mu, B, VB, Vsigma, S, VS)
@@ -118,17 +118,29 @@ if __name__ == "__main__":
         plt.savefig('likelihood_marg.png')
 
 
-    maxlik = False
+
+#[400.0, 100.0, 50.0, 50.0, 625.0, 100.0, 400.0, 100.0, 25.0, 25.0]
+#[ 400.          117.04071511   35.79569983   34.50307783  547.94115976
+#  121.10564877  121.10564877   23.00652436   23.00652436  529.30016307]
+
+    maxlik = True
     if(maxlik):
         #p0 = popt
         #p0 = popt + stats.norm.rvs(0,.01,size=10)
-        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=0.5,xtol=0.1)
-        #pfmin = cmodel.optim_fmin(p0,maxiter=100,ftol=0.01)
+        print p0
+        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=0.5,xtol=0.1,mfev=500)
+        print cmodel.log_likelihood(pfmin)
+        print cmodel.log_likelihood(p0)
 
         plt.clf()
         ax = plt.subplot(111)
         cmodel.plot(ax,pfmin)
         plt.savefig('bestfit_params.png')
+
+        plt.clf()
+        ax = plt.subplot(111)
+        cmodel.plot(ax,pfmin, fold=True)
+        plt.savefig('bestfit_params_folded.png')
 
         print pfmin
         print popt

### 2012-04-13 tex/ms.tex
in fake_data Gamma ---> Norm for bs
james: substantially revised model section, breaking it into non-detections and error modelling parts

@@
 perform photometric measurements anyway, so that there is a datum
-recorded at at every epoch.  In practice, most surveys only provide
+recorded at every epoch.  In practice, most surveys only provide
 detection upper limits at some epochs; in some not even upper
@@
 Unless a particular region of the sky gets special treatment (such as it is a known place of interest), discovery of variability happens by comparison of flux values at the database level.
-Analysis on such databases of transient/variables thus deals with censored data where there is often less information available that there should be.
+Analysis on such databases of transient/variables thus deals with censored data where there is often less information available than there should be.
 Even then, we might be able to infer what the thresholds were based on S/N measurements of objects neighboring the source of interest.
@@
 
-At each time $t_i$, the star is either detected ($q_i=1$) by the
-software or not ($q_i=0$).  When it is detected, the software returns
+At each time $t_i$, the star is either detected by the
+software or not.  When it is detected, the software returns
 a (possibly bad) flux value $f_i$ and (likely bad) uncertainty
-variance $s_i^2$.  When the source is \emph{not} detected ($q_i=0$),
+variance $s_i^2$.  When the source is \emph{not} detected,
 nothing is reported---in this case, because the software is not
@@
 At each epoch $i$, with associated time $t_i$, the survey takes an exposure at the location of the object and either (a) detects the object and records an estimate of its photon flux, $\fobs$ and the  variance of the statistical uncertainty in that estimate, $\sobs$, or (b) fails to detect the object.
-In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, $b_i$, which is the brightest the object could have been given that it was not detected at a significant level by the software.
+In the latter case, most modern surveys either record a reference value to signify that no detection was made (as is done in ASAS) or an estimate of the upper detection limit, $b_i$, which is the minimum brightness that could have been detected.
 
@@
 
-\subsection{The Light Curve Model}
-
+\section{The Light Curve Model}
+\label{sec:model}
+\subsection{Modeling Non-Detections}
 To model the mean brightness of the light curve as a function of time, we use a multiple-harmonic Fourier model with angular oscillation frequency $\omega$,
@@
 \end{eqnarray}
-where $\mu_i$ is the flux at time $t_i$, $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(B_k,A_k)$ is the relative phase offset of harmonic $k$.
+where $\mu_i$ is the flux at time $t_i$, $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(\frac{B_k}{A_k})$ is the relative phase offset of harmonic $k$.
 The number of harmonics, $K$, can either be fixed or treated as a free parameter over which to optimize the likelihood.
-In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized).
 
-Additionally, we need to instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.
-The $b_i$ parameter is essential because it constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.
+In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized) as well as random variation due to photo arrival counts, assumed normal mean 0 with variance $\sigma_i^2$.
+The reported uncertainties $s_i^2$ are an estimate of the $\sigma_i^2$, see Subsection \ref{sec:uncertainty}.
+This leads to a true flux measurement of
+\begin{eqnarray}\displaystyle
+\ftrue = \mu_i + e_i
+\label{eq:truefluxdef}
+\end{eqnarray}
+where $e_i \sim N(0,s_\mu^2 + \sigma_i^2)$.
+Recall that this is a latent variable because we will not see it when an observation is censored.
+We instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.
+The $b_i$ parameter is essential because it probabilistically constrains the possible values of the mean brightness, $\mu_i$, of the light curve when there is a non-detection.
 By employing a hierarchical model for the distribution of $b_i$, we can fully utilize all of the information encoded in both the detections and non-detections when computing the data likelihood.
-At each epoch, we connect the observed flux, $\fobs$, to the latent variable, $\ftrue$, signifying the true observable flux of the object, via the detection limit, $b_i$ by
+At each epoch, we connect the observed flux, $\fobs$, to the latent variable $\ftrue$ by
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-so that the observed flux is \texttt{NA} only when the true observable flux, $\ftrue$, is below the detection limit, $b_i$.
+The observed flux is \texttt{NA} only when the true flux, $\ftrue$, is below the detection limit $b_i$.
 
-Also, because we do not necessarily believe the reported measurement uncertainties, $\sobs$, we choose to introduce a parameter, $\sigma^2_i$, to represent the true uncertainty variance for the brightness measurement at epoch $i$.
-
-Hence, our initial model consists of the parameter vectors $\{f^*_1,...,f^*_N\}$, $\{b_1,...,b_N\}$ and $\{\sigma^2_1,...,\sigma^2_N\}$, and the model parameters
-\begin{eqnarray}\displaystyle
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, \cdots) \quad ,
-\end{eqnarray}
-along with prior information about observation times, $\{t_1,...,t_N\}$ and other prior assumptions \joey{we should be more explicit here}, which we make explicit by creating the prior information set
-\begin{eqnarray}\displaystyle
-I &\equiv& (\{t_i\}_{i=1}^N, \mbox{assumptions}) \quad .
-\end{eqnarray}
-Our goal is to write down the form of the likelihood of the data, $D$, given $\theta$ and $I$.
-Then, for each light curve we can search for the vector $\theta$ that maximizes the data likelihood, and use those maximum likelihood estimates for downstream astrophysical inference.
-
-\subsection{Statistical Model for Light Curves with Non-Detections}
-
-We model the observed flux, $\fobs$, as a Gaussian distribution with variance that has both measurement ($\sigma^2_i$) and model ($s^2_{\mu}$) contributions.
-In the case that a detection is made, we require that the observed brightness be greater than the brightness of the detection limit ($\fobs = \ftrue \ge b_i$) while in the case that no detection is made, we require that the brightness (if it could be measured) be less than the detection limit ($\ftrue < b_i$).
-To derive the likelihood of the observed $\fobs$ given $\sigma^2_i$, $\theta$ and $I$, we must integrate over the prior distribution of the unknown $b_i$,
+Using the model above, the likelihood of the observed $\fobs$ given $\sigma^2_i$, $\theta$ and $I$ (after integrating out the unknown $b_i$) is
 \begin{eqnarray}\displaystyle
 p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
-  N(\ftrue | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_{-\infty}^{\ftrue} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
+  N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_{-\infty}^{\fobs} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
   \int_{-\infty}^{\infty} \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
@@
 where we have introduced the hyperparameters $B$ and $V_B$ for the Gaussian prior distribution of $b_i$.
-In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ from $0$ (low brightness limit) to $\ftrue$, enforcing that the detection limit be fainter than the observed brightness.
-In the epochs for which no detection was made ($\fobs = \texttt{NA}$), we compute the probability that the unknown detection threshold $b_i$ is greater than the unknown true flux $f_i$.
-When both $b_i$ and $f_i$ are normally distributed, this can be computed using a single call to a Gaussian cumulative distribution function, specifically
-\begin{equation*}
-\mathbb{P}(b_i \geq f_i) = \int_{-\infty}^0 N(x|B-\mu_i,V_B + \sigma_i^2 + s_{\mu}^2) dx
-\end{equation*}
+In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ up to the observed flux $\fobs$, enforcing that the detection limit be fainter than the observed brightness.
+Computationally, this requires a call to a Gaussian pdf, a call to a Gaussian cdf, and a multiplication.
+In the epochs for which no detection was made ($\fobs = \texttt{NA}$), we compute the probability that the unknown detection threshold $b_i$ is greater than the unknown true flux $f_i^*$.
+When both $b_i$ and $f_i^*$ are normally distributed, this can be computed using a single call to a Gaussian cumulative distribution function.
+%% \begin{equation*}
+%% \mathbb{P}(b_i \geq f_i) = \int_{-\infty}^0 N(x|\mu_i - B,V_B + \sigma_i^2 + s_{\mu}^2) dx
+%% \end{equation*}
 
-In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made on that epoch.
+In the above, we have assumed no extra information on each of the $b_i$ values besides the knowledge of whether a detection was made at that epoch.
 Hence, we draw, in Equation \ref{eq:bprior}, each $b_i$ value from a global prior distribution which is the same at all epochs.
@@
 
-Next, we model the reported variance, $\sobs$, on the uncertainty of the magnitude measurement.
-Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true uncertainty variance, $\sigma^2_i$ through a Gamma likelihood, which is defined to take on non-negative values.
-Our likelihood of observed $\sobs$ given $\fobs$, $\sigma^2_i$, $\theta$ and $I$, is
+\subsection{Modeling Uncertainty in Uncertainties}
+\label{sec:uncertainty}
+Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true photon count uncertainty variance $\sigma^2_i$ through a Gamma likelihood.
+Our probability of observed $\sobs$ given $\fobs$, $\sigma^2_i$, $\theta$ and $I$, is
 \begin{eqnarray}\displaystyle
@@
   \Gamma (\sobs | \sigma^2_i, V_{\sigma} ) & \mbox{if $\fobs \ne \texttt{NA}$} \\
- 1& \mbox{if $\fobs = \texttt{NA}$}
+ I(\sobs = \texttt{NA})& \mbox{if $\fobs = \texttt{NA}$}
 \end{array}\right.\label{eq:slik}
@@
 \end{eqnarray}
-where $\Gamma(x | m, V)$ is the standard Gamma distribution with mean $m$ and variance $V$,
-and we have added a model parameter $V_\sigma$ to represent the
+where $\Gamma(x | m, V)$ is the standard Gamma distribution parameterized by its mean $m$ and variance $V$.
+The  model parameter $V_\sigma$ represents the
 variance in the distribution of reported uncertainties given the true
 uncertainty.  (Typically, the Gamma distribution is parameterized by parameters $(\alpha, \beta)$, where the mean is $\alpha\beta$ and the variance is $\alpha\beta^2$.)  The  purpose of
- including the likelihood in Equation \ref{eq:slik} to the model, in practice, to keep the $\sigma^2_i$ from
+ including the likelihood in Equation \ref{eq:slik} to the model, in practice, is to keep the $\sigma^2_i$ from
 drifting very far away from the $\sobs$, as set by the hyperparameter
-$V_\sigma$.
+$V_\sigma$. \james{maybe clarify this last sentence. isn't the main reason for having the $V_\sigma$ term because we don't believe the reported variance}
 
+We place a Gamma prior on the true photon count uncertainties, adding two parameters to our model.
+\begin{eqnarray}\displaystyle
+p(\sigma^2_i|\theta, I) &=& \Gamma(\sigma^2_i | S, V_S)
+\label{eq:sigma_prior}
+\end{eqnarray}
+\james{should probably justify choice of some priors here / include some plots and/or refs. gamma priors and normal priors.}
+
+\subsection{The Full Model}
+\label{sec:full}
 Putting it all together, we can write down the likelihood of the data, $D_i$, for the parameters $\theta$, on a single epoch, as
@@
 \\
-p(\sigma^2_i|\theta, I) &=& \Gamma(\sigma^2_i | S, V_S)
-\\
 \theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, V_{\sigma}, S, V_S) \quad ,
 \end{eqnarray}
-where we have inserted the expressions from Equations (\ref{eq:mlik}) and (\ref{eq:slik}), we have integrated out the nuisance parameter $\sigma^2_i$, again the $\Gamma$ is the Gamma function with specified mean and variance, and we have two more hyperparameters $S$ and $V_S$.
+where we have inserted the expressions from Equations (\ref{eq:mlik}), (\ref{eq:slik}), and (\ref{eq:sigma_prior}), and integrated over the nuisance parameter $\sigma^2_i$.
 
@@
 data set (all the measurements and non-detections of this star from
-all the epochs, as delivered by the survey) given the $2K + 5$ parameter vector $\theta$.  This model
+all the epochs, as delivered by the survey) given the $2K + 5$ parameter vector $\theta$.
+Figure \ref{fig:graph} displays a graphical model highlighting dependencies and 
+independence assumptions. This model
 ``correctly'' or at least ``justifiably'' uses all of the information
@@
 
-\subsection{Implementation}
-\label{ss:implemenation}
+%% Hence, our initial model consists of the parameter vectors $\{f^*_1,...,f^*_N\}$, $\{b_1,...,b_N\}$ and $\{\sigma^2_1,...,\sigma^2_N\}$, and the model parameters
+%% \begin{eqnarray}\displaystyle
+%% \theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, \cdots) \quad ,
+%% \end{eqnarray}
+%% along with prior information about observation times, $\{t_1,...,t_N\}$ and other prior assumptions \joey{we should be more explicit here}, which we make explicit by creating the prior information set
+%% \begin{eqnarray}\displaystyle
+%% I &\equiv& (\{t_i\}_{i=1}^N, \mbox{assumptions}) \quad .
+%% \end{eqnarray}
+%% Our goal is to write down the form of the likelihood of the data, $D$, given $\theta$ and $I$.
+Then, for each light curve we can search for the vector $\theta$ that maximizes the data likelihood, and use those maximum likelihood estimates for downstream astrophysical inference.
 
-\joey{Describe Python implementation here}
+\subsection{Optimization in Practice}
+\label{ss:optim}
+
+\joey{Describe how we optimize}
 

### 2012-04-13 py/censoring2.py
can now read in miras from ASAS website

@@ -99,7 +99,7 @@ class Censored:
         par = np.zeros(10)
         par[0] = Period
         par[1:4] = self.least_sq(Period)
-        par[4] = (0.2 * par[1])**2
+        par[4] = (0.5 * par[1])**2
         par[5] = 2.*np.abs(np.min(self.f))
         par[6] = par[5]
         par[7] = np.median(self.ef2)
@@ -113,7 +113,6 @@ class Censored:
         Amat[1,:] = self.mu(self.t,2.*np.pi / Period, 0,1,0)
         Amat[2,:] = self.mu(self.t,2.*np.pi / Period, 0,0,1)
         Atb = np.dot(Amat, self.f)
-        print Atb.shape
         AtAinv = np.matrix(np.dot(Amat,Amat.T)).I
         return np.dot(AtAinv,Atb)
         

### 2012-04-13 py/test_censoring2.py

@@ -128,7 +128,7 @@ if __name__ == "__main__":
         #p0 = popt
         #p0 = popt + stats.norm.rvs(0,.01,size=10)
         print p0
-        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=0.5,xtol=0.1,mfev=500)
+        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=500)
         print cmodel.log_likelihood(pfmin)
         print cmodel.log_likelihood(p0)
 

### 2012-04-13 py/fit_censoring.py
file that crunches on miras is now committed

@@ -1,0 +1,137 @@
+
+
+import censoring2 as c2
+import cProfile
+import numpy as np
+import matplotlib.pylab as plt
+import sys, os
+import urllib2 as ulib
+
+sys.path.append(os.path.abspath(os.environ.get("TCP_DIR") + \
+                                      '/Algorithms/fitcurve'))
+from lomb_scargle_refine import lomb as lombr
+from lomb_scargle_censor import lomb as lombc
+
+path = '/Users/jwrichar/Documents/CDI/CensoredData/'
+
+def periods_lomb(t, f, e, n_per=3,f0=1./1000,df=1e-5,fn=5.):
+
+    numf = int((fn-f0)/df)
+    ftest = 1.*f
+    P_cand = []
+    for i in xrange(n_per):
+        # lomb-scargle periodogram
+        psd,res = lombr(t,ftest,e,f0,df,numf, detrend_order=1,nharm=1)
+        P_cand.append(1./res['freq'])
+        ftest -= res['model']
+
+    return P_cand
+
+
+def period_lombc(t, f, e, tc,f0=1./1000,df=1e-5,fn=5., n_per = 5):
+
+    numf = int((fn-f0)/df)
+    freqgrid = np.linspace(f0,fn,numf)
+    ftest = 1.*f
+    # lomb-scargle periodogram
+    psd = lombc(t,ftest,e,tc,f0,df,numf)
+
+    plt.plot(1./freqgrid,psd)
+    indpeak = find_peaks(psd) # find peaks in LS
+    indtop = np.where(psd[indpeak] >= np.sort(psd[indpeak])[-n_per])[0] 
+    
+    P_cand = (1./freqgrid[indpeak[indtop]]) # top n_per periods
+    return P_cand
+
+def find_peaks(x):
+    xmid = x[1:-1] # orig array with ends removed
+    xm1 = x[2:] # orig array shifted one up
+    xp1 = x[:-2] # origh array shifted one back
+    return np.where(np.logical_and(xmid > xm1, xmid > xp1))[0] + 1
+
+def load_lc_from_web(ID):
+    urlpath = 'http://www.astrouw.edu.pl/cgi-asas/asas_cgi_get_data?'
+    ur = ulib.urlopen( urlpath + ID + ',asas3')#open url
+    miradata = np.loadtxt(ur.readlines(), usecols=(0,3,8,11), \
+                          dtype=[('t',np.float64), ('m',np.float64), ('e',np.float64),\
[…]

### 2012-04-13 py/censoring2.py
*bangs head against wall*

@@ -121,11 +121,13 @@ class Censored:
         #opt = op.fmin_bfgs(self.negll, p0, gtol=ftol, maxiter=maxiter)
         return opt
 
-    def plot(self, ax, par, fold = False):
+    def plot(self, ax, par, fold = False, plot_model = True):
         '''
         input:
         - ax: matplotlib axes object
         - par: set of hyperparameters
+        - fold: if True plot folded light curve, else unfolded
+        - plot_model: if True plot light curve model
 
         usage:
             plt.clf()
@@ -165,15 +167,18 @@ class Censored:
             ax.plot(x - mediant + 1., self.f, 'ko', alpha=0.5, mec='k')
             hogg_errorbar(ax, x - mediant + 1., self.f, self.ef)
             ax.plot(xc - mediant + 1., np.zeros_like(xc), 'r.', alpha=0.5, mec='r')
-        
-        mup = self.mu(tp, omega, A0, A1, B1)
-        ax.plot(xp - mediant, mup + np.sqrt(s2mu), 'b-', alpha=0.25)
-        ax.plot(xp - mediant, mup,                 'b-', alpha=0.50)
-        ax.plot(xp - mediant, mup - np.sqrt(s2mu), 'b-', alpha=0.25)
+        if(plot_model):
+            mup = self.mu(tp, omega, A0, A1, B1)
+            ax.plot(xp - mediant, mup + np.sqrt(s2mu), 'b-', alpha=0.25)
+            ax.plot(xp - mediant, mup,                 'b-', alpha=0.50)
+            ax.plot(xp - mediant, mup - np.sqrt(s2mu), 'b-', alpha=0.25)
         ax.set_xlim(tlim - mediant)
         foo = np.max(self.f + self.ef)
         ax.set_ylim(-0.1 * foo, 1.1 * foo)
-        ax.set_xlabel(r'time $t$ (MJD - %d~d)' % mediant)
+        if(fold):
+            ax.set_xlabel(r'$phi$')
+        else:
+            ax.set_xlabel(r'time $t$ (MJD - %d~d)' % mediant)
         ax.set_ylabel(r'flux $f$ ($\mu$Mgy)')
         ax.set_title(self.name)
         return None

### 2012-04-13 py/fit_censoring.py

@@ -36,6 +36,9 @@ def period_lombc(t, f, e, tc,f0=1./1000,df=1e-5,fn=5., n_per = 5):
     # lomb-scargle periodogram
     psd = lombc(t,ftest,e,tc,f0,df,numf)
 
+    #plt.plot(freqgrid,psd)
+    #plt.show()
+    
     plt.plot(1./freqgrid,psd)
     indpeak = find_peaks(psd) # find peaks in LS
     indtop = np.where(psd[indpeak] >= np.sort(psd[indpeak])[-n_per])[0] 
@@ -55,7 +58,8 @@ def load_lc_from_web(ID):
     miradata = np.loadtxt(ur.readlines(), usecols=(0,3,8,11), \
                           dtype=[('t',np.float64), ('m',np.float64), ('e',np.float64),\
                                  ('grade','S2')])
-    use = np.where(np.logical_and(miradata['grade'] != 'D',miradata['grade'] != 'F'))[0]
+    use = np.where(np.logical_and(np.logical_and(miradata['grade'] != 'D',miradata['grade'] != 'F'),\
+                                  miradata['m'] != 99.999))[0]
     return miradata[use,:]
 
 
@@ -70,62 +74,65 @@ if __name__ == "__main__":
 
     # select all miras (P(Mira) > 0.75 and anom_score < 3.0)
     miras = np.where(np.logical_and(catalog['Pmira'] > 0.75 , catalog['anom'] < 3.))[0] # 2538 mira variables
-
-    miras_LS_P = catalog['P'][miras]
-
+    
 
     #### FOR EACH MIRA IN miras ####
-    data = load_lc_from_web(catalog['ID'][0])
-
-    #'http://www.astrouw.edu.pl/cgi-asas/asas_cgi_get_data?102614-3830.6,asas3'
-
-    # load in data
-    #path = '/Users/jwrichar/Documents/CDI/CensoredData/'
-    #data = np.loadtxt(path + "data/lc_102614-3830.6.dat",\
-    #                     dtype=[('t', np.float32), ('m', np.float32),\
-    #                            ('e', np.float32)])
-
-    indo = np.where(data['m'] != 29.999)
-
-    tobs = data['t'][indo]
-    mobs = data['m'][indo]
-    eobs = data['e'][indo]
-    tcen = data['t'][np.where(data['m'] == 29.999)]
-
-    # instantiate Censor object
-    cmodel = c2.Censored(tobs, mobs, eobs, tcen, name = catalog['ID'][0])
-
[…]

### 2012-04-13 py/censoring2.py
changed sinusoid fitting to mag space from flux space

@@ -67,8 +67,8 @@ class Censored:
 
     ## convert all times to expected flux (i.e. t_i -> u_i)
         w =  (2 * np.pi)/ P 
-        self.u  = self.mu(self.t,  w, A0, A1, B1)
-        self.uc = self.mu(self.tc, w, A0, A1, B1)
+        self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1))
+        self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1))
 
     ## compute loglikelihood
         return np.sum(self.loglikelihood_censored(su2,B,VB,S,VS)) + \
@@ -99,7 +99,7 @@ class Censored:
         par = np.zeros(10)
         par[0] = Period
         par[1:4] = self.least_sq(Period)
-        par[4] = (0.5 * par[1])**2
+        par[4] = (0.4 * mag2flux(par[1]))**2
         par[5] = 2.*np.abs(np.min(self.f))
         par[6] = par[5]
         par[7] = np.median(self.ef2)
@@ -112,7 +112,7 @@ class Censored:
         Amat[0,:] = 1.
         Amat[1,:] = self.mu(self.t,2.*np.pi / Period, 0,1,0)
         Amat[2,:] = self.mu(self.t,2.*np.pi / Period, 0,0,1)
-        Atb = np.dot(Amat, self.f)
+        Atb = np.dot(Amat, self.m) # JWR changed to fitting in mag space
         AtAinv = np.matrix(np.dot(Amat,Amat.T)).I
         return np.dot(AtAinv,Atb)
         
@@ -168,7 +168,7 @@ class Censored:
             hogg_errorbar(ax, x - mediant + 1., self.f, self.ef)
             ax.plot(xc - mediant + 1., np.zeros_like(xc), 'r.', alpha=0.5, mec='r')
         if(plot_model):
-            mup = self.mu(tp, omega, A0, A1, B1)
+            mup = mag2flux(self.mu(tp, omega, A0, A1, B1))
             ax.plot(xp - mediant, mup + np.sqrt(s2mu), 'b-', alpha=0.25)
             ax.plot(xp - mediant, mup,                 'b-', alpha=0.50)
             ax.plot(xp - mediant, mup - np.sqrt(s2mu), 'b-', alpha=0.25)

### 2012-04-13 py/fit_censoring.py

@@ -77,7 +77,7 @@ if __name__ == "__main__":
     
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(1,100):
+    for jj in np.arange(100):
         # load in data from web
         print 'doing mira ' + str(catalog['ID'][miras[jj]]) + ' dotAstro: ' + str(catalog['dID'][miras[jj]])
     
@@ -93,14 +93,16 @@ if __name__ == "__main__":
         cmodel = c2.Censored(tobs, mobs, eobs, tcen, name = catalog['ID'][miras[jj]])
 
         # do Lomb-scargle search to get top few periods
-        nper = 3 # number of periods to initialize over
+        nper = 1 # number of periods to initialize over
         Pinit = period_lombc(tobs, mobs, eobs, tcen, n_per = nper,df=1e-5)
         #    Ps = periods_lomb(tobs, mobs, eobs)
         print Pinit
 
         p0 = cmodel.get_init_par(Pinit[0])
-        if np.isnan(cmodel.log_likelihood(p0)):
+        ll0 = cmodel.log_likelihood(p0)
+        if np.isnan(ll0):
             continue
+        print 'Log-likelihood at p0: ' + str(ll0)
 
         lliks = []
         params = np.zeros((nper,10))

### 2012-04-13 py/censoring2.py
going from smu2 to eta2

@@ -58,7 +58,7 @@ class Censored:
         A0 = par[1]
         A1 = par[2]
         B1 = par[3]
-        su2 = par[4]
+        eta2 = par[4]
         B = par[5]
         VB = par[6]
         Vsig = par[7]
@@ -71,35 +71,35 @@ class Censored:
         self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1))
 
     ## compute loglikelihood
-        return np.sum(self.loglikelihood_censored(su2,B,VB,S,VS)) + \
-               np.sum(self.loglikelihood_observed(su2,B,VB,Vsig,S,VS))
+        return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
+               np.sum(self.loglikelihood_observed(eta2,B,VB,Vsig,S,VS))
     
-    def loglikelihood_censored(self,su2,B,VB,S,VS):
+    def loglikelihood_censored(self,eta2,B,VB,S,VS):
     ## integrand of equation (12), integrate this across sig2
-        def sig_integrand(sig2,ui,su2,B,VB,S,VS):
-            return  gaussian_cdf(B, ui, VB + sig2 + su2) * gamma_pdf(sig2,S,VS)
-        return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0] \
+        def sig_integrand(sig2,ui,eta2,B,VB,S,VS):
+            return  gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui) * gamma_pdf(sig2,S,VS)
+        return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,eta2,B,VB,S,VS),epsabs=self.tolquad)[0] \
                        for ui in self.uc])
 
-    def loglikelihood_observed(self,su2,B,VB,Vsig,S,VS):
-        def integrand(sig2,ui,fi,ei2,su2,B,VB,Vsig,S,VS):
+    def loglikelihood_observed(self,eta2,B,VB,Vsig,S,VS):
+        def integrand(sig2,ui,fi,ei2,eta2,B,VB,Vsig,S,VS):
             p_not_cens = gaussian_cdf(fi, B, VB)
-            p_flux = gaussian_pdf(fi, ui, sig2 + su2)
+            p_flux = gaussian_pdf(fi, ui, sig2 + eta2 * ui * ui)
[…]

### 2012-04-13 tex/ms.tex
going from smu2 to eta2

@@
 
-In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{model variance} $s_\mu^2$ at each point (assumed constant but easily generalized) as well as random variation due to photo arrival counts, assumed normal mean 0 with variance $\sigma_i^2$.
+In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{fractional model variance} $\eta^2$ at each point (assumed constant but easily generalized) as well as random variation due to photo arrival counts, assumed normal mean 0 with variance $\sigma_i^2$.
 The reported uncertainties $s_i^2$ are an estimate of the $\sigma_i^2$, see Subsection \ref{sec:uncertainty}.
@@
 \end{eqnarray}
-where $e_i \sim N(0,s_\mu^2 + \sigma_i^2)$.
+where $e_i \sim N(0,\sigma_i^2 + \eta^2\,\mu_i^2$.
 Recall that this is a latent variable because we will not see it when an observation is censored.
@@
 p(\fobs |\sigma^2_i,\theta,I) &=& \left\{\begin{array}{ll}
-  N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  \int_{-\infty}^{\fobs} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
-  \int_{-\infty}^{\infty} \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
+  N(\fobs | \mu_i,  \sigma^2_i + \eta^2\,\mu_i^2)\,  \int_{-\infty}^{\fobs} p(b_i | \theta)\, \dd b_i & \mbox{if $\fobs \ne \texttt{NA}$} \\
+  \int_{-\infty}^{\infty} \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + \eta^2\,\mu_i^2)\, p(b_i | \theta)\, \dd \ftrue\, \dd b_i & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik}
@@
 \\
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, \cdots) \quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2, B, V_B, \cdots) \quad ,
 \end{eqnarray}
@@
 When both $b_i$ and $f_i^*$ are normally distributed, this can be computed using a single call to a Gaussian cumulative distribution function.
-%% \begin{equation*}
-%% \mathbb{P}(b_i \geq f_i) = \int_{-\infty}^0 N(x|\mu_i - B,V_B + \sigma_i^2 + s_{\mu}^2) dx
-%% \end{equation*}
 
@@
 p(\fobs |\sigma^2_i,\theta,I, b_i) &=& \left\{\begin{array}{ll}
-  N(\fobs | \mu_i,  \sigma^2_i + s^2_{\mu})\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
- \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + s^2_{\mu})\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
+  N(\fobs | \mu_i,  \sigma^2_i + \eta^2\,\mu_i^2)\,  I(\fobs \ge b_i) & \mbox{if $\fobs \ne \texttt{NA}$} \\
+ \int_{-\infty}^{b_i} N(\ftrue | \mu_i, \sigma^2_i + \eta^2\,\mu_i^2)\, \dd \ftrue & \mbox{if $\fobs = \texttt{NA}$} \\
 \end{array}\right.\label{eq:mlik_s}
@@
 \\
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, B, V_B, V_{\sigma}, S, V_S) \quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2\,\mu_i^2, B, V_B, V_{\sigma}, S, V_S) \quad ,
 \end{eqnarray}
@@
 \end{figure}
-
-%% Hence, our initial model consists of the parameter vectors $\{f^*_1,...,f^*_N\}$, $\{b_1,...,b_N\}$ and $\{\sigma^2_1,...,\sigma^2_N\}$, and the model parameters
-%% \begin{eqnarray}\displaystyle
-%% \theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, s^2_\mu, \cdots) \quad ,
-%% \end{eqnarray}
-%% along with prior information about observation times, $\{t_1,...,t_N\}$ and other prior assumptions \joey{we should be more explicit here}, which we make explicit by creating the prior information set
-%% \begin{eqnarray}\displaystyle
-%% I &\equiv& (\{t_i\}_{i=1}^N, \mbox{assumptions}) \quad .
-%% \end{eqnarray}
-%% Our goal is to write down the form of the likelihood of the data, $D$, given $\theta$ and $I$.
-Then, for each light curve we can search for the vector $\theta$ that maximizes the data likelihood, and use those maximum likelihood estimates for downstream astrophysical inference.
 

### 2012-04-13 py/censoring2.py
added fast ll computation for observed data

@@ -7,8 +7,8 @@ This code is synchronized with the document tex/ms.tex
 Issues:
 - limits on gamma integrations are made up (sensitivity not checked)
 - make it possible to insert estimated (or known) upper limits
-- need a folded version of the plot
 - need a magnitude version of the plot
+- quad returns nan when VB and VS get very large; consider rescaling fluxes to order unity
 """
 
 import numpy as np
@@ -48,7 +48,7 @@ class Censored:
     def mu(self, times, omega, A0, A1, B1):
         return A0 + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
         
-    def log_likelihood(self,par):
+    def log_likelihood(self,par, fast = True):
         """ 
         computes log p(D|theta,I)
         everything else: same notation as paper (all floats)
@@ -70,16 +70,33 @@ class Censored:
         self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1))
         self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1))
 
+        if(fast):
+            return np.sum(self.loglikelihood_censored(su2,B,VB,S,VS)) + \
+                   np.sum(self.loglikelihood_observed_fast(su2,B,VB,Vsig))
     ## compute loglikelihood
         return np.sum(self.loglikelihood_censored(su2,B,VB,S,VS)) + \
                np.sum(self.loglikelihood_observed(su2,B,VB,Vsig,S,VS))
-    
+
     def loglikelihood_censored(self,su2,B,VB,S,VS):
     ## integrand of equation (12), integrate this across sig2
         def sig_integrand(sig2,ui,su2,B,VB,S,VS):
             return  gaussian_cdf(B, ui, VB + sig2 + su2) * gamma_pdf(sig2,S,VS)
-        return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0] \
-                       for ui in self.uc])
+        sumll = 0.
+        for ui in self.uc:
+            tmp = integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,su2,B,VB,S,VS),epsabs=self.tolquad)[0]
+            #print tmp
[…]

### 2012-04-13 py/fit_censoring.py

@@ -7,6 +7,8 @@ import matplotlib.pylab as plt
 import sys, os
 import urllib2 as ulib
 
+import pdb
+
 sys.path.append(os.path.abspath(os.environ.get("TCP_DIR") + \
                                       '/Algorithms/fitcurve'))
 from lomb_scargle_refine import lomb as lombr
@@ -63,6 +65,12 @@ def load_lc_from_web(ID):
     return miradata[use,:]
 
 
+def catchnan(p):
+    ll0 = cmodel.log_likelihood(p)
+    if(np.isnan(ll0)):
+        1./0
+    pass
+
 if __name__ == "__main__":
 
     # load catalog
@@ -77,7 +85,7 @@ if __name__ == "__main__":
     
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(100):
+    for jj in np.arange(1,100):
         # load in data from web
         print 'doing mira ' + str(catalog['ID'][miras[jj]]) + ' dotAstro: ' + str(catalog['dID'][miras[jj]])
     
@@ -90,7 +98,7 @@ if __name__ == "__main__":
         tcen = data['t'][np.where(data['m'] == 29.999)]
 
         # instantiate Censor object
-        cmodel = c2.Censored(tobs, mobs, eobs, tcen, name = catalog['ID'][miras[jj]])
+        cmodel = c2.Censored(tobs, mobs, eobs, tcen, name = catalog['ID'][miras[jj]],tolquad=1.)
 
         # do Lomb-scargle search to get top few periods
         nper = 1 # number of periods to initialize over
@@ -100,9 +108,14 @@ if __name__ == "__main__":
 
         p0 = cmodel.get_init_par(Pinit[0])
         ll0 = cmodel.log_likelihood(p0)
-        if np.isnan(ll0):
-            continue
+
+        #pdb.run('catchnan(p0)')
+
         print 'Log-likelihood at p0: ' + str(ll0)
+        print p0
+        if np.isnan(ll0):
+            raise Exception
+            #continue
 
         lliks = []
         params = np.zeros((nper,10))
@@ -111,8 +124,8 @@ if __name__ == "__main__":
             # params:     P,A0,A1,B1,su2,B,VB,Vsig,S,VS
             p0 = cmodel.get_init_par(Pinit[i])
             print p0
-            pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=200)
[…]

### 2012-04-14 tex/ms.tex

@@
 
-In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{fractional model variance} $\eta^2$ at each point (assumed constant but easily generalized) as well as random variation due to photo arrival counts, assumed normal mean 0 with variance $\sigma_i^2$.
+In addition to the expected brightness in Equation \ref{eq:fourier}, we assume that there is uncertainty or inappropriateness in the model, leading to a \emph{fractional model variance} $\eta^2$ at each point (assumed constant but easily generalized) as well as random variation due to photon arrival counts, assumed normal mean 0 with variance $\sigma_i^2$.
 The reported uncertainties $s_i^2$ are an estimate of the $\sigma_i^2$, see Subsection \ref{sec:uncertainty}.
@@
 \end{eqnarray}
-where $e_i \sim N(0,\sigma_i^2 + \eta^2\,\mu_i^2$.
-Recall that this is a latent variable because we will not see it when an observation is censored.
+where $e_i \sim N(0,\sigma_i^2 + \eta^2\,\mu_i^2)$.
+Recall that $f_i^*$ is a latent variable because we will not see it when an observation is censored.
 We instantiate a latent variable, $b_i$, to represent the detection limit, in units of flux, at epoch $i$.
@@
 data set (all the measurements and non-detections of this star from
-all the epochs, as delivered by the survey) given the $2K + 5$ parameter vector $\theta$.
+all the epochs, as delivered by the survey) given the $2K + 9$ parameter vector $\theta$.
 Figure \ref{fig:graph} displays a graphical model highlighting dependencies and 

### 2012-04-14 py/censoring2.py
small bug fix

@@ -196,7 +196,7 @@ class Censored:
             if(mag):
                 mup = self.mu(tp, omega, A0, A1, B1)
                 mup_p = mup + 1.086 * np.sqrt(eta2)
-                mup_m = mup - 1.086 * np.sqrt(eta)
+                mup_m = mup - 1.086 * np.sqrt(eta2)
             else:
                 mup = mag2flux(self.mu(tp, omega, A0, A1, B1))
                 mup_p = mup * (1. + np.sqrt(eta2))

### 2012-04-18 tex/ms.tex

@@
 \begin{eqnarray}\displaystyle
-\mu_i &=& A_0 + \sum_{k=1}^K A_k\, \sin (t_i \, \omega  k) + B_k\, \cos (t_i \, \omega  k)
+\mu_i' &=& A_0 + \sum_{k=1}^K A_k\, \sin (t_i \, \omega  k) + B_k\, \cos (t_i \, \omega  k)
 \quad ,\label{eq:fourier}
 \end{eqnarray}
-where $\mu_i$ is the flux at time $t_i$, $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(\frac{B_k}{A_k})$ is the relative phase offset of harmonic $k$.
+where $\mu_i'$ is the model brightness estimate in magnitude at time $t_i$, $\sqrt{A_k^2 + B_k^2}$ is the amplitude of the $k^{\rm th}$ harmonic of the frequency $\omega$ and $\tan^{-1}(\frac{B_k}{A_k})$ is the relative phase offset of harmonic $k$.
 The number of harmonics, $K$, can either be fixed or treated as a free parameter over which to optimize the likelihood.
+
+We model the brightness in magnitude because periodic variation in magnitude is more sinusoidal than in flux.
+However since errors are more Gaussian in flux space, we convert the $u_i'$ to flux measurements using
+\begin{equation*}
+\mu_i = 10^6 \times 10^{-.4 \mu_i'}
+\end{equation*}
+Now both the model predictions $\mu_i$ and the data $(f_i,s_i)$ are in flux.
 

### 2012-04-23 py/censoring2.py
fixed buggy mag plot

@@ -48,7 +48,7 @@ class Censored:
     def mu(self, times, omega, A0, A1, B1):
         return A0 + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
         
-    def log_likelihood(self,par, fast = True):
+    def log_likelihood(self, par, fast = True):
         """ 
         computes log p(D|theta,I)
         everything else: same notation as paper (all floats)
@@ -107,8 +107,8 @@ class Censored:
                        for (ui,fi,ei2) in zip(self.u,self.f,self.ef2)])
 
 
-    def negll(self,par):
-        return -1*self.log_likelihood(par)
+    def negll(self, par, fast=True):
+        return -1*self.log_likelihood(par, fast=fast)
 
     def get_init_par(self,Period):
         # params:     P,A0,A1,B1,eta2,B,VB,Vsig,S,VS
@@ -133,8 +133,8 @@ class Censored:
         AtAinv = np.matrix(np.dot(Amat,Amat.T)).I
         return np.dot(AtAinv,Atb)
         
-    def optim_fmin(self,p0,maxiter=1000,ftol=0.0001,xtol=0.0001,mfev=1.e8):
-        opt = op.fmin(self.negll, p0, maxiter=maxiter,ftol=ftol,maxfun=mfev)
+    def optim_fmin(self, p0, maxiter=1000, ftol=0.0001, xtol=0.0001, mfev=1.e8, fast=True):
+        opt = op.fmin(self.negll, p0, args = (fast,), maxiter=maxiter,ftol=ftol,maxfun=mfev)
         #opt = op.fmin_bfgs(self.negll, p0, gtol=ftol, maxiter=maxiter)
         return opt
 
@@ -145,6 +145,7 @@ class Censored:
         - par: set of hyperparameters
         - fold: if True plot folded light curve, else unfolded
         - plot_model: if True plot light curve model
+        - mag: if True plot in mag space, else in flux space
 
         usage:
             plt.clf()
@@ -160,7 +161,7 @@ class Censored:
             return (t/P) % 1.
         if(fold):
             x, xc = phase(self.t, par[0]), phase(self.tc, par[0])
-            mediant = 0
+            mediant = 0.
             tlim = np.array([0., 2.])
             tp = np.linspace(0., par[0]*2., 10000)
             xp = np.linspace(0., 2., len(tp))
[…]

### 2012-04-23 py/fit_censoring.py

@@ -1,4 +1,11 @@
 
+if __name__ == '__main__':
+    import matplotlib
+    matplotlib.use('Agg')
+    from matplotlib import rc
+    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
+    rc('text', usetex=True)
+
 
 import censoring2 as c2
 import cProfile
@@ -11,6 +18,7 @@ import pdb
 
 sys.path.append(os.path.abspath(os.environ.get("TCP_DIR") + \
                                       '/Algorithms/fitcurve'))
+from lomb_scargle import lomb as lomb
 from lomb_scargle_refine import lomb as lombr
 from lomb_scargle_censor import lomb as lombc
 
@@ -19,14 +27,13 @@ path = '/Users/jwrichar/Documents/CDI/CensoredData/'
 def periods_lomb(t, f, e, n_per=3,f0=1./1000,df=1e-5,fn=5.):
 
     numf = int((fn-f0)/df)
+    #freqgrid = np.linspace(f0,fn,numf)
     ftest = 1.*f
     P_cand = []
     for i in xrange(n_per):
-        # lomb-scargle periodogram
-        psd,res = lombr(t,ftest,e,f0,df,numf, detrend_order=1,nharm=1)
+        psd,res = lombr(t,ftest,e,f0,df,numf, detrend_order=1,nharm=8)
         P_cand.append(1./res['freq'])
         ftest -= res['model']
-
     return P_cand
 
 
@@ -38,10 +45,6 @@ def period_lombc(t, f, e, tc,f0=1./1000,df=1e-5,fn=5., n_per = 5):
     # lomb-scargle periodogram
     psd = lombc(t,ftest,e,tc,f0,df,numf)
 
-    #plt.plot(freqgrid,psd)
-    #plt.show()
-    
-    plt.plot(1./freqgrid,psd)
     indpeak = find_peaks(psd) # find peaks in LS
     indtop = np.where(psd[indpeak] >= np.sort(psd[indpeak])[-n_per])[0] 
     
@@ -73,6 +76,8 @@ def catchnan(p):
 
 if __name__ == "__main__":
 
+    print 'Fitting Miras'
+    
     # load catalog
     catalog = np.loadtxt('/Users/jwrichar/Documents/CDI/ASAS/ASASCatalog/catalog/asas_class_catalog_v2_3.dat',\
                          usecols=(0,1,4,5,6,9,37), skiprows=1,\
[…]

### 2012-04-23 py/smoothers.py

@@ -1,66 +1,0 @@
-import numpy as np
-from scipy import interpolate, randn
-import scipy.interpolate
-import matplotlib.pyplot as plt
-import scipy.stats
-import synthetic_data
-## smooth a curve
-import supsmu
-
-
-## TODO standardize to spline and supersmooth output same type of data
-## should both these functions be moved to visualize?
-
-
-## p246 in scipy tutorial for good
-## info on this
-
-## recommended for smoothing s=m-sqrt(2*m)
-def spline(tfe,period,class_name="Unknown"):
-    tfe[:,0] = tfe[:,0] % period
-    positions = tfe[:,0].argsort()
-    tfe = tfe[positions,:]
-    ##smoothing = (tfe[:,0].size + np.sqrt(2*tfe[:,0].size))
-    smoothing = pow(tfe[:,0].size,3/2)
-    tck = interpolate.splrep(tfe[:,0],tfe[:,1],w=(1/tfe[:,2]),
-                             s=smoothing,k=3,task=0)
-    xs = np.linspace(0,period,1000)
-    ys = interpolate.splev(xs,tck,der=0)
-    plt.figure()
-    plt.plot(xs,ys,tfe[:,0],tfe[:,1],'x')
-    plt.show()
-
-
-## TODO: allow arguments to adjust weights, span, and alpha
-def supersmooth(tfe,period):
-    tfe[:,0] = (tfe[:,0] % period) / period
-    positions = tfe[:,0].argsort()
-    tfe[:,:] = tfe[positions,:]
-    typevec = np.float32
-    x = tfe[:,0].astype(typevec)
-    y = tfe[:,1].astype(typevec)
-    w = 1 / (tfe[:,2].astype(typevec))
-    iper = np.array([2]).astype(np.int32)
-    span = np.array([0.0]).astype(typevec)
-    alpha = np.array([1.0]).astype(typevec)
-    smo = np.zeros(tfe.shape[0]).astype(typevec)
-    sc = np.zeros(tfe.shape[0]*7).reshape((tfe.shape[0],7)).astype(typevec)
-    supsmu.supsmu(x,y,w,iper,span,alpha,smo,sc)
-    return(smo)
-
-
-if __name__ == '__main__':
-    ## practice the supersmooth function
-    if 1:
-        aSurvey = synthetic_data.surveySetup()
-        aSurvey.generateCurve()
-        tfe = np.c_[aSurvey.times,aSurvey.fluxes,aSurvey.errors]
[…]

### 2012-04-23 py/test.py

@@ -1,24 +1,0 @@
-import numpy as np
-import scipy.optimize
-import scipy.stats
-
-
-
-x_rands = scipy.stats.norm().rvs(100)
-y_rands = 2.7*x_rands + 10 + scipy.stats.norm().rvs(100)
-
-
-
-# def squared_error(coeffs):
-#     return(sum(np.power(y_rands - x_rands*coeffs[1] - coeffs[0],2)))
-
-def squared_error(coeffs):
-    return(sum(np.power(y_rands - coeffs[0],2)))
-
-
-
-coeffs = [np.mean(y_rands),0]
-print coeffs
-coeffs = scipy.optimize.fmin(squared_error,coeffs)
-print coeffs
-print np.mean(y_rands)

### 2012-04-23 py/censoring2.py
changed lower limits of integration to avoid infs and catch inf in likelihood call

@@ -79,9 +79,11 @@ class Censored:
     def loglikelihood_censored(self,eta2,B,VB,S,VS):
     ## integrand of equation (12), integrate this across sig2
         def sig_integrand(sig2,ui,eta2,B,VB,S,VS):
-            return  gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui) * gamma_pdf(sig2,S,VS)
-        return np.log([integrate.quad(sig_integrand,0.,S+5*np.sqrt(VS),(ui,eta2,B,VB,S,VS),epsabs=self.tolquad)[0] \
-                       for ui in self.uc])
+            #print 'gauss: ' +  str(gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui))
+            #print gamma_pdf(sig2,S,VS)
+            return gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui) * gamma_pdf(sig2,S,VS)
+        return np.log([integrate.quad(sig_integrand,np.max([0.,S-5.*np.sqrt(VS)]),S+5.*np.sqrt(VS),(ui,eta2,B,VB,S,VS),\
+                                      epsabs=self.tolquad)[0] for ui in self.uc])
 
     
     def loglikelihood_observed_fast(self,eta2,B,VB,Vsig):
@@ -103,8 +105,8 @@ class Censored:
             p_si = gamma_pdf(ei2,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
-        return np.log([integrate.quad(integrand,0.,S+5.*np.sqrt(VS),(ui,fi,ei2,eta2,B,VB,Vsig,S,VS),epsabs=self.tolquad)[0]\
-                       for (ui,fi,ei2) in zip(self.u,self.f,self.ef2)])
+        return np.log([integrate.quad(integrand,np.max([0.,S-5.*np.sqrt(VS)]),S+5.*np.sqrt(VS),(ui,fi,ei2,eta2,B,VB,Vsig,S,VS),\
+                                      epsabs=self.tolquad)[0] for (ui,fi,ei2) in zip(self.u,self.f,self.ef2)])
 
 
     def negll(self, par, fast=True):
@@ -236,7 +238,10 @@ def gaussian_cdf(x, mean, var):
 def gamma_pdf(x,mean,var):
     theta = var / mean
[…]

### 2012-04-23 py/fit_censoring.py

@@ -87,14 +87,18 @@ if __name__ == "__main__":
 
     # select all miras (P(Mira) > 0.75 and anom_score < 3.0)
     miras = np.where(np.logical_and(catalog['Pmira'] > 0.75 , catalog['anom'] < 3.))[0] # 2538 mira variables
-    
+
+    new_periods = np.zeros((len(miras),3),dtype=[('ID','S10'), ('oldP',np.float), ('newP',np.float)])
 
     #### FOR EACH MIRA IN miras ####
     for jj in np.arange(0,100):
 #    for jj in np.arange(486,487):
         # load in data from web
         print 'doing mira ' + str(catalog['ID'][miras[jj]]) + ' dotAstro: ' + str(catalog['dID'][miras[jj]])
-    
+
+        new_periods[jj,0] = catalog['ID'][miras[jj]]
+        new_periods[jj,1] = catalog['P'][miras[jj]]
+        
         data = load_lc_from_web(catalog['ID'][miras[jj]])
         
         indo = np.where(data['m'] != 29.999)
@@ -102,16 +106,18 @@ if __name__ == "__main__":
         mobs = data['m'][indo]
         eobs = data['e'][indo]
         tcen = data['t'][np.where(data['m'] == 29.999)]
+        # print 'Censored observations: ' + str(tcen)
 
         # instantiate Censor object
         cmodel = c2.Censored(tobs, mobs, eobs, tcen, name = catalog['ID'][miras[jj]], tolquad=1.)
 
         # do Lomb-scargle search to get top few periods
-        nper = 1 # number of periods to initialize over
+        nper = 2. # number of periods to initialize over
         Pinit = period_lombc(tobs, mobs, eobs, tcen, n_per = nper,df=1e-5)
         #Porig = periods_lomb(tobs, mobs, eobs, n_per = nper)
-        print Pinit
+        #print Pinit
         #print Porig
+        
 
         p0 = cmodel.get_init_par(Pinit[0])
         ll0 = cmodel.log_likelihood(p0, fast=True)
@@ -125,22 +131,31 @@ if __name__ == "__main__":
             #continue
 
         lliks = []
-        params = np.zeros((nper,10))
+        params = np.zeros((2*nper,10))
         for i in np.arange(nper):
             # initial guess for model parameters
             # params:     P,A0,A1,B1,su2,B,VB,Vsig,S,VS
[…]

### 2012-04-23 py/fit_censoring.py
fit_censoring now searches over multiple periods

@@ -94,7 +94,7 @@ if __name__ == "__main__":
     for jj in np.arange(0,100):
 #    for jj in np.arange(486,487):
         # load in data from web
-        print 'doing mira ' + str(catalog['ID'][miras[jj]]) + ' dotAstro: ' + str(catalog['dID'][miras[jj]])
+        print 'doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) + ' dotAstro: ' + str(catalog['dID'][miras[jj]])
 
         new_periods[jj,0] = catalog['ID'][miras[jj]]
         new_periods[jj,1] = catalog['P'][miras[jj]]
@@ -140,13 +140,13 @@ if __name__ == "__main__":
             p0 = cmodel.get_init_par(Pinit[i])
             pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=500,fast=True)
             #print pfmin
-            params[(2*i - 1),:] = pfmin
+            params[(2*i),:] = pfmin
             lliks.append(cmodel.log_likelihood(pfmin))
 
             # initialized at twice Nat's period
             p0 = cmodel.get_init_par(2 * Pinit[i])
             pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=500,fast=True)
-            params[(2*i),:] = pfmin
+            params[((2*i)+1),:] = pfmin
             lliks.append(cmodel.log_likelihood(pfmin))
 
 

### 2012-04-23 py/lomb_scargle_censor_test.py
lomb_scargle_censor_test.py tests nat's period finder code

@@ -1,0 +1,31 @@
+## code to test nat's function
+## by James Long on April 23, 2012
+
+import numpy as np
+from lomb_scargle_censor import lomb
+
+# get data in nat's preferred form
+i = "../data/lc_085647-6242.9.dat"
+lc = np.fromfile(i, dtype=float, count=-1, sep=" ")
+lc = lc.reshape((lc.size / 3,3))
+time_missing = lc[lc[:,1] > 29.9,0]
+lc = lc[lc[:,1] < 29.99,:]
+
+# l.s. grid search parameters
+min_freq = .002
+max_freq = 1.
+df = .000001
+numf = int((max_freq - min_freq) / df)
+
+## best period using censored data
+results = lomb(lc[:,0],lc[:,1],lc[:,2],time_missing,min_freq,df,numf)
+best = np.argmax(results)
+period = 1 / (min_freq + best * df)
+print period
+
+## best period ignoring censoring
+none_missing = np.array([],dtype='float64')
+results = lomb(lc[:,0],lc[:,1],lc[:,2],none_missing,min_freq,df,numf)
+best = np.argmax(results)
+period = 1 / (min_freq + best * df)
+print period

### 2012-04-24 py/censoring2.py
pushed over asas catalog

@@ -52,7 +52,7 @@ class Censored:
         """ 
         computes log p(D|theta,I)
         everything else: same notation as paper (all floats)
-        hella ugly unpacking from par[]; you're so hella NorCal
+        hella ugly unpacking from par[]
         """
         P = par[0]
         A0 = par[1]
@@ -77,12 +77,15 @@ class Censored:
                np.sum(self.loglikelihood_observed(eta2,B,VB,Vsig,S,VS))
     
     def loglikelihood_censored(self,eta2,B,VB,S,VS):
-    ## integrand of equation (12), integrate this across sig2
+        """
+        likelihood function for censored data
+        integrand of equation (12), integrated across sig2
+        """
         def sig_integrand(sig2,ui,eta2,B,VB,S,VS):
             #print 'gauss: ' +  str(gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui))
             #print gamma_pdf(sig2,S,VS)
             return gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui) * gamma_pdf(sig2,S,VS)
-        return np.log([integrate.quad(sig_integrand,np.max([0.,S-5.*np.sqrt(VS)]),S+5.*np.sqrt(VS),(ui,eta2,B,VB,S,VS),\
+        return np.log([integrate.quad(sig_integrand,np.max([0.,S-3.*np.sqrt(VS)]),S+3.*np.sqrt(VS),(ui,eta2,B,VB,S,VS),\
                                       epsabs=self.tolquad)[0] for ui in self.uc])
 
     
@@ -97,6 +100,7 @@ class Censored:
 
     def loglikelihood_observed(self,eta2,B,VB,Vsig,S,VS):
         """
+        log likelihood function for observed data, marginalizing over the sig2
         loglikelihood_observed_fast was created because this function fails for large values of VS
         """
         def integrand(sig2,ui,fi,ei2,eta2,B,VB,Vsig,S,VS):
@@ -105,14 +109,20 @@ class Censored:
             p_si = gamma_pdf(ei2,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
[…]

### 2012-04-24 py/fit_censoring.py

@@ -24,26 +24,33 @@ from lomb_scargle_censor import lomb as lombc
 
 path = '/Users/jwrichar/Documents/CDI/CensoredData/'
 
-def periods_lomb(t, f, e, n_per=3,f0=1./1000,df=1e-5,fn=5.):
+def periods_lomb(t, f, e, n_per=3,f0=1./1000,df=1e-5,fn=1./50.,plot=False):
 
     numf = int((fn-f0)/df)
-    #freqgrid = np.linspace(f0,fn,numf)
+    freqgrid = np.linspace(f0,fn,numf)
     ftest = 1.*f
     P_cand = []
     for i in xrange(n_per):
         psd,res = lombr(t,ftest,e,f0,df,numf, detrend_order=1,nharm=8)
+        if(plot):
+            plt.plot(freqgrid,psd)
+            plt.show()
+
         P_cand.append(1./res['freq'])
         ftest -= res['model']
     return P_cand
 
 
-def period_lombc(t, f, e, tc,f0=1./1000,df=1e-5,fn=5., n_per = 5):
+def period_lombc(t, f, e, tc,f0=1./1000,df=1e-5,fn=1./50., n_per = 5, plot=False):
 
     numf = int((fn-f0)/df)
     freqgrid = np.linspace(f0,fn,numf)
     ftest = 1.*f
     # lomb-scargle periodogram
     psd = lombc(t,ftest,e,tc,f0,df,numf)
+    if(plot):
+        plt.plot(freqgrid,psd)
+        plt.show()
 
     indpeak = find_peaks(psd) # find peaks in LS
     indtop = np.where(psd[indpeak] >= np.sort(psd[indpeak])[-n_per])[0] 
@@ -60,7 +67,7 @@ def find_peaks(x):
 def load_lc_from_web(ID):
     urlpath = 'http://www.astrouw.edu.pl/cgi-asas/asas_cgi_get_data?'
     ur = ulib.urlopen( urlpath + ID + ',asas3')#open url
-    miradata = np.loadtxt(ur.readlines(), usecols=(0,3,8,11), \
+    miradata = np.loadtxt(ur.readlines(), usecols=(0,2,7,11), \
                           dtype=[('t',np.float64), ('m',np.float64), ('e',np.float64),\
                                  ('grade','S2')])
     use = np.where(np.logical_and(np.logical_and(miradata['grade'] != 'D',miradata['grade'] != 'F'),\
@@ -76,10 +83,10 @@ def catchnan(p):
[…]

### 2012-04-24 py/fit_all_miras.py
added routine to do multi-processing on all Miras

@@ -1,0 +1,157 @@
+"""
+This file is part of the Censored Data Project
+Copyright 2012, Joseph W. Richards, David W. Hogg, James Long
+
+This code analyzes all ASAS Mira objects using parallel processing
+
+Issues:
+
+"""
+
+if __name__ == '__main__':
+    import matplotlib
+    matplotlib.use('Agg')
+    from matplotlib import rc
+    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
+    rc('text', usetex=True)
+
+
+from multiprocessing import Pool
+import censoring2 as c2
+import cProfile
+import numpy as np
+import matplotlib.pylab as plt
+import sys, os
+import urllib2 as ulib
+
+from fit_censoring import *
+
+global nper
+nper = 2
+
+global path
+path = '.'
+
+def doMira(ind, catalog):
+    """
+    analyze a single ASAS mira variable.  This method is meant to be used with the
+    multiprocessing module to analyze the entire ASAS Mira data set
+    """
+
+    
+    print '   #### doing mira ' + ': ' + str(catalog['ID'][ind]) +\
+          ' dotAstro: ' + str(catalog['dID'][ind])
+
+    new_periods = np.zeros((1,),dtype=[('ID','S16'), ('oldP',np.float), ('oldA',np.float),\
+                                       ('newP',np.float), ('newA',np.float),('llik',np.float)])
+    
+    data = load_lc_from_web(catalog['ID'][ind])
+    
+    indo = np.where(data['m'] != 29.999)
+    tobs = data['t'][indo]
+    mobs = data['m'][indo]
+    eobs = data['e'][indo]
+    tcen = data['t'][np.where(data['m'] == 29.999)]
+
+    # original L-S period (w/o using censored data)
+    Porig = periods_lomb(tobs, mobs, eobs, n_per = 1)[0]
+    new_periods[0][0] = catalog['ID'][ind]
+    new_periods[0][1] = Porig
+
+    # instantiate Censor object
+    cmodel = c2.Censored(tobs, mobs, eobs, tcen, name = catalog['ID'][ind], tolquad=1.)
+
+    # get original amplitude
+    orig = cmodel.get_init_par(Porig)
+    new_periods[0][2] = np.sqrt(orig[2]**2 + orig[3]**2)
+
[…]

### 2012-04-24 py/fit_censoring.py
added all of Nat's fitting code

@@ -16,8 +16,6 @@ import urllib2 as ulib
 
 import pdb
 
-sys.path.append(os.path.abspath(os.environ.get("TCP_DIR") + \
-                                      '/Algorithms/fitcurve'))
 from lomb_scargle import lomb as lomb
 from lomb_scargle_refine import lomb as lombr
 from lomb_scargle_censor import lomb as lombc
@@ -99,7 +97,7 @@ if __name__ == "__main__":
 #                                                 ('newP',np.float), ('newA',np.float),('llik',np.float)])
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(0,10):
+    for jj in np.arange(0,2):
     #for jj in np.arange(486,487):
         # load in data from web
         print '   #### doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) \

### 2012-04-24 py/fit_all_miras.py
changed multiprocessing call

@@ -26,11 +26,18 @@ import urllib2 as ulib
 
 from fit_censoring import *
 
-global nper
+#global nper
 nper = 2
 
-global path
-path = '.'
+#global path
+path = '../'
+
+#global catalog
+cat_data = np.loadtxt(path + 'data/asas_class_catalog_v2_3.dat',\
+                     usecols=(0,1,4,5,6,9,37), skiprows=1,\
+                     dtype=[('ID','S20'), ('dID','i4'), ('class','S20'), ('Pclass',np.float), \
+                            ('anom',np.float64), ('Pmira',np.float), ('P','f16')], delimiter=',')
+
 
 def doMira(ind, catalog):
     """
@@ -38,7 +45,6 @@ def doMira(ind, catalog):
     multiprocessing module to analyze the entire ASAS Mira data set
     """
 
-    
     print '   #### doing mira ' + ': ' + str(catalog['ID'][ind]) +\
           ' dotAstro: ' + str(catalog['dID'][ind])
 
@@ -135,23 +141,21 @@ def doMira(ind, catalog):
     
     return catalog['ID'][ind]
 
+def doMira_partial(ind):
+    return doMira(ind, cat_data)
 
 if __name__ == '__main__':
 
-    from functools import partial
+#    from functools import partial
 
-    cat_data = np.loadtxt(path + 'data/asas_class_catalog_v2_3.dat',\
-                          usecols=(0,1,4,5,6,9,37), skiprows=1,\
-                          dtype=[('ID','S20'), ('dID','i4'), ('class','S20'), ('Pclass',np.float), \
-                                 ('anom',np.float64), ('Pmira',np.float), ('P','f16')], delimiter=',')
 
-    partial_doMira = partial(doMira, catalog=cat_data)
+#    partial_doMira = partial(doMira, catalog=cat_data)
 
     p_mira = 0.75
     miras = np.where(np.logical_and(cat_data['Pmira'] > p_mira , cat_data['anom'] < 3.))[0] 
 
     pool = Pool(processes=2)
 
-    result = pool.map(partial_doMira, miras[0:4])
+    result = pool.map(doMira_partial, miras[0:4])
     pool.close()
     pool.join()

### 2012-04-24 py/fit_censoring.py
fixed bug in load_lc_from_web

@@ -65,7 +65,7 @@ def find_peaks(x):
 def load_lc_from_web(ID):
     urlpath = 'http://www.astrouw.edu.pl/cgi-asas/asas_cgi_get_data?'
     ur = ulib.urlopen( urlpath + ID + ',asas3')#open url
-    miradata = np.loadtxt(ur.readlines(), usecols=(0,2,7,11), \
+    miradata = np.loadtxt(ur, usecols=(0,2,7,11), \
                           dtype=[('t',np.float64), ('m',np.float64), ('e',np.float64),\
                                  ('grade','S2')])
     use = np.where(np.logical_and(np.logical_and(miradata['grade'] != 'D',miradata['grade'] != 'F'),\

### 2012-04-25 py/censoring2.py
added functionality to have known b's for the censored observations in censoring2.py

@@ -24,7 +24,7 @@ import scipy.optimize as op
 
 class Censored:
 
-    def __init__(self, t, m, em, tc, name = "Awesome LC", tolquad = 0.1):
+    def __init__(self, t, m, em, tc, b = None, name = "Awesome LC", tolquad = 0.1):
         """ 
         t: times when not censored (1-d ndarray)
         f: flux when not censored (1-d ndarray)
@@ -41,6 +41,7 @@ class Censored:
         self.ef = magerr2fluxerr(self.m, self.em)
         self.ef2 = self.ef**2
         self.name = name
+        self.b = b # known upper limits, if given
         
         unittests()
         return None
@@ -70,11 +71,19 @@ class Censored:
         self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1))
         self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1))
 
-        if(fast):
+        if(self.b==None): # if no upper limits are given
+            if(fast):
+                return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
+                       np.sum(self.loglikelihood_observed_fast(eta2,B,VB,Vsig))
             return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
-                   np.sum(self.loglikelihood_observed_fast(eta2,B,VB,Vsig))
-        return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
-               np.sum(self.loglikelihood_observed(eta2,B,VB,Vsig,S,VS))
+                   np.sum(self.loglikelihood_observed(eta2,B,VB,Vsig,S,VS))
+        else: # if upper limits are given ONLY FOR CENSORED DATA
+            assert(len(self.b) == len(self.tc))
+            if(fast):
+                return np.sum(self.loglikelihood_bgiven_censored(eta2,S,VS)) + \
+                       np.sum(self.loglikelihood_observed_fast(eta2,Vsig))
+            return np.sum(self.loglikelihood_bgiven_censored(eta2,S,VS)) + \
+                   np.sum(self.loglikelihood_observed(eta2,Vsig,S,VS))
     
     def loglikelihood_censored(self,eta2,B,VB,S,VS):
         """
@@ -87,8 +96,7 @@ class Censored:
[…]

### 2012-04-25 py/fit_censoring.py

@@ -97,7 +97,7 @@ if __name__ == "__main__":
 #                                                 ('newP',np.float), ('newA',np.float),('llik',np.float)])
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(0,2):
+    for jj in np.arange(22,23):
     #for jj in np.arange(486,487):
         # load in data from web
         print '   #### doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) \

### 2012-04-26 py/censoring2.py
made code more robust to numerical errors in gamma_pdf function

@@ -71,6 +71,9 @@ class Censored:
         self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1))
         self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1))
 
+        #print 'P=' + str(P) + 'A0 = ' + str(A0)+ ' A1 = '+ str(A1)+' B1 = '+ str(B1)
+        #print 'eta2 = ' + str(eta2)+ ' B = '+ str(B)+' VB = '+ str(VB)+ ' S = '+ str(S)+ ' VS = '+ str(VS)
+        #print np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS))
         if(self.b==None): # if no upper limits are given
             if(fast):
                 return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
@@ -104,6 +107,9 @@ class Censored:
         """
         p_not_cens = gaussian_cdf(self.f, B, VB)
         p_flux = gaussian_pdf(self.f, self.u, self.ef2 + eta2 * self.u**2)
+        logarg = p_not_cens * p_flux
+        if np.min(logarg) < 1.e-300:
+            logarg[np.where(logarg < 1.e-300)[0]] = 1.e-1
         return np.log(p_not_cens * p_flux)
 
     def loglikelihood_observed(self,eta2,B,VB,Vsig,S,VS):
@@ -117,7 +123,7 @@ class Censored:
             p_si = gamma_pdf(ei2,sig2,Vsig)
             p_sig2 = gamma_pdf(sig2,S,VS)
             return p_not_cens * p_flux * p_si * p_sig2
-        return np.log([integrate.quad(integrand,np.max([0.,S-3.*np.sqrt(VS)]),S+3.*np.sqrt(VS),(ui,fi,ei2,eta2,B,VB,Vsig,S,VS),\
+        return np.log([integrate.quad(integrand,np.max([0.,S-5.*np.sqrt(VS)]),S+5.*np.sqrt(VS),(ui,fi,ei2,eta2,B,VB,Vsig,S,VS),\
                                       epsabs=self.tolquad)[0] for (ui,fi,ei2) in zip(self.u,self.f,self.ef2)])
 
     def loglikelihood_bgiven_censored(self,eta2,S,VS):
@@ -127,7 +133,7 @@ class Censored:
         """
         def sig_integrand(sig2,bi,ui,eta2,S,VS):
[…]

### 2012-04-26 py/fit_censoring.py

@@ -97,8 +97,8 @@ if __name__ == "__main__":
 #                                                 ('newP',np.float), ('newA',np.float),('llik',np.float)])
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(22,23):
-    #for jj in np.arange(486,487):
+    for jj in np.arange(1,2):
+    #for jj in np.arange(440,441):
         # load in data from web
         print '   #### doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) \
               + ' dotAstro: ' + str(catalog['dID'][miras[jj]])
@@ -148,22 +148,24 @@ if __name__ == "__main__":
 
             # initialized at Nat's period
             p0 = cmodel.get_init_par(Pinit[i])
-            pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=500,fast=True)
+#            pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,mfev=500,fast=True)
+            pfmin = cmodel.optim_fmin_bfgs(p0,maxiter=1000,gtol=1.e-1,fast=True)
             #print pfmin
             params[(2*i),:] = pfmin
             lliks.append(cmodel.log_likelihood(pfmin))
 
             # initialized at twice Nat's period
             p0 = cmodel.get_init_par(2 * Pinit[i])
-            pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=500,fast=True)
+            #pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,mfev=500,fast=True)
+            pfmin = cmodel.optim_fmin_bfgs(p0,maxiter=1000,gtol=1.e-1,fast=True)
             params[((2*i)+1),:] = pfmin
             lliks.append(cmodel.log_likelihood(pfmin))
 
 
         print lliks
-        print 'Optimized log-likelihood: ' + str(np.max(lliks))
+        print 'Optimized log-likelihood: ' + str(np.nanmax(lliks))
         # optimal parameter vector
-        pstar = params[np.argmax(lliks),:]
+        pstar = params[np.where(lliks==np.nanmax(lliks))[0],:][0]
         #print pstar
         new_periods[0][3] = pstar[0]
         new_periods[0][4] = np.sqrt(pstar[2]**2 + pstar[3]**2)

### 2012-04-26 py/censoring2.py
model is now running on all miras

@@ -60,10 +60,10 @@ class Censored:
         A1 = par[2]
         B1 = par[3]
         eta2 = par[4]**2
-        B = par[5]
+        B = np.abs(par[5])
         VB = par[6]**2
         Vsig = par[7]**2
-        S = par[8]
+        S = np.abs(par[8])
         VS = par[9]**2
 
     ## convert all times to expected flux (i.e. t_i -> u_i)
@@ -109,8 +109,8 @@ class Censored:
         p_flux = gaussian_pdf(self.f, self.u, self.ef2 + eta2 * self.u**2)
         logarg = p_not_cens * p_flux
         if np.min(logarg) < 1.e-300:
-            logarg[np.where(logarg < 1.e-300)[0]] = 1.e-1
-        return np.log(p_not_cens * p_flux)
+            logarg[np.where(logarg < 1.e-300)[0]] = 1.e-100
+        return np.log(logarg)
 
     def loglikelihood_observed(self,eta2,B,VB,Vsig,S,VS):
         """
@@ -175,6 +175,7 @@ class Censored:
     def optim_fmin(self, p0, maxiter=1000, ftol=0.0001, xtol=0.0001, mfev=1.e8, fast=True):
         """
         maximize the log-likelihood function with respect to the 9 model parameters
+        fmin is more robust to numerical errors, as it can find its way out of nan's and inf's
         """
         opt = op.fmin(self.negll, p0, args = (fast,), maxiter=maxiter,ftol=ftol,maxfun=mfev)
         return opt
@@ -182,6 +183,7 @@ class Censored:
     def optim_fmin_bfgs(self, p0, maxiter=1000, gtol=0.0001, fast=True):
         """
         maximize the log-likelihood function with respect to the 9 model parameters using BFGS
+        BFGS often converges in fewer iterations, but will check out on a nan or inf
         """
         opt = op.fmin_bfgs(self.negll, p0, args = (fast,), gtol=gtol, maxiter=maxiter)
         return opt

### 2012-04-26 py/fit_all_miras.py

@@ -74,11 +74,11 @@ def doMira(ind, catalog):
     # do Lomb-scargle search to get top few periods
     Pinit = period_lombc(tobs, mobs, eobs, tcen, n_per = nper,df=1e-5)
     p0 = cmodel.get_init_par(Pinit[0])
-    ll0 = cmodel.log_likelihood(p0, fast=True)
+    #ll0 = cmodel.log_likelihood(p0, fast=True)
 
-    print 'Log-likelihood at p0: ' + str(ll0)
-    if np.isnan(ll0):
-        raise Exception
+    #print 'Log-likelihood at p0: ' + str(ll0)
+    #if np.isnan(ll0):
+    #    raise Exception
 
     lliks = []
     params = np.zeros((2*nper,10))
@@ -96,9 +96,9 @@ def doMira(ind, catalog):
         lliks.append(cmodel.log_likelihood(pfmin))
 
     print lliks
-    print 'Optimized log-likelihood: ' + str(np.max(lliks))
+    print 'Optimized log-likelihood: ' + str(np.nanmax(lliks))
     # optimal parameter vector
-    pstar = params[np.argmax(lliks),:]
+    pstar = params[np.where(lliks==np.nanmax(lliks))[0],:][0]
     new_periods[0][3] = pstar[0]
     new_periods[0][4] = np.sqrt(pstar[2]**2 + pstar[3]**2)
     new_periods[0][5] = np.max(lliks)
@@ -107,7 +107,7 @@ def doMira(ind, catalog):
     print 'Old Period: ' + str(round(Porig,3))
         
     # save result as line in txt file
-    np.savetxt(path + 'plots/bestpar_'+ catalog['ID'][ind] + '.dat', pstar)
+#    np.savetxt(path + 'plots/bestpar_'+ catalog['ID'][ind] + '.dat', pstar)
     
     # plot folded model with new method
     plt.clf()
@@ -128,8 +128,6 @@ def doMira(ind, catalog):
     cmodel.plot(ax, pold, fold=True, plot_model = False)
     plt.savefig(path + 'plots/bestfit_'+ catalog['ID'][ind] +'_old.png')
     # plot in mags
-    pold = pstar
-    pold[0] = catalog['P'][ind]
     plt.clf()
     ax = plt.subplot(111)
     cmodel.plot(ax, pold, fold=True, plot_model = False, mag = True)
@@ -146,16 +144,11 @@ def doMira_partial(ind):
 
 if __name__ == '__main__':
 
[…]

### 2012-04-26 py/fit_censoring.py

@@ -20,7 +20,7 @@ from lomb_scargle import lomb as lomb
 from lomb_scargle_refine import lomb as lombr
 from lomb_scargle_censor import lomb as lombc
 
-path = '/Users/jwrichar/Documents/CDI/CensoredData/'
+path = '../'
 
 def periods_lomb(t, f, e, n_per=3,f0=1./1000,df=1e-5,fn=1./50.,plot=False):
 
@@ -29,7 +29,7 @@ def periods_lomb(t, f, e, n_per=3,f0=1./1000,df=1e-5,fn=1./50.,plot=False):
     ftest = 1.*f
     P_cand = []
     for i in xrange(n_per):
-        psd,res = lombr(t,ftest,e,f0,df,numf, detrend_order=1,nharm=8)
+        psd,res = lombr(t,ftest,e,f0,df,numf, detrend_order=1,nharm=1)
         if(plot):
             plt.plot(freqgrid,psd)
             plt.show()
@@ -97,7 +97,7 @@ if __name__ == "__main__":
 #                                                 ('newP',np.float), ('newA',np.float),('llik',np.float)])
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(1,2):
+    for jj in np.arange(0,10):
     #for jj in np.arange(440,441):
         # load in data from web
         print '   #### doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) \
@@ -148,16 +148,16 @@ if __name__ == "__main__":
 
             # initialized at Nat's period
             p0 = cmodel.get_init_par(Pinit[i])
-#            pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,mfev=500,fast=True)
-            pfmin = cmodel.optim_fmin_bfgs(p0,maxiter=1000,gtol=1.e-1,fast=True)
+            pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,mfev=1000,fast=True)
+            #pfmin = cmodel.optim_fmin_bfgs(p0,maxiter=1000,gtol=1.e-1,fast=True)
             #print pfmin
             params[(2*i),:] = pfmin
             lliks.append(cmodel.log_likelihood(pfmin))
 
             # initialized at twice Nat's period
             p0 = cmodel.get_init_par(2 * Pinit[i])
[…]

### 2012-04-26 py/censoring2.py
added trend line and known b_i functionality

@@ -24,7 +24,7 @@ import scipy.optimize as op
 
 class Censored:
 
-    def __init__(self, t, m, em, tc, b = None, name = "Awesome LC", tolquad = 0.1):
+    def __init__(self, t, m, em, tc, b = None, f0 = 1.e6, name = "Awesome LC", tolquad = 0.1):
         """ 
         t: times when not censored (1-d ndarray)
         f: flux when not censored (1-d ndarray)
@@ -32,22 +32,25 @@ class Censored:
         tc: times of censored observations (1-d ndarray)
         """      
         self.tolquad = tolquad # numerical tolerance in integration
+        self.f0 = f0
         # insert data 
         self.t = t
         self.m = m
         self.em = em
         self.tc = tc
-        self.f = mag2flux(self.m)
-        self.ef = magerr2fluxerr(self.m, self.em)
+        self.f = mag2flux(self.m, f0=self.f0)
+        self.ef = magerr2fluxerr(self.m, self.em, f0=self.f0)
         self.ef2 = self.ef**2
         self.name = name
         self.b = b # known upper limits, if given
+        if(self.b != None):
+            self.bf = mag2flux(self.b, f0 = self.f0)
         
         unittests()
         return None
 
-    def mu(self, times, omega, A0, A1, B1):
-        return A0 + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
+    def mu(self, times, omega, A0, A1, B1, B0=0.):
+        return A0 + B0 * times + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
         
     def log_likelihood(self, par, fast = True):
         """ 
@@ -65,11 +68,17 @@ class Censored:
         Vsig = par[7]**2
         S = np.abs(par[8])
         VS = par[9]**2
+        if len(par) > 10:
+            B0 = par[10]
 
     ## convert all times to expected flux (i.e. t_i -> u_i)
-        w =  (2 * np.pi)/ P 
-        self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1))
-        self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1))
+        w =  (2 * np.pi)/ P
+        if len(par) == 10:
+            self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, B0 = 0.), f0=self.f0)
[…]

### 2012-04-26 py/fit_all_miras.py

@@ -45,8 +45,13 @@ def doMira(ind, catalog):
     multiprocessing module to analyze the entire ASAS Mira data set
     """
 
-    print '   #### doing mira ' + ': ' + str(catalog['ID'][ind]) +\
+    print '   #### doing mira: ' + str(catalog['ID'][ind]) +\
           ' dotAstro: ' + str(catalog['dID'][ind])
+
+    f = open(path + 'doMira.log', 'a')
+    f.write( 'starting mira: ' + str(catalog['ID'][ind]) +\
+          ' dotAstro: ' + str(catalog['dID'][ind])),'\n'
+    f.close()
 
     new_periods = np.zeros((1,),dtype=[('ID','S16'), ('oldP',np.float), ('oldA',np.float),\
                                        ('newP',np.float), ('newA',np.float),('llik',np.float)])

### 2012-04-26 py/censoring2.py
fixed bug: when mag model gets very large, the mag2flux conversion returns inf. I now restrict the magnitude to greater than -10.

@@ -80,6 +80,8 @@ class Censored:
             self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, B0), f0=self.f0)
             self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1, B0), f0=self.f0)
 
+        #print 'muc = ' + str(self.uc)
+        #print 'muc mag = ' + str(self.mu(self.tc, w, A0, A1, B1, B0))
         #print 'P=' + str(P) + 'A0 = ' + str(A0)+ ' A1 = '+ str(A1)+' B1 = '+ str(B1)
         #print 'eta2 = ' + str(eta2)+ ' B = '+ str(B)+' VB = '+ str(VB)+ ' S = '+ str(S)+ ' VS = '+ str(VS)
         #print np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS))
@@ -106,7 +108,7 @@ class Censored:
             #print 'gauss: ' +  str(gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui))
             #print gamma_pdf(sig2,S,VS)
             return gaussian_cdf(B, ui, VB + sig2 + eta2 * ui * ui) * gamma_pdf(sig2,S,VS)
-        return np.log([integrate.quad(sig_integrand,np.max([0.,S-3.*np.sqrt(VS)]),S+3.*np.sqrt(VS),(ui,eta2,B,VB,S,VS),\
+        return np.log([integrate.quad(sig_integrand,np.max([0.,S-5.*np.sqrt(VS)]),S+5.*np.sqrt(VS),(ui,eta2,B,VB,S,VS),\
                                       epsabs=self.tolquad)[0] for ui in self.uc])
    
     def loglikelihood_observed_fast(self,eta2,B,VB,Vsig):
@@ -312,9 +314,13 @@ def gaussian_pdf(x, mean, var):
     #if(np.any(var <= 0)):
     #    print x, mean, var
     #    raise Exception
+    if(np.any(var < 1.e-20)):
+        return 1.e-10
     return (oneoversqrt2pi/np.sqrt(var) * np.exp(-0.5 * (x - mean)**2 / var) )
 
 def gaussian_cdf(x, mean, var):
+    if(np.any(var < 1.e-20)):
+        return 1.e-10
     return .5*(1. + erf(oneoversqrt2 * (x - mean)/np.sqrt(var)) ) # look up correct form
 
 def gamma_pdf(x,mean,var):
[…]

### 2012-04-27 tex/ms.tex
added functionality to have known b's for the censored observations in censoring2.py
modified document to reflect changes in code

@@
 We model the brightness in magnitude because periodic variation in magnitude is more sinusoidal than in flux.
-However since errors are more Gaussian in flux space, we convert the $u_i'$ to flux measurements using
+However since errors more approximately Gaussian in flux space, we convert the $\mu_i'$ to flux  using
 \begin{equation*}
-\mu_i = 10^6 \times 10^{-.4 \mu_i'}
+\mu_i = 10^6 \times 10^{-0.4 \mu_i'}
 \end{equation*}
@@
 In Equation \ref{eq:mlik}, in the epochs for which a detection was made ($\fobs \ne \texttt{NA}$), we marginalize over the unknown $b_i$ up to the observed flux $\fobs$, enforcing that the detection limit be fainter than the observed brightness.
-Computationally, this requires a call to a Gaussian pdf, a call to a Gaussian cdf, and a multiplication.
+Computationally, this requires a call to a Gaussian probability density function (pdf), a call to a Gaussian cumulative distribution function (cdf), and a multiplication.
 In the epochs for which no detection was made ($\fobs = \texttt{NA}$), we compute the probability that the unknown detection threshold $b_i$ is greater than the unknown true flux $f_i^*$.
-When both $b_i$ and $f_i^*$ are normally distributed, this can be computed using a single call to a Gaussian cumulative distribution function.
+When both $b_i$ and $f_i^*$ are normally distributed, this can be computed using a single call to a Gaussian cdf.
 
@@
 \end{eqnarray}
-where the boolean indictor function, $I(\ftrue \ge b_i)$, is 1 if $\ftrue \ge b_i$ and 0 otherwise.
+where the boolean indictor function, $I(\fobs \ge b_i)$, is 1 if $\fobs \ge b_i$ and 0 otherwise.
 
@@
 \label{sec:uncertainty}
-Instead of assuming that $\sobs$ is a perfect, error-free measurement, we probabilistically connect it to the true photon count uncertainty variance $\sigma^2_i$ through a Gamma likelihood.
-Our probability of observed $\sobs$ given $\fobs$, $\sigma^2_i$, $\theta$ and $I$, is
+Instead of assuming that $\sobs$ is a perfect, error-free measurement, we can probabilistically connect it to the true photon flux uncertainty variance $\sigma^2_i$ through a Gamma likelihood.
+Our model probability of observed $\sobs$ given $\fobs$, $\sigma^2_i$, $\theta$ and $I$, is
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-where $\Gamma(x | m, V)$ is the standard Gamma distribution parameterized by its mean $m$ and variance $V$.
-The  model parameter $V_\sigma$ represents the
-variance in the distribution of reported uncertainties given the true
-uncertainty.  (Typically, the Gamma distribution is parameterized by parameters $(\alpha, \beta)$, where the mean is $\alpha\beta$ and the variance is $\alpha\beta^2$.)  The  purpose of
- including the likelihood in Equation \ref{eq:slik} to the model, in practice, is to keep the $\sigma^2_i$ from
-drifting very far away from the $\sobs$, as set by the hyperparameter
-$V_\sigma$. \james{maybe clarify this last sentence. isn't the main reason for having the $V_\sigma$ term because we don't believe the reported variance}
+where $\Gamma(x | m, V)$ is the standard Gamma distribution parameterized by its mean $m$ and variance $V$\footnote{The $\Gamma(x | m, V)$ pdf is $ p(x ) = (\theta^k \Gamma(k))^{-1}  x^{k-1} e^{x/ \theta}$, for $x \ge 0$, where $\theta = V / m$ and $k  = m^2 / V$.}.
+The  model hyperparameter $V_\sigma$ represents the
+variance in the distribution of reported uncertainties around the true
+uncertainty.
+Smaller values of $V_\sigma$ represent higher credence given to the reported uncertainties whereas larger values mean that our model is given more flexibility to use observation uncertainties that are far from the reported values.
+Ultimately, the model will choose the value of $V_\sigma$, for each light curve, that maximizes the data likelihood.
 
-We place a Gamma prior on the true photon count uncertainties, adding two parameters to our model.
+To complete our model of uncertainties on the reported observational error variances, $s_i^2$, we must place a prior distribution on the true uncertainties, $\sigma_i^2$.
+We choose to place a Gamma prior on the true photon flux uncertainties,
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-\james{should probably justify choice of some priors here / include some plots and/or refs. gamma priors and normal priors.}
+adding two parameters, $S$ and $V_S$, to our model.
+\james{should probably justify choice of some priors here / include some plots and/or refs.
+gamma priors and normal priors.}  This puts us in a position to write down our full likelihood model for light curves containing censored observations.
 
@@
 \begin{eqnarray}\displaystyle
+\label{eqn:lik}
 p(D_i|\theta,I) &=& \int_0^{\infty} p(\fobs |\sigma^2_i,\theta,I)\, p(\sobs | \fobs, \sigma^2_i,\theta,I)\, p(\sigma^2_i | \theta, I)\,\dd \sigma^2_i
 \\
+\label{eqn:likpar}
 \theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2\,\mu_i^2, B, V_B, V_{\sigma}, S, V_S) \quad ,
@@
 \begin{eqnarray}\displaystyle
+\label{eqn:likfull}
 p(D|\theta,I) &=& \prod_i p(D_i|\theta,I)
@@
 
-\joey{Describe how we optimize}
+Using the light curve data, $D$, from a particular source, our goal is to infer the parameters that describe the variability of that object.
+We do so by maximizing the likelihood in Equations (\ref{eqn:lik}), and (\ref{eqn:likfull}) with respect to the parameter vector in Equation (\ref{eqn:likpar}).
+The maximum likelihood estimate (MLE), which we call $\widehat{\theta}$ is the best estimate of the vector of model parameters that explain the variability of the object, given all of the data (observed and censored) and the assumed model.
+The MLE allows us to then infer parameters of physical interest, such as the period and amplitude of oscillation.
+
+To maximize the data likelihood with respect to $\theta$, we use the {\tt fmin} routine in the {\tt Python} module {\tt scipy.optimze}.
+The {\tt fmin} routine uses a a Nelder-Mead simplex algorithm to optimize a function with respect to its arguments starting from some initial guess, $\theta_0$.
+The convergence of the algorithm is highly sensitive to the value of $\theta_0$.
+To find the MLE parameters for each light curve, we use {\tt fmin} to minimize the negative log likelihood (negative natural log of Equation \ref{eqn:likfull}).
+We use a standard Lomb-Scargle Fourier analysis (ignoring any epochs of non-detection) to initialize the search algorithm, deriving an initial guess for $\omega$ as the frequency corresponding to the maximal Lomb-Scargle spectral density, and initializing $A_0$, and $\{A_k, B_k\}_{k=1}^K$ using standard least squares.
+In practice, we run the {\tt fmin} routine multiple times using different initial values of the angular frequency corresponding to the frequencies of the most significant spectral density peaks and values of one-half those frequencies.
+Initial values for the other parameters are set at $\eta^2 = 0.2$, $B = V_B = 2\cdot\min(\fobs)$, and $ V_{\sigma} = S = V_S = \textrm{med}(\sobs)$.
+
+Replacing likelihood with fast version where we trust $s_i^2$ for observed data.
 

### 2012-04-27 py/censoring2.py

@@ -87,8 +87,13 @@ class Censored:
         #print np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS))
         if(self.b==None): # if no upper limits are given
             if(fast):
-                return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
-                       np.sum(self.loglikelihood_observed_fast(eta2,B,VB,Vsig))
+                llc = np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS))
+                llo = np.sum(self.loglikelihood_observed_fast(eta2,B,VB,Vsig))
+                if(np.isnan(llc) or np.isnan(llo)):
+                    print 'LL is nan'
+                if(np.abs(llc) == np.inf or np.abs(llo) == np.inf):
+                    print 'LL is inf or -inf'
+                return llc + llo
             return np.sum(self.loglikelihood_censored(eta2,B,VB,S,VS)) + \
                    np.sum(self.loglikelihood_observed(eta2,B,VB,Vsig,S,VS))
         else: # if upper limits are given ONLY FOR CENSORED DATA
@@ -174,7 +179,7 @@ class Censored:
         par[4] = 0.2
         par[5] = 2.*np.abs(np.min(self.f))
         par[6] = par[5]
-        par[7] = np.median(self.ef) # JWR changed from ef2 to ef
+        par[7] = np.median(self.ef2) 
         par[8] = par[7]
         par[9] = par[8] # must be kept small or integration will take too long
         # also if par[9] is too large, the integration may return nan
@@ -321,20 +326,22 @@ def gaussian_pdf(x, mean, var):
 def gaussian_cdf(x, mean, var):
     if(np.any(var < 1.e-20)):
         return 1.e-10
-    return .5*(1. + erf(oneoversqrt2 * (x - mean)/np.sqrt(var)) ) # look up correct form
+    return .5*(1. + erf(oneoversqrt2 * (x - mean)/np.sqrt(var)) ) 
 
 def gamma_pdf(x,mean,var):
     theta = var / mean
     k = mean / theta
     if(np.log(gamma(k)) == np.inf):
         return 1.e-10
-    #print 'gamma PDF: ' + str(-k*np.log(theta)) +' ' + str(np.log(gamma(k))) + ' '+ str((k-1.)*np.log(x)) + ' ' +str(x/theta)
[…]

### 2012-04-27 py/fit_all_miras.py

@@ -50,7 +50,7 @@ def doMira(ind, catalog):
 
     f = open(path + 'doMira.log', 'a')
     f.write( 'starting mira: ' + str(catalog['ID'][ind]) +\
-          ' dotAstro: ' + str(catalog['dID'][ind])),'\n'
+          ' dotAstro: ' + str(catalog['dID'][ind]) +'\n')
     f.close()
 
     new_periods = np.zeros((1,),dtype=[('ID','S16'), ('oldP',np.float), ('oldA',np.float),\

### 2012-04-27 py/fit_censoring.py

@@ -97,7 +97,7 @@ if __name__ == "__main__":
 #                                                 ('newP',np.float), ('newA',np.float),('llik',np.float)])
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(0,10):
+    for jj in np.arange(64,65):
     #for jj in np.arange(440,441):
         # load in data from web
         print '   #### doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) \

### 2012-04-28 py/censoring2.py
fixed bug about having no non-detects

@@ -49,8 +49,8 @@ class Censored:
         unittests()
         return None
 
-    def mu(self, times, omega, A0, A1, B1, B0=0.):
-        return A0 + B0 * times + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
+    def mu(self, times, omega, A0, A1, B1, A01=0., A02 = 0.):
+        return A0 + A01 * times + A02 * times**2 + A1 * np.cos(omega * times) + B1 * np.sin(omega * times)
         
     def log_likelihood(self, par, fast = True):
         """ 
@@ -69,22 +69,31 @@ class Censored:
         S = np.abs(par[8])
         VS = par[9]**2
         if len(par) > 10:
-            B0 = par[10]
+            A01 = par[10]
+        if len(par) > 11:
+            A02 = par[11]
+
+        w =  (2 * np.pi)/ P
+
+        # handle case of no non-detections
+        if(len(self.tc) == 0):
+            if len(par) == 10:
+                self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, A01 = 0., A02=0.), f0=self.f0)
+            else:
+                self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, A01, A02=0.), f0=self.f0)
+            if(fast):
+                return np.sum(self.loglikelihood_observed_fast(eta2,B,VB,Vsig))
+            return np.sum(self.loglikelihood_observed(eta2,B,VB,Vsig,S,VS))
+ 
 
     ## convert all times to expected flux (i.e. t_i -> u_i)
-        w =  (2 * np.pi)/ P
         if len(par) == 10:
-            self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, B0 = 0.), f0=self.f0)
-            self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1, B0 = 0.), f0=self.f0)
+            self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, A01 = 0., A02=0.), f0=self.f0)
+            self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1, A01 = 0., A02=0.), f0=self.f0)
         else:
-            self.u  = mag2flux(self.mu(self.t,  w, A0, A1, B1, B0), f0=self.f0)
-            self.uc = mag2flux(self.mu(self.tc, w, A0, A1, B1, B0), f0=self.f0)
[…]

### 2012-04-28 py/fit_censoring.py

@@ -97,7 +97,7 @@ if __name__ == "__main__":
 #                                                 ('newP',np.float), ('newA',np.float),('llik',np.float)])
 
     #### FOR EACH MIRA IN miras ####
-    for jj in np.arange(64,65):
+    for jj in np.arange(0,1):
     #for jj in np.arange(440,441):
         # load in data from web
         print '   #### doing mira ' + str(jj) + ': ' + str(catalog['ID'][miras[jj]]) \

### 2012-07-05 tex/ms.tex
ms updated and TODO list added

@@
 
- Show P - A relationship before and after using the method
+Figure \ref{fig:miraP} shows the estimated periods for standard Lomb-Scargle against the periods from our censored Fourier model for 2538 likely Mira variables chosen from the MACC catalog (these were chosen such that $P$(Mira) $>0.5$.)  For 12.6\% of these objects, the censored period is twice that of the Lomb-Scargle period, while for 0.51\% it is triple (show examples).
+This occurs because of XYZ.
+For 85.5\% of all objects, the periods are consistent to $\pm$5\%.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{../plots/mira_periods_LS_vs_Censored.pdf}
+\end{center}
+\caption{ .  \label{fig:miraP}}
+\end{figure*}
+
+Figure \ref{fig:miraPA} shows the period-amplitude relationship for the set of Mira variables, with periods and amplitudes estimated by standard Lomb-Scargle (left) and the censored model (right).
+Results show that XYZ.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{../plots/mira_period_amplitude_relation.pdf}
+\end{center}
+\caption{ .  \label{fig:miraPA}}
+\end{figure*}
 
@@
 There is an RRL P - A relationship (strong linear anti-correlation)
+
+As a selection criterion (simulated data?)
 

### 2012-07-05 py/lomb_scargle_censor_test.py

@@ -12,8 +12,8 @@ time_missing = lc[lc[:,1] > 29.9,0]
 lc = lc[lc[:,1] < 29.99,:]
 
 # l.s. grid search parameters
-min_freq = .002
-max_freq = 1.
+min_freq = 1./1000.
+max_freq = 1./50.
 df = .000001
 numf = int((max_freq - min_freq) / df)
 

### 2013-03-26 tex/ms.tex
tiny typo in parameter list
author list, acks, moved figures

@@
 \shorttitle{Fitting the Undetected}
-\shortauthors{}
+\shortauthors{Richards, Hogg, Long, Foreman-Mackey}
 \title{Fitting the Undetected:  Light-curve Feature Estimation in the Presence of Non-detections}
 \author{
-Some wild \& crazy guys
+Joseph W. Richards\altaffilmark{1, 2, 3, *},
+David W. Hogg\altaffilmark{4, 5},
+James P. Long\altaffilmark{2},
+Daniel Foreman-Mackey\altaffilmark{4}
 }
 %
-%\altaffiltext{1}{Astronomy Department, University of California, Berkeley, CA, 94720-7450, USA}
-%\altaffiltext{2}{Statistics Department, University of California, Berkeley, CA, 94720-7450, USA}
-%\altaffiltext{*}{E-mail: {\tt jwrichar@stat.berkeley.edu}}
+\altaffiltext{1}{Astronomy Department, University of California, Berkeley, CA, 94720-7450, USA}
+\altaffiltext{2}{Statistics Department, University of California, Berkeley, CA, 94720-7450, USA}
+\altaffiltext{3}{wise.io, 2150 Shattuck Ave., suite 525, Berkeley, CA, 94704, USA}
+\altaffiltext{4}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
+\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{*}{\texttt{jwrichar@stat.berkeley.edu}}
 
 % Note the status (in preparation, MNRAS submitted date, ApJ accepted date, in press etc).
-\slugcomment{in preparation}
+\slugcomment{DRAFT 2013-03-26 / in preparation}
 
@@
 \label{eqn:likpar}
-\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2\,\mu_i^2, B, V_B, V_{\sigma}, S, V_S) \quad ,
+\theta &\equiv& (\omega, A_0, \{A_k, B_k\}_{k=1}^K, \eta^2, B, V_B, V_{\sigma}, S, V_S) \quad ,
 \end{eqnarray}
@@
 veracity.
-
-\begin{figure}[hbtp]
-  \begin{center}
-    \includegraphics[width=\textwidth,%
-      trim=3cm 12cm 5cm 3cm, clip=true]{gm/gm.pdf}
-  \end{center}
-  \caption{The graphical representation of the probability distribution.
-    \label{fig:graph}}
-\end{figure}
 
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[angle=0,width=6.5in]{../plots/mira_simulated.pdf}
-\end{center}
-\caption{ Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.
-Blue tick marks along the bottom axis denote phases where there were non-detections.
-As the original flux measurements are dimmed by higher factors, the troughs of the sinusoidal light curve are censored, resulting in an incomplete view of the data.
-\label{fig:mirasim}}
-\end{figure*}
-
 \subsection{Results of Fitting Miras}
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[angle=0,width=6.5in]{../plots/mira_periods_LS_vs_Censored.pdf}
-\end{center}
-\caption{ .  \label{fig:miraP}}
-\end{figure*}
-
 Figure \ref{fig:miraPA} shows the period-amplitude relationship for the set of Mira variables, with periods and amplitudes estimated by standard Lomb-Scargle (left) and the censored model (right).
 Results show that XYZ.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[angle=0,width=6.5in]{../plots/mira_period_amplitude_relation.pdf}
-\end{center}
-\caption{ .  \label{fig:miraPA}}
-\end{figure*}
 
@@
 
-\acknowledgements
+\acknowledgements It is a pleasure to thank Josh Bloom (Berkeley) for
+valuable discussions and comments.  Partial support for this project
+was provided by [JWR: PUT GRANT NUMBERS HERE], NASA grant NNX12AI50G,
+and NSF grant IIS-1124794.
 
@@
 
+\clearpage
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth,%
+      trim=3cm 12cm 5cm 3cm, clip=true]{gm/gm.pdf}
+  \end{center}
+  \caption{The graphical representation of the probability distribution.
+    \label{fig:graph}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{../plots/mira_simulated.pdf}
+\end{center}
+\caption{ Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.
+Blue tick marks along the bottom axis denote phases where there were non-detections.
+As the original flux measurements are dimmed by higher factors, the troughs of the sinusoidal light curve are censored, resulting in an incomplete view of the data.
+\label{fig:mirasim}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{../plots/mira_periods_LS_vs_Censored.pdf}
+\end{center}
+\caption{ .  \label{fig:miraP}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{../plots/mira_period_amplitude_relation.pdf}
+\end{center}
+\caption{ .  \label{fig:miraPA}}
+\end{figure}
+
 \end{document}

### 2013-03-26 tex/ms.tex
notes to selves

@@
-% Define the document class. aastex for manuscript preparation, emulateapj for making it look polished.
-% \documentclass[12pt,preprint2]{aastex}
-%\documentclass[12pt,a4paper, iop,onecolumn,numberedappendix]{emulateapj}
 \documentclass[12pt,preprint]{aastex}
@@
 \newcommand{\james}[1] { \textcolor{blue} {
-\ensuremath{\blacklozenge} {\bf james:}  {#1}
+\ensuremath{\blacklozenge} {\bf james says:}  {#1}
 \ensuremath{\blacklozenge} } }
@@
 \newcommand{\hogg}[1] { \textcolor{darkorange} {
-\ensuremath{\blacksquare} {\bf hogg:}  {#1}
+\ensuremath{\blacksquare} {\bf hogg says:}  {#1}
 \ensuremath{\blacksquare} } }
@@
 \newcommand{\dan}[1] { \textcolor{darkgreen} {
-\ensuremath{\bigstar} {\bf dan:}  {#1}
+\ensuremath{\bigstar} {\bf dan says:}  {#1}
 \ensuremath{\bigstar} } }
@@
 \newcommand{\joey}[1] { \textcolor{red} {
-\ensuremath{\clubsuit} {\bf joey:}  {#1}
+\ensuremath{\clubsuit} {\bf joey says:}  {#1}
 \ensuremath{\clubsuit} } }%\documentclass[manuscript]{aastex}
@@
   \begin{center}
+    [\hogg{PUT FIGURE HERE}]
+  \end{center}
+  \caption{An example of why this project is necessary.
+The points are...
+\hogg{this should be a figure showing heavily censored data and a traditional fit to them, folded AND UNFOLDED, showing that the traditional fit is way wrong.}\label{fig:badfit}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+  \begin{center}
     \includegraphics[width=\textwidth,%
@@
   \end{center}
-  \caption{The graphical representation of the probability distribution.
-    \label{fig:graph}}
+  \caption{The graphical representation of the probability distribution.\label{fig:graph}}
 \end{figure}
@@
 As the original flux measurements are dimmed by higher factors, the troughs of the sinusoidal light curve are censored, resulting in an incomplete view of the data.
-\label{fig:mirasim}}
+\hogg{This figure needs to show the OLD and NEW best-fit lightcurve models too!}\label{fig:mirasim}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+  \begin{center}
+    [\hogg{PUT FIGURE HERE}]
+  \end{center}
+  \caption{A real-data example.  This is the same source as that shown in \figurename~\ref{fig:badfit} but now showing the results of...\label{fig:goodfit}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+  \begin{center}
+    [\hogg{PUT FIGURE HERE}]
+  \end{center}
+  \caption{Another few real-data examples; in each case the method presented here either matches the traditional method or is substantially better.\label{fig:examples}}
 \end{figure}

### 2013-03-26 tex/ms.tex
new, reduced-scope abstract

@@
 detectable at all epochs.  In principle, the best approach is to
-perform photometric measurements anyway, so that there is a datum
-recorded at every epoch.  In practice, most surveys only provide
-detection upper limits at some epochs; in some not even upper
-limits but just the information that the star was not observed.  Here
-we demonstrate with real data on periodic variables from the All-Sky
-Automated Survey (ASAS) that these non-detections are nonetheless useful in
-fitting and parameter estimation; their inclusion improves both
-precision and accuracy.  One novel aspect of this work is that we do
-not require the data source to include accurate---or even
-any---information about the uncertainty variances on detections or the
-values of the upper limits; we find that we can model these as latent
-variables, even when they are different at every epoch.  Using realistic
-simulations of Mira variable stars, we obtain amplitude estimates that
-are {\bf XXX\%} more accurate by incorporating the non-detections into
-the analysis.  For  Mira and RR Lyrae variables observed by ASAS, our
-method obtains a tighter period - amplitude relationship than the standard
-technique.
+perform photometric measurements (forced photometry) anyway.
+%
+In practice, most surveys only provide detection upper limits at some
+epochs; in some not even upper limits but just the ``null''
+information that the star was not detected.
+%
+Here we demonstrate with real data on periodic variables from the
+All-Sky Automated Survey (ASAS) that these null observations can
+in fact be very useful in fitting and parameter estimation.
+%
+We build a novel light-curve model that includes the per-epoch
+detection thresholds as a set of latent parameters, one per data
+point; the best-fit model is the maximum-likelihood model after
+marginalizing out the latent thresholds.
+%
+The key assumption is that the detection thresholds are drawn from a
+stationary distribution (the parameters of which are free to vary).
+%
+We demonstrate that this censored-data model outperforms traditional
+fitting---in which null observations are discarded from the
+fits---in estimating both periods and amplitudes for stars subject to
+heavy censoring.
+%
+In particular, our model eradicates a failure mode in which
+traditional fitting obtains an integer fraction of the correct period.
 \end{abstract}

### 2013-03-26 tex/ms.tex
acks and section structure

@@
 
-\section{Formulating the Model}
+\section{The Model}
 \label{sec:model}
-
-%This note, based on the original note from JWR on Feb. 8, 2012, includes revisions from David Hogg, James Long, Nat Butler, and Josh Bloom.
 
@@
 
-\section{The Light Curve Model}
-\label{sec:model}
 \subsection{Modeling Non-Detections}
@@
 
-\acknowledgements It is a pleasure to thank Josh Bloom (Berkeley) for
-valuable discussions and comments.  Partial support for this project
-was provided by [JWR: PUT GRANT NUMBERS HERE], NASA grant NNX12AI50G,
-and NSF grant IIS-1124794.
+\acknowledgements It is a pleasure to thank
+%
+  Josh Bloom (Berkeley) and
+  Nat Butler (ASU)
+%
+for valuable discussions and comments.
+%
+Partial support for this project was provided by \hogg{[JWR: PUT GRANT
+  NUMBERS HERE]}, NASA grant NNX12AI50G, and NSF grant IIS-1124794.
 

### 2013-03-27 tex/ms.tex
figure 1 of ms

--- 2013-03-26 tex/ms.tex
@@
   \begin{center}
-    [\hogg{PUT FIGURE HERE}]
+\includegraphics[angle=0,width=6.5in]{../plots/fig_045832-0604_1.png}
   \end{center}
-  \caption{An example of why this project is necessary.
-The points are...
-\hogg{this should be a figure showing heavily censored data and a traditional fit to them, folded AND UNFOLDED, showing that the traditional fit is way wrong.}\label{fig:badfit}}
+  \caption{A comparison of the best fit obtained by the traditional method (left) and our censored data model (right) on the Mira variable ASAS 045832-0604.1.
+The top row shows the respective model fits to the raw light curves while the bottom row displays the period-folded data and models.
+The black dots denote the observed data and the red vertical dashes mark the epochs of null observation.
+The traditional fitting method clearly obtains half the true period while our method recovers the correct period.
+The traditional method fails because it identifies a reasonable high-frequency fit to the observed data which completely ignores the null observations.
+Note that alternate cycles of the traditional model fit contain no observed data.
+By contrast, our method fully accounts for the null observations using a probability model.\label{fig:badfit}}
 \end{figure}

### 2013-03-27 py/censoring2.py

@@ -237,7 +237,7 @@ class Censored:
 
 
 
-    def plot(self, ax, par, fold = False, plot_model = True, mag = False):
+    def plot(self, ax, par, fold = False, plot_model = True, mag = False, plot_title = True):
         '''
         input:
         - ax: matplotlib axes object
@@ -289,20 +289,20 @@ class Censored:
             y = self.m
             ey = self.em
             if(self.b == None):
-                yc = np.zeros_like(xc) + np.max(y) + 0.5
+                yc = np.zeros_like(xc) + np.max(y) + 0.25
             else:
                 yc = self.b
         else:
             y = self.f
             ey = self.ef
             yc = np.zeros_like(xc)
-        ax.plot(x - mediant, y, 'ko', alpha=0.5, mec='k')
+        ax.plot(x - mediant, y, 'ko', alpha=0.5, mec='k', markersize=4)
         hogg_errorbar(ax, x - mediant, y, ey)
-        ax.plot(xc - mediant, yc, 'r.', alpha=0.5, mec='r')
+        ax.plot(xc - mediant, yc, 'r|', alpha=0.3, mec='r', markersize=10)
         if(fold): # if fold, plot 2 phases of LC
-            ax.plot(x - mediant + 1., y, 'ko', alpha=0.5, mec='k')
+            ax.plot(x - mediant + 1., y, 'ko', alpha=0.5, mec='k', markersize=4)
             hogg_errorbar(ax, x - mediant + 1., y, ey)
-            ax.plot(xc - mediant + 1., yc, 'r.', alpha=0.5, mec='r')
+            ax.plot(xc - mediant + 1., yc, 'r|', alpha=0.3, mec='r', markersize=10)
         if(plot_model):
             if(mag):
                 mup = self.mu(tp, omega, A0, A1, B1, A01, A02)
@@ -312,9 +312,9 @@ class Censored:
                 mup = mag2flux(self.mu(tp, omega, A0, A1, B1, A01, A02), f0=self.f0)
                 mup_p = mup * (1. + np.sqrt(eta2))
                 mup_m = mup * (1. - np.sqrt(eta2))
-            ax.plot(xp - mediant, mup_p, 'b-', alpha=0.25)
-            ax.plot(xp - mediant, mup, 'b-', alpha=0.50)
-            ax.plot(xp - mediant, mup_m, 'b-', alpha=0.25)
[…]

### 2013-03-27 py/fit_all_miras.py

@@ -31,7 +31,7 @@ from fit_censoring import *
 nper = 2
 
 #global path
-path = '../'
+path = '/Users/jwrichar/Documents/CDI/CensoredData/'
 
 #global catalog
 cat_data = np.loadtxt(path + 'data/asas_class_catalog_v2_3.dat',\

### 2013-03-27 py/make_lc_figs.py

@@ -1,10 +1,10 @@
 
-if __name__ == '__main__':
-    import matplotlib
-    matplotlib.use('Agg')
-    from matplotlib import rc
-    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
-    rc('text', usetex=True)
+# if __name__ == '__main__':
+#     import matplotlib
+#     matplotlib.use('Agg')
+#     from matplotlib import rc
+#     rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':18})
+#     rc('text', usetex=True)
 
 
 import censoring2 as c2
@@ -26,6 +26,8 @@ def plot_fig(ID, path, nper=2):
     """ plot folded and unfolded LS and censored model fits
     """
     data = load_lc_from_web(ID)
+
+    print path + 'plots/fig_'+ ID +'.png'
 
     indo = np.where(data['m'] != 29.999)
     tobs = data['t'][indo]
@@ -60,13 +62,13 @@ def plot_fig(ID, path, nper=2):
         print "Fitting for period candidate %i of %i" % (i, nper)
         # initialized at Nat's period
         p0 = cmodel.get_init_par(Pinit[i])
-        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=1000,fast=True)
+        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=200,fast=True)
         params[(2*i),:] = pfmin
         lliks.append(cmodel.log_likelihood(pfmin))
 
         # initialized at twice Nat's period
         p0 = cmodel.get_init_par(2 * Pinit[i])
-        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=1000,fast=True)
+        pfmin = cmodel.optim_fmin(p0,maxiter=1000,ftol=1.,xtol=0.1,mfev=200,fast=True)
         params[((2*i)+1),:] = pfmin
         lliks.append(cmodel.log_likelihood(pfmin))
 
@@ -84,19 +86,25 @@ def plot_fig(ID, path, nper=2):
   
     # make plots
     plt.clf()
+    fig = plt.gcf()
+    fig.set_size_inches(12.0,8.0)
     # unfolded LS
     ax = plt.subplot(221)
-    cmodel.plot(ax, orig, fold=False, plot_model=True, mag=True)
[…]

### 2013-03-27 tex/ms.tex
reflow paragraph I am about to delete and rewrite
partial introduction outline; its now a huge mess

@@
 
-Surveys working in real-time tend to analyze each image with an eye to populating databases quickly.
-Unless a particular region of the sky gets special treatment (such as it is a known place of interest), discovery of variability happens by comparison of flux values at the database level.
-Analysis on such databases of transient/variables thus deals with censored data where there is often less information available than there should be.
-Even then, we might be able to infer what the thresholds were based on S/N measurements of objects neighboring the source of interest.
+Telescopes don't make catalogs, they make images (Hogg \& Lang 20xx).
+Nonetheless, some of the most important astronomical projects have produced
+enormously productive catalogs (the \textsl{SDSS}, for examle CITATION).
+There is a growing understanding that precise measurements
+of astronomical sources have to be made directly 
+in the imaging data
+(CITE Lang, Anderson, something in weak lensing):
+It is impossible to pass the full information content of the imaging to the end user
+via a catalog (which by necessity is built with point estimates of the quantities of
+interest).
+
+...time domain seems simple, because only photometry matters!
+
+...but catalogs make censored point estimates, usually... noise, morphology, CRs, defects...
+
+...catalogs could be better if they did forced photometry.  upper limits worse.
+nulls worse.  total obfuscation worst of all
+
+...here we are going to work with nulls; they can still be informative!
+
+...as an example, consider \figurename~\ref{fig:example}...
+
+Surveys working in real-time tend to analyze each image with an eye to
+populating databases quickly.
+Unless a particular region of the sky gets special treatment
+(such as it is a known place of interest),
+discovery of variability happens by comparison of flux values at the database level.
+Analysis on such databases of transient/variables thus deals with censored data
+where there is often less information available than there should be.
+Even then, we might be able to infer what the thresholds were based
+on S/N measurements of objects neighboring the source of interest.
 

### 2013-03-27 tex/ms.tex
intro paragraphs

@@
 Telescopes don't make catalogs, they make images (Hogg \& Lang 20xx).
-Nonetheless, some of the most important astronomical projects have produced
-enormously productive catalogs (the \textsl{SDSS}, for examle CITATION).
-There is a growing understanding that precise measurements
-of astronomical sources have to be made directly 
-in the imaging data
+Nonetheless, some of the most important astronomical projects have produced enormously productive catalogs
+(the \textsl{SDSS}, for examle CITATION).
+There is a growing understanding that precise measurements of astronomical sources have to be made directly in the imaging data
 (CITE Lang, Anderson, something in weak lensing):
-It is impossible to pass the full information content of the imaging to the end user
-via a catalog (which by necessity is built with point estimates of the quantities of
-interest).
+It is impossible to pass the full information content of the imaging to the end user via a catalog
+(which by necessity is built with point estimates of the quantities of interest).
 
-...time domain seems simple, because only photometry matters!
+The measurements of importance for variable-star science are relatively simple.
+Unlike the measurements used for astrometry, extragalactic transients, and weak lensing,
+where two-dimensional (morphological) information matters,
+when variable stars are being observed, all the end user wants is accurate photometry of each source at each epoch.
+One could hope that catalog-level information is sufficient:
+the detailed image morphology is not required for lightcurve extraction, except when fields are crowded.
+However, there are still limitations to working at catalog level:
+Catalog entries (nowadays) usually come with noise variance estimates,
+but they don't usually give detailed information about the locations or influence of
+bad detector pixels, cosmic rays, and flat-field variations;
+all these things might be needed to construct a precise photometric noise model or to understand residuals or artifacts.
+More importantly, catalogs are usually subject to hard \emph{censoring};
+the catalogs (nowadays) only include measurements that are deemed (by the software or catalog makers)
+to be \emph{signficant detections}.
 
-...but catalogs make censored point estimates, usually... noise, morphology, CRs, defects...
+This could change.
+In principle a catalog could contain a measurement at \emph{every} epoch for any source that is detected at \emph{any} epoch.
+This kind of ``forced photometry'' would include many non-significant measurements,
+but it would come closer to passing forward all the information in the imaging to users working at catalog level.
+We recommend this approach to the catalog makers of the future (especially those making the \textsl{LSST} catalogs).
 
-...catalogs could be better if they did forced photometry.  upper limits worse.
-nulls worse.  total obfuscation worst of all
+Less good than the comprehensive forced-photometry approach is to report, when measurements are not significant,
+a well-defined ``upper limit'' for the brightness of each source.
+This is rarely done, and upper limits are hard to use, but upper limits are an improvement over standard practice.
+Standard practice is to \emph{report nothing at all} when a source is not detected significantly.
+Indeed, it is often hard, when presented with a multi-epoch catalog,
+to determine whether the reason a particular source is missing from the imaging at a particular epoch
+is because the source was faint or because the source fell outside the imaging field of view!
 
-...here we are going to work with nulls; they can still be informative!
+In this contribution we look at improving catalog-level time-domain astronomy in the presence of \emph{censored data}.
+The problem we consider is that of period and amplitude modeling for large-amplitude variable stars in a data set
+in which \emph{no information is reported at all} at the epochs at which the star is not significantly detected.
+Furthermore, we will imagine that we also have \emph{no information at all} about the sensitivity of the telescope
+at the time of each non-detection; that is, we have no upper limits and we can't reconstruct them from any extant meta-data.
+The non-detections are literally database ``null'' values.
+This set-up may seem artificial, but it is in fact the situation we find ourselves in for various surveys,
+including the ASAS survey, data from which we will model below.
+What we will show is that \emph{the ``null'' non-detections are in fact very informative};
+amplitude and period fitting will be wrong in general if the non-detections are ignored.
+
+The benefit of using the non-detections is large, but the cost is also large:
+Use of these ``null'' data involves non-trivial modeling,
+and large numbers of latent parameters to be inferred and marginalized out.
+
+To make this all concrete, consider the photometric data from the \textsl{ASAS} survey
+depicted in \figurename~\ref{fig:example} for one Mira variable.
+The photometric detections are shown and so are the epochs at which there are nulls (or non-detections).
+The standard
 
 ...as an example, consider \figurename~\ref{fig:example}...
-
-Surveys working in real-time tend to analyze each image with an eye to
-populating databases quickly.
-Unless a particular region of the sky gets special treatment
-(such as it is a known place of interest),
-discovery of variability happens by comparison of flux values at the database level.
-Analysis on such databases of transient/variables thus deals with censored data
-where there is often less information available than there should be.
-Even then, we might be able to infer what the thresholds were based
-on S/N measurements of objects neighboring the source of interest.
 

### 2013-03-27 tex/ms.tex
more intro text

@@
 Nonetheless, some of the most important astronomical projects have produced enormously productive catalogs
-(the \textsl{SDSS}, for examle CITATION).
+(the \textsl{SDSS}, for example CITATION).
 There is a growing understanding that precise measurements of astronomical sources have to be made directly in the imaging data
@@
 
+To make this all concrete, consider the photometric data from the \textsl{ASAS} survey
+depicted in \figurename~\ref{fig:badfit} for one Mira variable.
+The photometric detections are shown and so are the epochs at which there are nulls (or non-detections).
+The standard fitting procedure, given censored data like these, is to \emph{ignore the null values}.
+When the nulls are ignored, the period fitting finds a short period.
+At that period, the model predicts the star to be bright at many phases in which the survey reported nulls.
+That is, although we have no idea how or why the nulls were generated,
+they effectively provide strong evidence against the short period.
+When we include the nulls (as described below), we get a longer period.
+This longer period is obviously better because it makes the star faint whenever nulls are common.
+This example clearly illustrates that the nulls are informative.
+In this paper we demonstrate how to automatically extract that information.
+
 The benefit of using the non-detections is large, but the cost is also large:
@@
 and large numbers of latent parameters to be inferred and marginalized out.
+In particular it involves modeling the data-generating process,
+which includes variable instrument and atmospheric conditions
+(leading to different effective detection thresholds)
+and also the software process by which catalog-entry decisons are made.
+In practice our models will be very simplified, but it is impossible to proceed \emph{without} significant modeling.
 
-To make this all concrete, consider the photometric data from the \textsl{ASAS} survey
-depicted in \figurename~\ref{fig:example} for one Mira variable.
-The photometric detections are shown and so are the epochs at which there are nulls (or non-detections).
-The standard
-
-...as an example, consider \figurename~\ref{fig:example}...
-
-\joey{What people actually do, what problems that could cause, and what people should do}
-
-A photometric survey (think of it as operating in a single photometric
-bandpass for now) scans over the celestial position of a particular
-variable star  a large number of times $t_i$.  At each of these
-times, the imaging data are analyzed by software, which
-treats each scan as a \emph{completely independent} survey.  That is,
-when analyzing the data from time $t_i$, none of the information from
-any of the other times is used in any way.
-
-At each time $t_i$, the star is either detected by the
-software or not.  When it is detected, the software returns
-a (possibly bad) flux value $f_i$ and (likely bad) uncertainty
-variance $s_i^2$.  When the source is \emph{not} detected,
-nothing is reported---in this case, because the software is not
-awesome, it doesn't see any reason to report anything at all, since on
-the non-detect passes, it has no inkling that this patch of the sky is
-interesting in any way.  Welcome to the desert of the real.
-
-Beyond this, there are two additional issues.  The first is that each
-night of imaging is different in an unreported and unknown way.  That
-is, there is different transparency, sky brightness, and point-spread
-function.  So the detection limit, or completeness level, or censoring
-of the catalog is different every night.  The second issue is that
-either because we are getting the data from a non-generous source, or
-because the conditions change rapidly and unpredictably, we can't
-analyze all the sources in a finite patch simultaneously.  For each
-variable star of interest, we \emph{only} get the data on that star
-itself.
+\hogg{either merge in or delete the material below this line}
 

### 2013-03-27 tex/ms.tex
notes to self
typesetting

@@
 % Note the status (in preparation, MNRAS submitted date, ApJ accepted date, in press etc).
-\slugcomment{DRAFT 2013-03-26 / in preparation}
+\slugcomment{DRAFT 2013-03-27 / in preparation}
 
@@
 Here we demonstrate with real data on periodic variables from the
-All-Sky Automated Survey (ASAS) that these null observations can
+\textsl{All-Sky Automated Survey} that these null observations can
 in fact be very useful in fitting and parameter estimation.
@@
 This set-up may seem artificial, but it is in fact the situation we find ourselves in for various surveys,
-including the ASAS survey, data from which we will model below.
+including the \textsl{ASAS} survey (CITATION), data from which we will model below.
 What we will show is that \emph{the ``null'' non-detections are in fact very informative};
@@
 
-\hogg{either merge in or delete the material below this line}
+\hogg{note to self: either merge in or delete the material below this line}
 

### 2013-03-28 tex/ms.tex
lots of figures added to manuscript

--- 2013-03-27 tex/ms.tex
@@
 
-Telescopes don't make catalogs, they make images (Hogg \& Lang 20xx).
-Nonetheless, some of the most important astronomical projects have produced enormously productive catalogs
-(the \textsl{SDSS}, for examle CITATION).
-There is a growing understanding that precise measurements of astronomical sources have to be made directly in the imaging data
-(CITE Lang, Anderson, something in weak lensing):
-It is impossible to pass the full information content of the imaging to the end user via a catalog
-(which by necessity is built with point estimates of the quantities of interest).
-
-The measurements of importance for variable-star science are relatively simple.
-Unlike the measurements used for astrometry, extragalactic transients, and weak lensing,
-where two-dimensional (morphological) information matters,
-when variable stars are being observed, all the end user wants is accurate photometry of each source at each epoch.
-One could hope that catalog-level information is sufficient:
-the detailed image morphology is not required for lightcurve extraction, except when fields are crowded.
-However, there are still limitations to working at catalog level:
-Catalog entries (nowadays) usually come with noise variance estimates,
-but they don't usually give detailed information about the locations or influence of
-bad detector pixels, cosmic rays, and flat-field variations;
-all these things might be needed to construct a precise photometric noise model or to understand residuals or artifacts.
-More importantly, catalogs are usually subject to hard \emph{censoring};
-the catalogs (nowadays) only include measurements that are deemed (by the software or catalog makers)
-to be \emph{signficant detections}.
-
-This could change.
-In principle a catalog could contain a measurement at \emph{every} epoch for any source that is detected at \emph{any} epoch.
-This kind of ``forced photometry'' would include many non-significant measurements,
-but it would come closer to passing forward all the information in the imaging to users working at catalog level.
-We recommend this approach to the catalog makers of the future (especially those making the \textsl{LSST} catalogs).
-
-Less good than the comprehensive forced-photometry approach is to report, when measurements are not significant,
-a well-defined ``upper limit'' for the brightness of each source.
-This is rarely done, and upper limits are hard to use, but upper limits are an improvement over standard practice.
-Standard practice is to \emph{report nothing at all} when a source is not detected significantly.
-Indeed, it is often hard, when presented with a multi-epoch catalog,
-to determine whether the reason a particular source is missing from the imaging at a particular epoch
-is because the source was faint or because the source fell outside the imaging field of view!
-
-In this contribution we look at improving catalog-level time-domain astronomy in the presence of \emph{censored data}.
-The problem we consider is that of period and amplitude modeling for large-amplitude variable stars in a data set
-in which \emph{no information is reported at all} at the epochs at which the star is not significantly detected.
-Furthermore, we will imagine that we also have \emph{no information at all} about the sensitivity of the telescope
-at the time of each non-detection; that is, we have no upper limits and we can't reconstruct them from any extant meta-data.
-The non-detections are literally database ``null'' values.
-This set-up may seem artificial, but it is in fact the situation we find ourselves in for various surveys,
-including the ASAS survey, data from which we will model below.
-What we will show is that \emph{the ``null'' non-detections are in fact very informative};
-amplitude and period fitting will be wrong in general if the non-detections are ignored.
-
-The benefit of using the non-detections is large, but the cost is also large:
-Use of these ``null'' data involves non-trivial modeling,
-and large numbers of latent parameters to be inferred and marginalized out.
-
-To make this all concrete, consider the photometric data from the \textsl{ASAS} survey
-depicted in \figurename~\ref{fig:example} for one Mira variable.
-The photometric detections are shown and so are the epochs at which there are nulls (or non-detections).
-The standard
-
-...as an example, consider \figurename~\ref{fig:example}...
+Surveys working in real-time tend to analyze each image with an eye to populating databases quickly.
+Unless a particular region of the sky gets special treatment (such as it is a known place of interest), discovery of variability happens by comparison of flux values at the database level.
+Analysis on such databases of transient/variables thus deals with censored data where there is often less information available than there should be.
+Even then, we might be able to infer what the thresholds were based on S/N measurements of objects neighboring the source of interest.
 
@@
   \begin{center}
-\includegraphics[angle=0,width=6.5in]{../plots/fig_045832-0604_1.png}
+\includegraphics[angle=0,width=6.5in]{fig_045832-0604_1.png}
   \end{center}
@@
 \begin{center}
-\includegraphics[angle=0,width=6.5in]{../plots/mira_simulated.pdf}
+$\begin{array}{c@{\hspace{-.3in}}c}
+\multicolumn{1}{l}{\mbox{\bf (a)}} & \multicolumn{1}{l}{\mbox{\bf (b)}} \\      \vspace{-.1in}
+\includegraphics[angle=0,width=3.5in]{fig_simulation_0_01_dat.png} & \includegraphics[angle=0,width=3.5in]{fig_simulation_0_02_dat.png} \\
+\multicolumn{1}{l}{\mbox{\bf (c)}} & \multicolumn{1}{l}{\mbox{\bf (d)}} \\
+\includegraphics[angle=0,width=3.5in]{fig_simulation_0_05_dat.png} & \includegraphics[angle=0,width=3.5in]{fig_simulation_0_1_dat.png}
+\end{array}$
+%\includegraphics[angle=0,width=6.5in]{../plots/mira_simulated.pdf}
 \end{center}
-\caption{ Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.
+\caption{ \joey{rewrite this caption!} Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.
 Blue tick marks along the bottom axis denote phases where there were non-detections.
@@
   \begin{center}
-    [\hogg{PUT FIGURE HERE}]
+$\begin{array}{c@{\hspace{-.3in}}c}
+\multicolumn{1}{l}{\mbox{\bf (a)} \hspace{0.7in} \textrm{ ASAS 034344+0655.5}} & \multicolumn{1}{l}{\mbox{\bf (b)}\hspace{0.7in}  \textrm{ ASAS 091617-2936.7}} \\      \vspace{-.1in}
+\includegraphics[angle=0,width=3.5in]{fig_034344+0655_5.png} & \includegraphics[angle=0,width=3.5in]{fig_091617-2936_7.png} \\
+\multicolumn{1}{l}{\mbox{\bf (c)}\hspace{0.7in}  \textrm{ ASAS 075855-1914.1}} & \multicolumn{1}{l}{\mbox{\bf (d)}\hspace{0.7in}  \textrm{ ASAS 075826-4019.8}} \\
+\includegraphics[angle=0,width=3.5in]{fig_075855-1914_1.png} & \includegraphics[angle=0,width=3.5in]{fig_075826-4019_8.png}
+\end{array}$
   \end{center}
-  \caption{A real-data example.  This is the same source as that shown in \figurename~\ref{fig:badfit} but now showing the results of...\label{fig:goodfit}}
+  \caption{ \joey{add caption: period doubling} A real-data example.  This is the same source as that shown in \figurename~\ref{fig:badfit} but now showing the results of...\label{fig:goodfit}}
 \end{figure}
@@
   \begin{center}
-    [\hogg{PUT FIGURE HERE}]
+$\begin{array}{c@{\hspace{-.3in}}c}
+\multicolumn{1}{l}{\mbox{\bf (a)} \hspace{0.7in} \textrm{ ASAS 112811-6606.3}} & \multicolumn{1}{l}{\mbox{\bf (b)}\hspace{0.7in}  \textrm{ ASAS 065002-4554.6}} \\      \vspace{-.1in}
+\includegraphics[angle=0,width=3.5in]{fig_112811-6606_3.png} & \includegraphics[angle=0,width=3.5in]{fig_065002-4554_6.png} \\
+\multicolumn{1}{l}{\mbox{\bf (c)}\hspace{0.7in}  \textrm{ ASAS 092021-5815.2}} & \multicolumn{1}{l}{\mbox{\bf (d)}\hspace{0.7in}  \textrm{ ASAS 082012-3816.7}} \\
+\includegraphics[angle=0,width=3.5in]{fig_092021-5815_2.png} & \includegraphics[angle=0,width=3.5in]{fig_082012-3816_7.png}
+\end{array}$
   \end{center}
-  \caption{Another few real-data examples; in each case the method presented here either matches the traditional method or is substantially better.\label{fig:examples}}
+  \caption{\joey{add caption: period tripling} Another few real-data examples; in each case the method presented here either matches the traditional method or is substantially better.\label{fig:examples}}
 \end{figure}

### 2013-03-28 py/censoring2.py

@@ -237,7 +237,8 @@ class Censored:
 
 
 
-    def plot(self, ax, par, fold = False, plot_model = True, mag = False, plot_title = True):
+    def plot(self, ax, par, fold = False, plot_model = True, mag = False, plot_title = True,\
+             period = None):
         '''
         input:
         - ax: matplotlib axes object
@@ -317,7 +318,7 @@ class Censored:
             #ax.plot(xp - mediant, mup_m, 'b-', alpha=0.3)
         ax.set_xlim(tlim - mediant)
         if(mag):
-            ax.set_ylim(np.max(y) + 1., np.min(y) - 0.5)
+            ax.set_ylim(np.max(y) + 0.5, np.min(y) - 0.5)
             ax.set_ylabel(r'magnitude $m$ (mag)')
         else:
             foo = np.max(y + ey)
@@ -329,6 +330,11 @@ class Censored:
             ax.set_xlabel(r'time $t$ (MJD $-$ %d d)' % mediant)
         if(plot_title):
             ax.set_title(self.name)
+        if(period != None):
+            ax.text(0.05, 0.95,'P = %.1f d' % period,
+                    horizontalalignment='left',
+                    verticalalignment='center',
+                    transform = ax.transAxes)
         return None
 
 

### 2013-03-28 py/make_lc_figs.py

@@ -22,12 +22,13 @@ from lomb_scargle_censor import lomb as lombc
 
 from fit_censoring import load_lc_from_web, periods_lomb, period_lombc
 
-def plot_fig(ID, path, nper=2):
+def plot_fig(ID, path, nper=2, simulation=False):
     """ plot folded and unfolded LS and censored model fits
     """
-    data = load_lc_from_web(ID)
-
-    print path + 'plots/fig_'+ ID +'.png'
+    if(simulation):
+        data = load_sim(path+ID)
+    else:
+        data = load_lc_from_web(ID)
 
     indo = np.where(data['m'] != 29.999)
     tobs = data['t'][indo]
@@ -90,11 +91,11 @@ def plot_fig(ID, path, nper=2):
     fig.set_size_inches(12.0,8.0)
     # unfolded LS
     ax = plt.subplot(221)
-    cmodel.plot(ax, orig, fold=False, plot_model=True, mag=True, plot_title = False)
+    cmodel.plot(ax, orig, fold=False, plot_model=True, mag=True, plot_title = False, period=orig[0])
     ax.set_title("Traditional Method",fontsize=20)
     # unfolded censored
     ax2 = plt.subplot(222)
-    cmodel.plot(ax2, pstar, fold=False, plot_model=True, mag=True, plot_title = False)
+    cmodel.plot(ax2, pstar, fold=False, plot_model=True, mag=True, plot_title = False, period=pstar[0])
     ax2.set_title("Our Method",fontsize=20)
     # folded LC
     ax3 = plt.subplot(223)
@@ -102,18 +103,39 @@ def plot_fig(ID, path, nper=2):
     # folded censored
     ax4 = plt.subplot(224)
     cmodel.plot(ax4, pstar, fold=True, plot_model=True, mag=True, plot_title = False)
-    fn = path + 'plots/fig_'+ ID +'.png'
+    
+    if(simulation):
+        fn = path + 'plots/fig_simulation_'+ ID.split('_')[-1].replace('.','_') +'.png'
+    else:
+        fn = path + 'plots/fig_'+ ID +'.png'
+        
     print 'Writing %s' % fn
     plt.savefig(fn, dpi=200)
 
     return pstar
 
 
+def load_sim(fpath):
+    """ read in simulated light curve and return dict of t, m, e """
+    data = np.loadtxt(fpath, dtype={'names': ('t', 'm', 'e'),
[…]

### 2013-03-29 tex/ms.tex
wrote section 3

@@
 Telescopes don't make catalogs, they make images (Hogg \& Lang 20xx).
-Nonetheless, some of the most important astronomical projects have produced enormously productive catalogs
+Nonetheless, some of the most important astronomical projects have produced enormously valuable catalogs
 (the \textsl{SDSS}, for example CITATION).
@@
 the detailed image morphology is not required for lightcurve extraction, except when fields are crowded.
-However, there are still limitations to working at catalog level:
+However, there are still limitations to working at the catalog level:
 Catalog entries (nowadays) usually come with noise variance estimates,
-but they don't usually give detailed information about the locations or influence of
+but they do not typically give detailed information about the locations or influence of
 bad detector pixels, cosmic rays, and flat-field variations;
@@
 the catalogs (nowadays) only include measurements that are deemed (by the software or catalog makers)
-to be \emph{signficant detections}.
+to be \emph{significant detections}.
 
@@
 but it would come closer to passing forward all the information in the imaging to users working at catalog level.
-We recommend this approach to the catalog makers of the future (especially those making the \textsl{LSST} catalogs).
+We recommend this approach to the catalog makers of the future (especially those constructing the \textsl{LSST} catalogs).
 
@@
 This set-up may seem artificial, but it is in fact the situation we find ourselves in for various surveys,
-including the \textsl{ASAS} survey (CITATION), data from which we will model below.
+including the All-Sky Automated Survey (\textsl{ASAS}, \citealt{2001ASPC..246...53P}), data from which we will model below.
 What we will show is that \emph{the ``null'' non-detections are in fact very informative};
@@
 
-To make this all concrete, consider the photometric data from the \textsl{ASAS} survey
-depicted in \figurename~\ref{fig:badfit} for one Mira variable.
-The photometric detections are shown and so are the epochs at which there are nulls (or non-detections).
+To make this more concrete, consider the photometric data from the \textsl{ASAS} survey,
+depicted in \figurename~\ref{fig:badfit} for a single Mira variable.
+The photometric detections are plotted and so are the epochs at which there are nulls (non-detections).
 The standard fitting procedure, given censored data like these, is to \emph{ignore the null values}.
-When the nulls are ignored, the period fitting finds a short period.
-At that period, the model predicts the star to be bright at many phases in which the survey reported nulls.
-That is, although we have no idea how or why the nulls were generated,
-they effectively provide strong evidence against the short period.
-When we include the nulls (as described below), we get a longer period.
-This longer period is obviously better because it makes the star faint whenever nulls are common.
+When the nulls are ignored, the period-fitting method finds a short period.
+At that period, the model predicts the star to be bright at many phases in which the survey reported null observations.
+However, the presence of a null observation means that it is likely that the source brightness has dipped below the detection limit at that epoch.
+%That is, although we have no idea how or why the nulls were generated,
+%they effectively provide strong evidence against the short period.
+When we include the nulls in our model (as described in \S\ref{sec:model}), we find that the best fit is a longer-period, higher-amplitude model.
+This longer period is obviously better because it models the star to be faint at epochs where nulls commonly occur.
 This example clearly illustrates that the nulls are informative.
-In this paper we demonstrate how to automatically extract that information.
+In this paper we demonstrate how to automatically incorporate that information in a rigorous statistical model.
 
@@
 (leading to different effective detection thresholds)
-and also the software process by which catalog-entry decisons are made.
-In practice our models will be very simplified, but it is impossible to proceed \emph{without} significant modeling.
+and also the software process by which catalog-entry decisions are made.
+In practice our models will be very simplified, but it is impossible---in the presence of such noisy and imperfect data---to proceed \emph{without} significant modeling.
 
@@
 
-Replacing likelihood with fast version where we trust $s_i^2$ for observed data.
+Replacing likelihood with fast version where we trust $s_i^2$ for observed data. \joey{I need to finish this}
 
-\section{Experiments}
+\joey{we should include a link to the git repo}
+
+\section{Experiment: Period Estimation for Mira Variables}
 \label{sec:experiments}
+
+In this section, we compare  the methodology from \S\ref{sec:model} to the traditional fitting method that does not incorporate null observations (the generalized Lomb-Scargle periodogram; \citealt{1976lomb, 1982scar, zech09}).
+The goal in the simulation study is to compare the abilities of the two methods for period and amplitude estimation on a series of simulated Mira variables.
+Of particular interest is to gauge the efficacy of each fitting method when applied to light curves at different observed (median) brightness levels.
+
+\subsection{Simulating Faint Mira Variables}
+\label{ss:mirasim}
+
+In this first experiment, we begin with a well-observed Mira variable from the ASAS Catalog of Variable Stars (ACVS, \citealt{acvs}) , ASAS 235627-4947.2.
+This star has a pulsation period of 266.5245 days and a Lomb-Scargle amplitude of 2.36 mag ($V$ band) based on 408 epochs of observations with zero null detections.
+The median observed brightness of the ASAS light curve is $V = 10.4$.
+
+We will employ the ASAS observations of this bright, well-observed star to simulate what the same star would look like at different median observed brightness levels.
+As the star's observed brightness is artificially changed to fainter and fainter values, its simulated ASAS light curve will contain a larger and larger number of null observations.
+Effectively, artificially dimming the star decreases the signal to noise ratio of each observation, which will cause some of the original detections to change into non-detections.
+
+To simulate faint Mira stars from ASAS 235627-4947.2, we use the following procedure:
+\begin{enumerate}
+\item Convert the observed magnitudes $m_i$ and uncertainty variances $s^2_{m,i}$ to fluxes $f_i$ and flux uncertainty variances $s^2_{f,i}$ (we assume a $V$-band zero-point of $3.67\times10^{-9}$ erg/s/cm$^2$/\AA).
+\item For a given flux dimming parameter, $d$, sample the new fluxes, $\tilde{f}_i$, from a Gaussian distribution centered around $f_i / d$ with standard deviation sampled from the empirical ASAS distribution of $s^2_{f,i} | f_i$.
+We simulate 10 different Mira variables by employing different values of $d$, ranging from 1 to 500.
+\item Flux measurements of the `dimmed' light curve that are not at least 5$\sigma$ above zero are denoted as non-detections and their flux estimates (and errors) are censored and nulled out.
+\end{enumerate}
+Folded and unfolded versions of these simulated light curves are plotted in Figure \ref{fig:mirasim}.
+
+\subsection{Results of Fitting Simulations}
+
+We apply both the traditional method and our censored fitting method to each of the 10 simulated light curves.
+Fits to four of the simulations are plotted in Figure \ref{fig:mirasim}.
+In three of the four simulations, the traditional method estimates an incorrect period whereas our method recovers the true period, to within $\sim1 \%$, in each case.
+Clearly, the censoring method works well because it incorporates the abundant null observations, effectively discovering the period which best places them  at phases for which the light-curve model dips below the inferred detection limit.
+Remarkably, even for the faintest simulation ($V=16.7$), our method estimates the correct period even though only 7 epochs of observations were recorded (Figure \ref{fig:mirasim}d).
+This demonstrates the value in utilizing the null detections: whereas the traditional method completely ignores the information in the 401 null non-detections, our model fully incorporates it when determining the best-fit period and sinusoid model.
+
+Figure \ref{fig:mirasimparams} depicts the estimates of the period and amplitude of the Mira star as a function of the median $V$-band magnitude of the simulated source.
+Whereas our method recovers the correct period for each simulation, the traditional period-finding method only works well for the five brightest simulations ($V \lesssim 14$).
+For fainter magnitudes, the traditional method mistakenly detects a period of one-half or one-third the value of the true period.
+The story for amplitude estimation is similar: even for the faintest stars, the estimates of our method are reasonably well-behaved (all within 1 mag of the true amplitude) whereas the amplitude estimates from the traditional method quickly fall to very small values as the brightness of the simulation decreases.
+
+  It is important to note that our fitting method typically takes longer than the traditional generalized Lomb-Scargle periodogram to fit a light curve.
+The amount of compute time for our method depends heavily on the number of null observations.
+For example, on a single core of a MacBook Pro with 2.9 GHz Intel Core and 8 GB of memory, our method takes 291 seconds for the faintest simulation (401 null detections), 80 seconds for the simulation at 12.8th magnitude (105 null detections) and 0.24 seconds for the brightest simulation (0 null detections).
+For comparison, the traditional method takes less than 0.1 seconds in each case.
+
+\section{Application to Mira Light Curves in ASAS}
+\label{sec:results}
+
+We use the methodology of \cite{2011arXiv1106.2832R} to select the top ASAS Mira  and RR Lyrae, Fundamental Mode candidates.
+Using a Mira classification probability threshold of 0.8 gives us 1720 Mira  and 1029 RR Lyrae candidates.
 
@@
 
-\subsection{Simulating Faint Mira Variables}
-\label{ss:mirasim}
-
-In this first experiment, we begin with a well-observed Mira variable from the ASAS Catalog of Variable Stars (ACVS, \citealt{acvs}) , ASAS 235627-4947.2.
-This star has a pulsation period of 266.6286 days and a Lomb-Scargle amplitude of 2.38 mag (V band).
-Note there are no non-detections in the ASAS light curve.
-
-To simulate faint Mira stars from ASAS 235627-4947.2, we use the following procedure:
-\begin{enumerate}
-\item Convert the observed magnitudes $m_i$ and uncertainty variances $s^2_{m,i}$ to fluxes $f_i$ and flux uncertainty variances $s^2_{f,i}$ (we assume a V-band zero-point of $3.67\times10^{-9}$ erg/s/cm$^2$/\AA)
-\item For a given flux dimming parameter, $d$, sample the new fluxes, $\tilde{f}_i$, from a Gaussian distribution centered around $f_i / d$ with standard deviation sampled from the empirical ASAS distribution of $s^2_{f,i} | f_i$.
-\joey{I found that relationship to be linear with a slope $\approx 1$, which is probably BS since the ASAS mag errors are all crap.
-So I added a reference value to each flux error to ensure a mag limit $\approx$ 14.5 mag.}
-\item Fluxes that are not at least 5$\sigma$ above zero are denoted as non-detections and their flux estimates (and errors) are censored.
-\end{enumerate}
-Folded light curves of ASAS 235627-4947.2, dimmed by four different values of $d$, are plotted in Figure \ref{fig:mirasim}.
-
-\subsection{Results of Fitting Miras}
-
-\section{Results: ASAS Light Curves}
-\label{sec:results}
-
-We use the methodology of \cite{2011arXiv1106.2832R} to select the top ASAS Mira  and RR Lyrae, Fundamental Mode candidates.
-Using a posterior probability threshold of 0.8 gives us 1720 Mira  and 1029 RR Lyrae candidates.
-
  \subsection{Mira Variables}
@@
 Results show that XYZ.
-
- \subsection{RR Lyrae Variables}
-
-  Show P - A relationship before and after using the method
-
-There is an RRL P - A relationship (strong linear anti-correlation)
-
-As a selection criterion (simulated data?)
 
@@
 $\begin{array}{c@{\hspace{-.3in}}c}
-\multicolumn{1}{l}{\mbox{\bf (a)}} & \multicolumn{1}{l}{\mbox{\bf (b)}} \\      \vspace{-.1in}
-\includegraphics[angle=0,width=3.5in]{fig_simulation_0_01_dat.png} & \includegraphics[angle=0,width=3.5in]{fig_simulation_0_02_dat.png} \\
-\multicolumn{1}{l}{\mbox{\bf (c)}} & \multicolumn{1}{l}{\mbox{\bf (d)}} \\
-\includegraphics[angle=0,width=3.5in]{fig_simulation_0_05_dat.png} & \includegraphics[angle=0,width=3.5in]{fig_simulation_0_1_dat.png}
+\multicolumn{1}{l}{\mbox{\bf (a)} \hspace{0.1in} V=16.7} & \multicolumn{1}{l}{\mbox{\bf (b)}\hspace{0.1in} V=15.5} \\      \vspace{-.1in}
+\includegraphics[angle=0,width=3.5in]{fig_simulation_0_002_dat.png} & \includegraphics[angle=0,width=3.5in]{fig_simulation_0_007_dat.png} \\
+\multicolumn{1}{l}{\mbox{\bf (c)}\hspace{0.1in} V=14.6} & \multicolumn{1}{l}{\mbox{\bf (d)}\hspace{0.1in} V=13.6} \\
+\includegraphics[angle=0,width=3.5in]{fig_simulation_0_02_dat.png} & \includegraphics[angle=0,width=3.5in]{fig_simulation_0_05_dat.png}
 \end{array}$
@@
 \end{center}
-\caption{ \joey{rewrite this caption!} Folded light curves for data simulated from the Mira variable ASAS 235627-4947.2 using the prescription in \S\ref{ss:mirasim}.
-Blue tick marks along the bottom axis denote phases where there were non-detections.
-As the original flux measurements are dimmed by higher factors, the troughs of the sinusoidal light curve are censored, resulting in an incomplete view of the data.
-\hogg{This figure needs to show the OLD and NEW best-fit lightcurve models too!}\label{fig:mirasim}}
+\caption{ Results of period fitting analysis for 4 of the Mira simulations based on the ASAS 235627-4947.2 light curve.
+In each figure, the left-hand column shows the result of the traditional fitting method and the right-hand column depicts the result of our method.
+For each of the simulations shown, our method discovers the true period (to within $\sim 1\%$), even down to a median Mira magnitude of $V=16.7$ (panel a).
+On the other hand, the traditional fitting method finds a wrong period for all but the brightest simulation plotted ($V=13.6$, panel d).
+Likewise, the amplitude estimates for our method are clearly better than those of the traditional method.
+Even with as few as 7 observed epochs of data, our method successfully recovers the true period because it utilizes the 400 null observations.
+\label{fig:mirasim}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\begin{center}
+\includegraphics[angle=0,width=6.5in]{simulation_parameter_estimates.png}
+\end{center}
+\caption{  Period and amplitude estimates of simulated Mira variables.
+Top: Our method (dashed blue line) recovers the true period for each simulated brightness level while the traditional method fails for the five faintest simulations.
+The dashed gray line denotes the `true' period, determined by analysis on the original, well-sampled Mira.
+Bottom: For amplitude estimation, our method determines values which are much closer to the true value (dashed gray line), even for faint Mira simulations.
+Because it ignores the null observations, the traditional method recovers  amplitudes that are much too small for faint Miras.
+\label{fig:mirasimparams}}
 \end{figure}

### 2013-03-29 py/simulate_LCs.R

@@ -1,0 +1,119 @@
+
+## simulated mocked up light curves w/non-detections from real, well-observed LCs
+
+##########################################################################################
+## functions
+
+# magnitude to flux conversion
+mag2flux = function(mag,magerr){
+  flux.zp = 3.67*10^(-9) # V-band flux zero-point
+  mag.zp = 2.5*log10(flux.zp)
+  flux = 10^((mag.zp-mag)/2.5)
+  fluxerr = sqrt(magerr^2 * (flux*log(10)*(1/2.5))^2)
+  return(list(flux=flux,fluxerr=fluxerr))
+}
+
+# flux to mag conversion
+flux2mag = function(flux,fluxerr){
+  flux.zp = 3.67*10^(-9) # V-band flux zero-point
+  mag.zp = 2.5*log10(flux.zp)
+  mag = -2.5 * log10(flux) + mag.zp
+  magerr = sqrt(fluxerr^2 * (-2.5/(flux*log(10)))^2)
+  return(list(mag=mag,magerr=magerr))
+}
+
+
+################# create simulated LC from existing LC
+
+makeSim = function(time,flux,fluxerr,p,sigmodel,eps=1e-16){
+  n=length(flux)
+  muflux = flux*p
+  pred.err = predict(sigmodel,newdata=data.frame(flux=log10(muflux)),se.fit=TRUE)
+  mufluxerr = pred.err$fit
+  sigfluxerr = pred.err$se.fit
+
+  fluxerrnew = 10^rnorm(n,mean=mufluxerr,sd=sigfluxerr)+eps
+  fluxnew = rnorm(n,mean=flux*p,sd=fluxerrnew)
+
+  nondetect = ifelse(fluxnew - 5*fluxerrnew <=0,1,0)
+  
+  return(list(flux=fluxnew,fluxerr=fluxerrnew,nondetect=nondetect))
+}
+
+#####################################################################################
+
+#####################################################################################
+################# miras
+miras = system("ls /Users/jwrichar/Documents/CDI/LS_upperlim/miras/",intern=TRUE)
+for(ii in 1:length(miras)){
[…]

### 2013-03-29 tex/ms.tex
finished intro, discussion notes

@@
 
-\hogg{note to self: either merge in or delete the material below this line}
-
-\joey{RRL end at 120 kpc in Sesar et al. paper}
-
-\joey{discuss in context of: parameter estimation (amplitude / period), discovery (we're gonna find a lot more variable sources if we take the non-detections into account), and classification; lossiness}
-
-\joey{let's make a point to stress variable discovery}
-
-\joey{Numbers of LCs affected: Stripe 82, ASAS}
+In what follows, we concentrate on parameter estimation---fitting of amplitudes and periods to noisy censored data%
+---but because modeling, parameter estimation, discovery, and classification are intertwined,
+this work has implications in all these areas.
+For example, we often conclude that a star \emph{is} a periodic variable because parameter estimation delivers a solid period.
+We often conclude that a star is a Mira or a RR Lyrae star because it's best-fit period or lightcurve shape is in some range.
+For example, the valuable RR Lyrae catalog generated with \textsl{SDSS Stripe 82} data (CITE SESAR) classified stars
+by performing parameter estimation and then cutting on derived parameters.
+This study worked in the traditional mode of ignoring censored data, thereby limiting its ability to find the faintest
+(and therefore most distant) RR Lyrae stars.
+A model for the censored data might have extended the reach of this survey and led to new discoveries of Milky Way structure.
+\hogg{Joey: add words here about your own variable-star catalog and what could be improved.}
 
@@
 
-What people can do, starting from the data-taking procedure.
+\hogg{This section should include limitations of what we have done:  We didn't vary all available model parameters.
+Our code is slow.
+We don't permit time dependence to the censoring model.
+Our noise model is naive at best.}
 
-limitations
+\hogg{This section should return to the detection point and the promise of that.}
+
+\hogg{This section should NOT be a summary of the paper:  That lives in the abstract!}
 

### 2013-03-29 tex/ms.tex
discussion section start

@@
 \label{sec:discussion}
+% This section should NOT be a summary of the paper:  That lives in the abstract!
 
-\hogg{This section should include limitations of what we have done:  We didn't vary all available model parameters.
-Our code is slow.
-We don't permit time dependence to the censoring model.
-Our noise model is naive at best.}
+We have shown that even the apparently ``null'' information that a
+source was not detected can nonetheless be very informative, if the
+generation of the nulls is \emph{modeled}.  This is true even in the
+worst-case scenario that telescope sensitivity is not known or
+derivable from the data or any meta data.  The inclusion of the null
+data can have important impacts on parameter estimation; in particular
+we have shown that it improves variable-star period and amplitude
+estimation in heavily censored data sets.
 
-\hogg{This section should return to the detection point and the promise of that.}
+Modeling the nulls brings great benefit.  It comes at cost, however:
+Optimizing the marginalized likelihood (marginalizing out the latent
+parameters related to the censoring) is computationally expensive.
+It would be even more expensive if we had permitted all the
+variables in our graphical model (\figurename~\ref{fig:pgm}) to vary.
+Our method is far slower than the traditional method of ignoring the
+nulls and performing linear least-squares fitting.
 
-\hogg{This section should NOT be a summary of the paper:  That lives in the abstract!}
+There are also various limitations to our approach, or aspects of what
+we have presented here that could use improvement.  One is that our
+noise model and censoring model were both extremely naive.  We
+effectively assumed that the censoring process was related very
+directly to signal-to-noise; this might not be the case in many
+experiments.  In particular, sources are often missed because of
+detector issues or chip gaps or other hardware-related reasons.  These
+nulls could be modeled by our model without modification, but it would
+be a ``bad fit'' to the true data-generating process.  A better model
+might be a ``mixture'' model that permits some censoring to be caused
+by more arbitrary processes.
+
+We also assumed that the censoring process is constant in time.  This
+is not true in general; in most optical systems the censoring would
+depend (at least in part) on the weather, and the weather depends on
+season and year.  It would be straightforward to add parameters that
+permit the censoring to depend on season and year.
+
+\hogg{This section should return to the detection point and the promise of that.  Can we give an example in ASAS of a de novo detection?}
 
